### 2013-09-27 article-models-defn.tex
resolve issue #7

\section{A variable-free presentation of type theory}
We have seen in the example of $\UU$ that we will not need to refer to variables
when we are manipulating contexts. In fact, it is more natural not to.
This suggests that there is a presentation
of type theory where contexts do not mention variables. The first small gain is that such a presentation
would not be burdened with comments about variables being bounded or not, or fresh or
not occuring at all.

\subsection{The basic judgments}
The judgments we can make are the standard six basic judgments involving contexts, types and terms and
judgmental equality in each of these three cases.
\begin{align*}
\jctx*{\Gamma} & \jctxeq*{\Gamma}{\Gamma'}\\
\jtype*{\Gamma}{A} & \jtypeeq*{\Gamma}{A}{B}\\
\jterm*{\Gamma}{A}{x} & \jtermeq*{\Gamma}{A}{x}{y}.
\end{align*}

In contrast with standard practise, we don't assume that there is an empty context. The idea is that
if we interpret only the syntax of contexts, types and terms with extension, weakening and
substitution, we do not have a full model of type theory but we still have a 
structure resembling a category and categories need not have terminal objects. Therefore,
the rules that follow will describe only how to manipulate contexts, types and terms.
Of most of the inferences that we give there are two versions: one introducing an operation, the other asserting that the operation in question preserves judgmental equality.

\subsection{The basic rules for judgmental equality}
The rules for judgmental equality establish that it is an equivalence relation
in all three cases (contexts, types and terms).
%\begingroup
%\renewcommand*{\arraystretch}{3}
%\begin{equation*}
%\begin{array}{ccc}
\begin{infarray}{ccc}
\inference{\jctx{\Gamma}}{\jctxeq{\Gamma}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}}{\jctxeq{\Delta}{\Gamma}} & \inference{\jctxeq{\Gamma}{\Delta}\qquad\jctxeq{\Delta}{\Theta}}{\jctxeq{\Gamma}{\Theta}}\\
\inference{\jtype{\Gamma}{A}}{\jtypeeq{\Gamma}{A}{A}} &
\inference{\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Gamma}{B}{A}} & 
\inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\Gamma}{B}{C}}{\jtypeeq{\Gamma}{A}{C}}\\
\inference{\jterm{\Gamma}{A}{x}}{\jtermeq{\Gamma}{A}{x}{x}} & 
\inference{\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{A}{y}{x}} &
\inference{\jtermeq{\Gamma}{A}{x}{y}\qquad\jtermeq{\Gamma}{A}{y}{z}}{\jtermeq{\Gamma}{A}{x}{z}}
\end{infarray}
%\end{array}
%\end{equation*}
\begin{infarray}{cc}
\inference{\jctxeq{\Gamma}{\Delta}\qquad\jtype{\Gamma}{A}}{\jtype{\Delta}{A}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jtypeeq{\Delta}{A}{B}}\\
\inference{\jctxeq{\Gamma}{\Delta}\qquad\jterm{\Gamma}{A}{x}}{\jterm{\Delta}{A}{x}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Delta}{A}{x}{y}}\\
\inference{\jtypeeq{\Gamma}{A}{B}\qquad \jterm{\Gamma}{A}{x}}{\jterm{\Gamma}{B}{x}}
& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtermeq{\Gamma}{A}{x}{y}}{\jtermeq{\Gamma}{B}{x}{y}}
\end{infarray}

\subsection{Extension}
We introduce extension which not only extends a context $\Gamma$ and a type
$A$ over it to a context $\ctxext{\Gamma}A$, but which also extends a type $A$
in context $\Gamma$ and a family $P$ over it to a type $\ctxext{A}P$ in context
$\Gamma$. We do this to ensure that all of type theory can be done in a context.
For instance, we could say (1) that a context in context $\Gamma$ is the same thing
as a type in context $\Gamma$; (2) When $A$ is a context in this sense, a type in
context $A$ is the same thing as a family $P$ over $A$ and (3) when $P$ is a type
in context $A$ in this sense, a term of $P$ keeps its original meaning.

Note that by introducing extension on the level of types and families,
we introduce $\Sigma$-types at a very early stage.
However, we need substitution to make this precise.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}}{\jctx{\ctxext{\Gamma}A}}
& \inference{\jctxeq{\Gamma}{\Delta}\qquad\jtypeeq{\Gamma}{A}{B}}{\jctxeq{\ctxext{\Gamma}A}{\ctxext{\Delta}B}}\\
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\ctxext{A}P}}
& \inference{\jtypeeq{\Gamma}{A}{B}\qquad\jtypeeq{\ctxext{\Gamma}A}{P}{Q}}{\jtypeeq{\Gamma}{\ctxext{A}P}{\ctxext{B}Q}}
\end{infarray}

\subsection{Weakening}
We first define weakening by a context $\Gamma$. Since weakening by $\Gamma$
should in principle be a functor, it acts on contexts, types and terms alike.
Note that it is because of the weakening $\ctxwk{\Gamma}{\Delta}$ of a context
$\Delta$ by $\Gamma$ that we can speak of context morphisms from $\Gamma$ to $\Delta$: they are the terms
of $\ctxwk{\Gamma}{\Delta}$.
\begin{infarray}{cc}
\inference{\jctx{\Gamma}\qquad\jctx{\Delta}}{\jtype{\Gamma}{\ctxwk{\Gamma}{\Delta}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jctxeq{\Delta}{\Delta'}}{\jtypeeq{\Gamma}{\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma'}{\Delta'}}}\\
\inference{\jctx{\Gamma}\qquad\jtype{\Delta}{B}}{\jtype{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}} & \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtypeeq{\Delta}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma'}{B'}}}\\
\inference{\jctx{\Gamma}\qquad\jterm{\Delta}{B}{y}}{\jterm{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}} 
& \inference{\jctxeq{\Gamma}{\Gamma'}\qquad\jtermeq{\Delta}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}\ctxwk{\Gamma}{\Delta}}{\ctxwk{\Gamma}{B}}{\ctxwk{\Gamma}{y}}{\ctxwk{\Gamma'}{y'}}} 
\end{infarray}

A weakening operation is also defined for types. When $A$ and $B$ are both types
 in context $\Gamma$, the weakened type $\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$
 is the family which `doesn't really depend on $A$'. The terms of $\ctxwk{A}{B}$
 are the functions from $A$ to $B$. Likewise, the terms $\ctxwk{A}y$ are the
 constant maps at $y$ from $A$ to $B$, for $y:B$.
\begin{infarray}{cc}
\inference{\jtype{\Gamma}{A}\qquad\jtype{\Gamma}{B}}{\jtype{\ctxext{\Gamma}A}{\ctxwk{A}{B}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\Gamma}{B}{B'}}{\jtypeeq{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jtype{\ctxext{\Gamma}B}{Q}}
{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtypeeq{\ctxext{\Gamma}B}{Q}{Q'}}
{\jtypeeq{\ctxext({\Gamma}{A})\ctxwk{A}{B}}{\ctxwk{A}{Q}}{\ctxwk{A'}{Q'}}}\\
\inference{\jtype{\Gamma}{A}\qquad\jterm{\Gamma}{B}{y}}{\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A}{y}}}
& \inference{\jtypeeq{\Gamma}{A}{A'}\qquad\jtermeq{\Gamma}{B}{y}{y'}}{\jtermeq{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{\ctxwk{A}{y}}{\ctxwk{A'}{y'}}}
\end{infarray}

\subsection{Substitution}
\begin{infarray}{cc}
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\subst{x}{P}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext{\Gamma}A}{P}{P'}}{\jtypeeq{\Gamma}{\subst{x}{P}}{\subst{x'}{P'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jtype{\ctxext({\Gamma}{A})P}{Q}}{\jtype{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtypeeq{\ctxext({\Gamma}{A})P}{Q}{Q'}}{\jtypeeq{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x'}{Q'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext{\Gamma}A}{P}{f}}{\jterm{\Gamma}{\subst{x}{P}}{\subst{x}{f}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext{\Gamma}A}{P}{f}{f'}}{\jtermeq{\Gamma}{\subst{x}{P}}{\subst{x}{f}}{\subst{x'}{f'}}}\\
\inference{\jterm{\Gamma}{A}{x}\qquad\jterm{\ctxext({\Gamma}{A}){P}}{Q}{g}}{\jterm{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{g}}}
& \inference{\jtermeq{\Gamma}{A}{x}{x'}\qquad\jtermeq{\ctxext({\Gamma}{A})P}{Q}{g}{g'}}{\jtermeq{\ctxext{\Gamma}\subst{x}{P}}{\subst{x}{Q}}{\subst{x}{g}}{\subst{x'}{g'}}}
\end{infarray}

\subsection{Compatibility of extension, weakening and substitution}
We describe the compatibility of extension, weakening and substitution. These
compatibility rules are formulated in a weak sense.

Extension is compatible with extension:
\begin{infarray}{c}
\inference
{ \jtype{\ctxext{\Gamma}A}{P}
}{\jterm
  {\ctxext{\Gamma}({A}P)}
  {\ctxwk{\ctxext{\Gamma}({A}P)}{(\ctxext({\Gamma}{A})P)}}{\typefont{extext}_0(A,P)}
}\\
\inference
{ \jtype{\ctxext({\Gamma}{A})P}{Q}
}{\jterm
  {\ctxext{\Gamma}({A}({P}Q))}
  {\ctxwk{\ctxext{A}({P}Q)}{(\ctxext({A}{P})Q)}}
  {\typefont{extext}_1(A,P,Q)}
}
\end{infarray}

Weakening is compatible with extension:
\begin{infarray}{c}
\inference{
  \jctx{\Gamma}
  \qquad
  \jtype{\Delta}{B}}
  {\jterm
    {\ctxext{\Gamma}\ctxwk{\Gamma}{(\ctxext{\Delta}B)}}
    {\ctxwk{\ctxwk{\Gamma}{(\ctxext{\Delta}B)}}{\ctxext{(\ctxwk{\Gamma}{\Delta})}(\ctxwk{\Gamma}{B})}}
    {\typefont{wkext}_0(\Gamma,B)}}
\end{infarray}

\subsection{More stuff in the type theory}
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{A}}{\idfunc[A]}}
\end{infarray}

\subsection{Leibniz equality on terms}
Occasionally we shall need a weaker version of equality than judgmental equality.
Our choice here is Leibniz equality, since it implies propositional equality when
identity types are present.

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$.
We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctxext{\Gamma}A$ and thus we have the inference
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\ctxwk{A}{P}}{\ctxwk{\ctxwk{A}{P}}{(\ctxwk{\ctxwk{A}{A}}{P})}}}
\end{equation*}
This family plays the role of the family $\jtype{\Gamma,x,y:A}{P(x)\to P(y)}$
in the type theory of \cite{TheBook}.

\subsection{Trivial cofibrations and weak equivalences of types}
We describe a relation between types that expresses when they are weakly equivalent.
Weak equivalence is introduced because we need a weaker notion of judgmental 
equality which also makes sense when identity types are not present, since that
would allow us to state that context extension, weakening and substitution
commute with each other.

A term $f:\ctxwk{A}{B}$ is a trivial cofibration if it has the
property that for any fibration $Q$ over $B$,
to find a section of $Q$ it suffices to find a section of the fibration
$f^\ast Q$ over $A$. In our type theoretical setting, the rôle of fibrations
is played by families, the rôle of the function type $A\to B$ is played by
$\ctxwk{A}{B}$ and our version of the pullback $f^\ast Q$ is $\subst{f}{(\ctxwk{A}{Q})}$.

\begin{defn}
Let $f$ be a term of $\ctxwk{A}{B}$ in context $\Gamma$.
\begin{enumerate}
\item For a family $Q$ over $B$ in context $\Gamma$ we define $f^\ast Q\jdeq\subst{f}{(\ctxwk{A}{Q})}$.
\item For a term $g$ of $Q$ in context $\ctxext{\Gamma}B$ we define $f^\ast g\jdeq\subst{f}{(\ctxwk{A}{g})}$.
\end{enumerate} 
\end{defn}
\begin{rmk}
The type $f^\ast(\ctxwk{B}{C})$ in context $\ctxext{\Gamma}A$ is can be viewed as the
type of functions from $A$ to $C$ which factor through $f$. Thus there should be
a function from $f^\ast(\ctxwk{B}{C})$ to $\ctxwk{A}{C}$.
\end{rmk}

\begin{rmk}
Every term $\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{B}}{f}$ allows us to infer the following:
\begin{equation*}
\inference{\jtype{\ctxext{\Gamma}B}{Q}}{\jtype{\ctxext{\Gamma}A}{f^\ast Q}}
\qquad
\inference{\jterm{\ctxext{\Gamma}B}{Q}{g}}{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{f^\ast g}}
\end{equation*}
\end{rmk}

\begin{defn}
A term $f:\ctxwk{A}{B}$ in context $\ctxext{\Gamma}A$ is said to be a trivial
cofibration if we can infer
\begin{equation*}
\inference{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{t}}{\jterm{\ctxext{\Gamma}B}{Q}{\tilde{t}}}\qquad
\inference{\jterm{\ctxext{\Gamma}A}{f^\ast Q}{t}}{\jtermeq{\ctxext{\Gamma}A}{f^\ast Q}{f^\ast \tilde{t}}{t}}
\end{equation*}
{\color{red}This statement should be reformulated so that it only involves a single judgment...
but I don't see directly how to do that.}
\end{defn}

We have the following theorem in the type theory of \cite{TheBook}, which supports
our claim that we may indeed speak of a trivial cofibration.

\begin{thm}
Suppose $f:A\to B$ is a function. Then $f$ is an equivalence if and only if
for every $Q:B\to\type$ and every $g:\prd{x:A}Q(f(x))$ there is a section
$h:\prd{y:B}Q(b)$ with the property that $h\circ f\htpy g$. 
\end{thm}

\begin{proof}
We can first take $Q$ to be the constant family $\lam{y}A$. Furthermore, we may
take $g\defeq\idfunc[A]$. Then we get a term of type
\begin{equation*}
\sm{h:B\to A}h\circ f\htpy \idfunc[A],
\end{equation*}
i.e.~we get a left inverse $h$ for $f$. To show that $h$ is also a right inverse
of $f$, let $Q$ be the family $\lam{y}\id{f(h(y))}{y}$. To find a section of
$Q$, which is the homotopy we aim for, it suffices to find a section of
$Q\circ f$. In other words, we have to show that $\id{f(h(f(x)))}{f(x)}$ for
every $x:A$. This follows from the fact that $h$ is a left inverse for $f$.

The reverse direction is immediate.
\end{proof}

Another approach would be to define $f:\ctxwk{A}{B}$ to be left invertible
if there is a term $\ctxext({\Gamma}{A})P$

\subsection{The interactions of extension, weakening and substitution}

\subsection{The basic constructors of type theory}
\subsubsection{Dependent function types}
\begin{infarray}{cc}
\inference{\jtype{\ctxext{\Gamma}A}{P}}{\jtype{\Gamma}{\mprd{A}{P}}}
& \inference{\jtypeeq{\ctxext{\Gamma}A}{P}{P'}}{\jtypeeq{\Gamma}{\mprd{A}{P}}{\mprd{A}{P'}}}\\
\inference{\jterm{\ctxext{\Gamma}A}{P}{f}}{\jterm{\Gamma}{\mprd{A}{P}}{\lambda(f)}}
& \inference{\jtermeq{\ctxext{\Gamma}A}{P}{f}{f'}}{\jtermeq{\Gamma}{\mprd{A}{P}}{\lambda(f)}{\lambda(f')}}\\
\inference{\jterm{\Gamma}{\mprd{A}{P}}{g}}{\jterm{\ctxext{\Gamma}A}{P}{\tfev(g)}}
& \inference{\jtermeq{\Gamma}{\mprd{A}{P}}{g}{g'}}{\jtermeq{\ctxext{\Gamma}A}{P}{\tfev(g)}{\tfev(g')}}\\
\inference{\jterm{\ctxext{\Gamma}A}{P}{f}}{\jtermeq{\Gamma}{A}{\tfev(\lambda(f))}{f}}
\end{infarray}

With these rules we will not get the weak $\eta$-rule when identity types are present.
So it might be better to state that $\lambda$ is a trivial cofibration

\subsection{Identity types}
We need to define the diagonal function $\jterm{\ctxext{\Gamma}A}{\ctxwk{A}{(\ctxext{A}\ctxwk{A}{A})}}{\delta}$
\begin{infarray}{c}
\inference{\jtype{\Gamma}{A}}{\jtype{\ctxext({\Gamma}{A})\ctxwk{A}{A}}{\idtypevar{A}}}\\
\inference{\jtype{\Gamma}{A}}{\jterm{\ctxext{\Gamma}A}{\delta^\ast\idtypevar{A}}{\reflsym}}\\
\inference{\jtype{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}}{P}\qquad
\jterm{\ctxext{\Gamma}A}{\pairr{\delta,\reflsym}^\ast P}{d}}
{\jterm{\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}}{P}{J}}
\end{infarray}

\section{Internal models of type theory}
\subsection{Models of type theory without basic constructors}\label{internal-model-contexts}
\begin{defn}\label{defn:premodel}
An internal model $\mfM$ of type theory consists of the following data. 
\begin{enumerate}
\item A type $\tfctx(\mfM)$ of \emph{contexts}.
\item A function $\mftypfunc{\mfM}$ assigning to each context $\Gamma$ of $\mfM$ an internal model
$\mftyp{\mfM}{\Gamma}$ of type theory.
\begin{defn}
The type $\tfctx(\mftyp{\mfM}{\Gamma})$ is denoted by $\mftyp{\mfM}{\Gamma}$. A
term of $\mftyp{\mfM}{\Gamma}$ is called a \emph{type in context $\Gamma$}. To indicate
that $A$ is a type in context $\Gamma$ we also write $\Gamma\vdash A:\mfM$. 
When $P:\mftyp{\mftyp{\mfM}{\Gamma}}{A}$ for some type $A$ in context $\Gamma$, we
also speak of \emph{a family $P$ over $A$ in context $\Gamma$.}
\end{defn}
\item A family $\terms{\blank}:\prd*{\Gamma:\ctx(\mfM)}\mftyp{\mfM}{\Gamma}\to\type$
assigning to every type $A$ in context $\Gamma$ the type $\terms{A}$ of its
terms.
\begin{defn}
When $A$ is a type in context $\Gamma$, we define $\Gamma\vdash x:A$ 
to mean $x:\terms{A}$.
\end{defn}
\item Context extension: a morphism
\begin{equation*}
\tfext^\Gamma:\mftyp{\mfM}{\Gamma}\to\mfM
\end{equation*}
of internal models for every context $\Gamma$.
\begin{defn}
When $A$ is a type in context $\Gamma$, we also denote
the context $\tfext^\Gamma_0(A)$ of $\mfM$ by $\ctxext{\Gamma}A$.
We will write
$\ctxext({\Gamma}{A})P$ for $\ctxext({\Gamma}{A})P$.
\end{defn}
\item The judgmental equalities:
\begin{align*}
\mftyp{\mftyp{\mfM}{\Gamma}}{A} & \jdeq\mftyp{\mfM}{\ctxext{\Gamma}A}\\
\mftyp{\tfext^\Gamma}{A} & \jdeq \modelfont{id}_{\mftyp{\mfM}{\ctxext{\Gamma}A}}\\
\terms[{\tfext^\Gamma}]{P} & \jdeq \idfunc[\terms{P}].
\end{align*}
\begin{rmk}
In particular, we will have the judgmental equalities:
\begin{enumerate}
\item When $A$ is a type in context $\Gamma$ we have
\begin{equation*}
\mftyp{\mftyp{\mfM}{\Gamma}}{A}\jdeq\mftyp{\mfM}{\ctxext{\Gamma}A},
\end{equation*}
ensuring that a context in the model $\mftyp{\mftyp{\mfM}{\Gamma}}{A}$ is the same thing as a
context in the model $\mftyp{\mfM}{\ctxext{\Gamma}A}$.
\item If $Q$ is a family over $P$ where $P$ is a family over $A$ in context $\Gamma$, then
\begin{equation*}
\tfext^{\protect{\mftyp{\mftyp{\mfM}{\Gamma}}{A}}}(P,Q)
\jdeq
\tfext^{\protect{\mftyp{\mfM}{\ctxext{\Gamma}A}}}(P,Q)
\end{equation*}
ensuring that twe two possible notion of context extension are the same.
\end{enumerate}
Other judgmental equalities will be required with the ingredients that follow.
We will not list them all.
\end{rmk}
\item Weakening: a morphism
\begin{equation*}
\tfwk^A:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}
\end{equation*}
of internal models for every type $A$ in context $\Gamma$. When $B$ is a type
in context $\Gamma$, we denote $\tfwk^A(B)$ by $\ctxwk{A}{B}$. 
\item Substitution: a morphism
\begin{equation*}
\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}A}\to\mftyp{\mfM}{\Gamma}
\end{equation*}
of internal models for any $x:A$ in context $\Gamma$. When $P$ is a family over $A$ in context
$\Gamma$, we denote $\tfsubst^x(P)$ also by $\subst{x}{P}$. 
\item A judgmental equality
\begin{equation*}
\tfsubst^x\circ\tfwk^A\jdeq\modelfont{id}_{\mftyp{\mfM}{\Gamma}}
\end{equation*}
for any $x:A$ and $B$ in context $\Gamma$.
\item A context $\unit^\mfM:\tfctx(\mfM)$ and a term of type $\isequiv(\ctxext{\unit^\mfM}\blank)$. We will denote
this equivalence by $e_\unit$. The context $\unit^\mfM$ is also
called the \emph{empty context}.
\begin{defn}
For any context $\Gamma$, type $\terms{\Gamma}$ is defined to mean
$\terms{e_\unit^{-1}(\Gamma)}$. 
\end{defn}
\item A section $\unit^{\blank}:\prd{\Gamma:\ctx(\mfM)}\mftyp{\mfM}{\Gamma}$ assigning
a type $\unit^\Gamma$ in context $\Gamma$ to every context $\Gamma$ and
an identification $\alpha_\unit(\Gamma):\id{\ctxext{\Gamma}\unit^\Gamma}{\Gamma}$
for every context $\Gamma$. We also require that there is an identification
$\id{\trans{\alpha_\unit(\Gamma)}{\ctxwk{\unit^\Gamma}{A}}}{A}$ for every
type $A$ in context $\Gamma$.
\item For any type $A$ in context $\Gamma$, a term $\idfunc[A]:\terms{\ctxwk{A}{A}}$
\end{enumerate}
\begin{flushright}
\textsl{End of \autoref{defn:premodel}.}
\end{flushright}
\end{defn}

\begin{defn}
In an internal model $\mfM$ we define
\begin{equation*}
\ctxhom{\Delta}{\Gamma}\defeq \terms{\ctxwk{e_\unit^{-1}(\Delta)}{e_\unit^{-1}(\Gamma)}}
\end{equation*}
\end{defn}

\subsection{Morphisms of internal models}
\begin{defn}\label{defn:premodel-morphism}
A morphism $f:\mfM\to \mfN$ of internal models consists of
\begin{enumerate}
\item a function $\ctx(f):\ctx(\mfM)\to\ctx(\mfN)$. The function $\ctx(f)$ is also
denoted by $f_0$.
\item a morphism $\mftyp{f}{\Gamma}:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfN}{f_0(\Gamma)}$ of internal models for every
$\Gamma:\ctx(M)$.
\item the judgmental equality
\begin{equation*}
\mftyp{\mftyp{f}{\Gamma}}{A}\jdeq\mftyp{f}{\ctxext{\Gamma}A}
\end{equation*}
\item a function $\terms[f]{A}:\terms[\mfM]{A}\to\terms[\mfN]{\mftyp{f}{\Gamma}_0(A)}$ for
every type $A$ in context $\Gamma$. 
\item preservation of context extension: 
\begin{align*}
\alpha^f_0 & :\id{f\circ\tfext^\Gamma}{\tfext^{f_0(\Gamma)}\circ\mftyp{f}{\Gamma}}\\
\alpha^f_1 & :\id{\mftyp{f}{\Gamma}_0\circ\tfext^A}{\tfext^{\mftyp{f}{\Gamma}_0(A)}\circ\mftyp{f}{\ctxext{\Gamma}A}}.
\end{align*}
\begin{comment}
the judgmental equality
\begin{equation*}
f_0(\ctxext{\Gamma}A)\jdeq\ctxext{f_0(\Gamma)}\mftyp{f}{\Gamma}_0(A)
\end{equation*}
for every type $A$ in context $\Gamma$.
\end{comment}
\item preservation of weakening: 
\begin{align*}
\beta^f_0 & :\id{\mftyp{f}{\Gamma}\circ\tfwk^\Gamma}{\tfwk^{f_0(\Gamma)}\circ f}\\
\beta^f_1 & :\id{\mftyp{f}{\ctxext{\Gamma}A}\circ\tfwk^A}{\tfwk^\protect{\mftyp{f}{\Gamma}_0(A)}\circ\mftyp{f}{\Gamma}}.
\end{align*}
\begin{comment}
This gives the following three judgmental equalities:
\begin{enumerate}
\item the judgmental equality
\begin{equation*}
\mftyp{f}{\ctxext{\Gamma}A}_0(\ctxwk{A}{B})\jdeq\ctxwk{\mftyp{f}{\Gamma}_0(A)}{\mftyp{f}{\Gamma}_0(B)}
\end{equation*}
for every two types $A$ and $B$ in context $\Gamma$.
\item the judgmental equality
\begin{equation*}
\mftyp{f}{\ctxext({\Gamma}{A})\ctxwk{A}{B}}\circ\tfwk^A\jdeq\tfwk^{\protect{\mftyp{f}{\Gamma}_0(A)}}\circ\mftyp{f}{\ctxext{\Gamma}B}
\end{equation*}
\item the judgmental equality
\begin{equation*}
\terms[f]{\ctxwk{A}{B}}(\ctxwk{A}{y})\jdeq\ctxwk{\mftyp{f}\Gamma_0(A)}{(\terms[f]{B}(y))}
\end{equation*}
\end{enumerate}
\end{comment}
\item preservation of substitution: 
\begin{equation*}
\gamma^f:\id{\mftyp{f}\Gamma_0\circ\tfsubst^x}{\tfsubst^\protect{\terms[f]{A}(x)}\circ\mftyp{f}{\ctxext{\Gamma}A}}.
\end{equation*}
\begin{comment}
\begin{enumerate}
\item the judgmental equality
\begin{equation*}
\mftyp{f}\Gamma_0(\subst{x}{P})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext{\Gamma}A}_0(P)}
\end{equation*}
for every family $P$ over $A$ in context $\Gamma$ and every term $x:A$.
\item the judgmental equality
\begin{equation*}
\mftyp{f}\Gamma_0(\subst{x}{Q})\jdeq\subst{\terms[f]{A}(x)}{\mftyp{f}{\ctxext({\Gamma}{A})P}_0(Q)}
\end{equation*}
for every family $Q$ over $P$ over $A$ in context $\Gamma$ and every term $x:A$.
\end{enumerate}
\end{comment}
\end{enumerate}
\begin{flushright}
\textsl{End of \autoref{defn:premodel-morphism}.}
\end{flushright}
\end{defn}

\begin{defn}
Suppose that $f:\mfM\to\mfN$ and $g:\mfN\to\mfN'$ are morphisms of internal models.
We define the composition $g\circ f:\mfM\to\mfN'$ to be the morphism given by
\begin{enumerate}
\item $(g\circ f)_0\defeq g_0\circ f_0$
\item $\mftypfunc{g\circ f}(\Gamma)\defeq\mftypfunc{g}(f_0(\Gamma))\circ\mftypfunc{f}(\Gamma)$
\item $\terms[g\circ f]{A}\defeq\terms[g]{\mftyp{f}{\Gamma}_0(A)}\circ\terms[f]{A}$.
\item a definition of $\alpha^{g\circ f}_0$ and $\alpha^{g\circ f}_1$.
\item a definition of $\beta^{g\circ f}_0$ and $\beta^{f\circ f}_1$.
\item a definition of $\gamma^{g\circ f}$. 
\end{enumerate}
\end{defn}

\begin{rmk}
The requirement that a morphism of models acts on terms is reminiscent of the requirement
that a functor acts on morphisms. In this fasion, the requirement that a morphism of
models preserves substitution is the counterpart of a functor preserving composition.
\end{rmk}
Context extension, weakening and substitution are required to be such morphims
of models. Thus they must be extended for this purpose.

\begin{description}
\item[Context extension] We will extend context extension to a morphism 
$\tfext(\Gamma):\mftyp{\mfM}{\Gamma}\to M$. Thus, we already have
$\tfext(\Gamma)_0\defeq\ctxext{\Gamma}\blank$. We also require 
\begin{align*}
\mftyp{\tfext(\Gamma)}{A} & : \mftyp{\mftyp{\mfM}{\Gamma}}{A}\to \mftyp{\mfM}{\ctxext{\Gamma}A}\\
\terms[\tfext(\Gamma)]{P} & : \terms[\mftyp{\mfM}{\Gamma}]{P}\to\terms[\mfM]{\mftyp{\tfext(\Gamma)}{A}_0(P)}
\end{align*}
For both of these we take the identity function.
We require furthermore that context extension preserves context extension,
weakening and substitution:
\begin{enumerate}
\item Extension preserves extension: for every family $P$ over $A$ in context
$\Gamma$ an identification $\id{\ctxext({\Gamma}{A})P}{\ctxext{\Gamma}({A}P)}$.
\item Extension preserves weakening:
\item Extension preserves substitution:
Because context extension is the identity on the levels of types and terms,
the rules are easier.
\end{enumerate}
\item[Weakening] is a morphism $\tfwk:\mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}$, so we must have
\begin{align*}
\tfwk^A_0 & : \mftyp{\mfM}{\Gamma}\to\mftyp{\mfM}{\ctxext{\Gamma}A}
\intertext{which is denoted by $\ctxwk{A}{\blank}$,}
\mftypfunc{\tfwk^A} & : \mftyp{\mfM}{\ctxext{\Gamma}B}\to\mftyp{\mfM}{\ctxext({\Gamma}{A})\ctxwk{A}{B}}\\
\terms[\tfwk^A]{B} & : \terms[\mftyp{\mfM}{\Gamma}]{B}\to\terms[\mftyp{\mfM}{\ctxext{\Gamma}A}]{\ctxwk{A}{B}}
\end{align*}
for a term $y:B$, the term $\terms[{\tfwk^A}]{B}(y)$ is the constant map from
$A$ to $B$, assigning $y$ to every term of $A$.
\item[Substitution] is a morphism $\tfsubst^x:\mftyp{\mfM}{\ctxext{\Gamma}A}
\to\mftyp{\mfM}{\Gamma}$ for every $x:A$ in context $\Gamma$, so we must have
\begin{align*}
\tfsubst^x_0 & : \mftyp{\mfM}{\ctxext{\Gamma}A}\to\mftyp{\mfM}{\Gamma}
\intertext{which is denoted by $\subst{x}{\blank}$,}
\mftyp{\tfsubst^x}{P} & : \mftyp{\mfM}{\ctxext({\Gamma}{A})P}\to\mftyp{\mfM}{\ctxext{\Gamma}\subst{x}{P}}\\
\terms[\tfsubst^x]{P} & : \terms{P}\to\terms{\subst{x}{P}}
\end{align*}
The function $\terms[\tfsubst^x]{P}$ is usually denoted by $\tfev(\blank,x)$. 
\end{description}

\subsection{The basic type constructors in internal models}

\begin{defn}
\begin{enumerate}
\item A type $\mprd{A}{P}$ for every type $A$ in context $\Gamma$ and every type
$P$ in context $\ctxext{\Gamma}A$.
\begin{defn}
Suppose $A$ and $B$ are types in context $\Gamma$. We define
\begin{align*}
A\to B & \defeq\mprd{A}{\ctxwk{A}{B}}.
\intertext{For any two contexts $\Delta$ and $\Gamma$, we define}
\Delta\to\Gamma & \defeq e_\unit^{-1}(\Delta)\to e_\unit^{-1}(\Gamma)
\end{align*}
and furthermore we define $\ctxhom{\Delta}{\Gamma}\defeq\terms{\Delta\to\Gamma}$.
\end{defn}
\item An equivalence $\lambda:\eqv{\terms{P}}{\terms{\mprd{A}{P}}}$ for every
family $P$ over $A$ in context $\Gamma$. When $\ctxext{\Gamma}A\vdash 
u:P$, we call $\lambda(u)$ the \emph{$\lambda$-abstraction of $u$.}
\begin{rmk}
Note that we get an equivalence $\eqv{\terms{\unit^\Gamma\to A}}{\terms{A}}$ 
for every type $A$ in context $\Gamma$.
\end{rmk}
\begin{rmk}
Thus we see that $\eqv{\terms{A\to B}}{\terms{\ctxwk{A}{B}}}$
by $\lambda$-abstraction. 
\end{rmk}
%\item A function $\tfev:\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$.
\item A term $\pi^A:\ctxext{\Gamma}A\to \Gamma$ for every type $A$ in context
$\Gamma$.
\item A term $\iota^x:\ctxext{\Gamma}\subst{x}{P}\to\ctxext({\Gamma}{A})P$
for every family $P$ of types over $A$ in context $\Gamma$ and every
term $x:A$.
\item A type $A[f]$ in context $\Delta$ for every $f:\ctxhom{\Delta}{\Gamma}$ 
and every type $A$ in context $\Gamma$. {\color{blue}Could this be defined
in terms of the substitution $\subst{x}{P}$ we already have? It seems so. If this is
indeed the case we need either some identifications or we could just omit
this part of the definition.}
\item an identification $\id{B[\pi^A]}{\ctxwk{A}{B}}$ for any two types $A$
and $B$ in context $\Gamma$.
\item A type $\msm{A}{P}$ in context $\Gamma$ for every family $P$ over $A$
in context $\Gamma$, with an equivalence $\pairr{\blank,\blank}:\eqv{\sm{x:\terms{A}}\terms{\subst{x}{P}}}
{\terms{\msm{A}{P}}}$.
\item A family $\idtypevar{A}$ over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$ for
every type $A$ in context $\Gamma$ and a term of type.
\begin{rmk}
If $A$ and $B$ are types in context $\Gamma$, we may denote the type 
$\msm{A}{\ctxwk{A}{B}}$ by $A\times B$. By the end of the current definition
there will be an identification $\id{\ctxext({\Gamma}{A})P}
{\ctxext{\Gamma}\msm{A}{P}}$.

There is a term $\delta:\ctxhom{\ctxext{\Gamma}A}{\ctxext({\Gamma}{A})\ctxwk{A}{A}}$
defined by...
\end{rmk}
\item For any family $Q$ over $\idtypevar{A}$ in context $\ctxext({\Gamma}{A})\ctxwk{A}{A}$ an equivalence
$J:\eqv{\terms{Q}}{}$
\end{enumerate}
\end{defn}

\begingroup
\color{blue}
\subsubsection*{More desiderata}
Some of which hopefully follow from more elegant or general rules, but
this list is to keep them in mind:
\begin{enumerate}
\item $\id{e_\unit^{-1}(\Gamma.A)}{\msm{e_\unit^{-1}(\Gamma)}{\trans{(\eta_\Gamma)}{A}}}$, where
$\eta_\Gamma:\id{\Gamma}{e_\unit(e_\unit^{-1}(\Gamma))}$ is the unit of the
equivalence $e_\unit$.  
\item $\id{(\msm{A}{P}\to B)}{(P\to \ctxwk{A}{B})}$, expressing that $\Sigma$ is
left adjoint to weakening.
\item $\id{(\ctxwk{A}{B}\to P)}{(B\to\mprd{A}{P})}$ expressing that $\Pi$ is
right adjoint to weakening.
\item a term $\Gamma\vdash\idfunc[A]:A\to A$ for every type $A$ in context $\Gamma$.
\item a context $\UU$ for the universe.
\item an identification $\id{\pi^A\circ \pi^P\circ\iota^x}{\pi^{\subst{x}{P}}}$ for every
family $P$ of types over $A$ in context $\Gamma$ and every term $x:A$. In other
words, the diagram
\begin{equation*}
\begin{tikzcd}
\ctxext({\Gamma}{A})P \ar{r}{\pi^P} & \ctxext{\Gamma}A \ar{d}{\pi^A}\\
\ctxext{\Gamma}\subst{x}{P} \ar{u}{\iota^x} \ar{r}[swap]{\pi^{\subst{x}{P}}} & \Gamma
\end{tikzcd}
\end{equation*}
should commute.
\begin{comment}
\item For $\Gamma:\ctx$, $\Gamma\vdash A:\type$, $\ctxext{\Gamma}A\vdash P
:\type$ and $\ctxext({\Gamma}{A})P\vdash Q:\type$, $x:A$ and $u:P$
we desire that $\typefont{nat{-}ev{-}subst}(Q,u,x):\id{Q[\iota^x][\tfev(u,x)]}{Q[u][x]}$.
\item A function
\begin{equation*}
\tfcomp:\terms{\mprd{P}{Q}}\to\prd{x:\terms{A}}\terms{\mprd{\subst{x}{P}}{Q[\iota^x]}}
\end{equation*}
which could possibly replace $\tfev$ because it could be that $\tfev$ is
definable in terms of $\tfcomp$. It should satisfy
\begin{equation*}
\id{\trans{\typefont{nat{-}ev{-}subst}(Q,u,x)}{\tfev(\tfcomp(f,x),\tfev(u,x))}}{\tfev(\lambda(\tfev(f,u)),x)},
\end{equation*}
which is a naturality condition. There should be an equivalence between
$\terms{\mprd{A}{P}}$ and a type consisting of those terms of $\prd{x:\terms{A}}\terms{\subst{x}{P}}$ which satisfy
a certain naturality condition. One could hope that by formulating this dependently
enough, it is possible to write down such a naturality in one line.
\end{comment}
\end{enumerate}
\endgroup

\begin{comment}
\begingroup
\color{darkgreen}
\subsubsection*{The problem with evaluation.}
...or why I'm doing such strange things. It seems that evaluation is of a
different nature. At first sight, one might expect that evaluation compares
to $\lambda$-abstraction the same way that first and second projection compare
to the pairing function $\pairr{\blank,\blank}$. However, this is not the case
because evaluation cannot be seen as an inverse to $\lambda$-abstraction. The
$\lambda$-abstraction is of type $\terms{P}\to\terms{\mprd{A}{P}}$ whereas
evaluation is of type $\terms{\mprd{A}{P}}\to\prd{x:\terms{A}}\terms{\subst{x}{P}}$. There
are a few things to note about $\tfev$:
\begin{enumerate}
\item There's no function back so in particular $\tfev$ is not an equivalence.
My guess is that there should be a naturality condition playing a role.
\item Evaluation tells how functions interact with substitution of terms.
\end{enumerate}
So it seems that evaluation is like a variable binder in the same way
that quantifiers are binding variables. If I remember right, Voevodsky mentioned
something like this at the Formal Topology conference in Ljubljana, though I
forgot what exactly he said.
\endgroup
\end{comment}

\begin{comment}
\begingroup
\color{gray}
\subsubsection*{Old stuff below:}
\begin{enumerate}
\item a context morphism $\catid{\Gamma}:\ctxhom{\Gamma}{\Gamma}$ for every
context $\Gamma$.
\item a context morphism $f\circ g:\ctxhom{E}{\Gamma}$ for every two context
morphisms $f:\ctxhom{\Delta}{\Gamma}$ and $g:\ctxhom{E}{\Delta}$. 
\item paths $\id{f\circ\catid{\Delta}}{f}$ and $\id{\catid{\Gamma}\circ f}{f}$
for every $f:\ctxhom{\Delta}{\Gamma}$.
\item a path $\id{f\circ (g\circ h)}{(f\circ g)\circ h}$ for every
$f:\ctxhom{\Delta}{\Gamma}$, $g:\ctxhom{E}{\Delta}$ and $h:\ctxhom{Z}{E}$. 
\item a context morphism $\Gamma_!:\ctxhom{\Gamma}{\unit^\mfM}$ 
for every context $\Gamma$.
\item a path $\id{(\unit^\mfM)_!}{\catid{\unit^\mfM}}$.
\item a path $\id{\Gamma_!\circ f}{\Delta_!}$ for every $f:\ctxhom{\Delta}{\Gamma}$. 
\item a path $\id{A[\catid{\Gamma}]}{A}$ for every object $A$ in context $\Gamma$.
\item a path $\id{A[f][g]}{A[f\circ g]}$ for every object $A$ in context $\Gamma$
and every $f:\ctxhom{\Delta}{\Gamma}$ and $g:\ctxhom{E}{\Delta}$. 
\end{enumerate}
\endgroup
\end{comment}

\begin{desiderata}
We should have a theorem stating that our internal model indeed interprets
the rules of type theory.
\end{desiderata}

\begin{comment}
\section{Continuity of functions}
We wish to find a condition under which we have
\begin{equation*}
\eqv{\terms{\mprd{A}{P}}}{\sm{f:\prd{x:\terms{A}}\terms{\subst{x}{P}}}\text{``$f$ is continuous''}}.
\end{equation*}
for all families $P$ over $A$. 
Needless to say, we first need to formulate what we mean by ``$f$ is continuous''.
Thus, we need to find the family $\pairr{p,u,v}\mapsto \id{\trans{p}{u}}{v}$,
which also requires us to define transportation.

Let $P$ be a family over $A$ in context $\Gamma$. Then we can weaken
$\ctxwk{A}{P}$, which is a family over $\ctxwk{A}{A}$ in context $\ctxext{\Gamma}A$.
We also have the family $\ctxwk{\ctxwk{A}{A}}{P}$ over $\ctxwk{A}{A}$ in context
$\ctx{\Gamma}{A}$ and thus we can form the family
\begin{equation*}
\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}.
\end{equation*}
This family plays the role of the family $x,y:A\vdash P(x)\to P(y):\type$.
{\color{red}Explain this by performing the relevant substitution.}
Transportation is a term of the family
\begin{equation*}
\idtypevar{A}\to\ctxwk{\ctxwk{A}{A}}{P}\to\ctxwk{A}{P}
\end{equation*}
obtained by identity elimination, as usual. Thus we obtain a term
\begin{equation*}
\mathcal{T}rans_{x,y}:\terms{\id{x}{y}\to\subst{x}{P}\to\subst{y}{P}}.
\end{equation*}
for every $x,y:A$.

Now let $x,y:A$. Then we have the type $\subst{y}{P}$
in context $\Gamma$ and the family $\ctxwk{\subst{y}{P}}{\subst{x}{P}}$
over $\subst{y}{P}$ in context $\Gamma$. We could weaken $\subst{y}{P}$ subsequently by $A$, $\ctxwk{A}{A}$
and $\idtypevar{A}$ to obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\ctxwk{A}{A}}{\ctxwk{A}{\subst{y}{P}}}}$ over $\idtypevar{A}$ in context $\ctxext({\Gamma}{A})\ctxwk{A}{A}$. We will
denote this family simply by $\ctxwk{\idtypevar{A}}{\subst{y}{P}}$. Likewise,
we obtain the family $\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{x}{P}}}$ over
$\ctxwk{\idtypevar{A}}{\subst{y}{P}}$ in context $\ctxext(({\Gamma}{A}){\ctxwk{A}{A}})\idtypevar{A}$ and we obtain the family $\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}$
over $\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{y}{P}}}$. Now we
have the family $\subst{\ctxwk{A}{\mathcal{T}rans_{x,y}}}{\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}}$ over
$\ctxwk{\idtypevar{A}}{\ctxwk{\subst{y}{P}}{\subst{x}{P}}}$, which is
the desired family to state the continuity condition:

\begin{defn}
A function $f:\prd{x:\terms{A}}\terms{\subst{x}{P}}$ is said to be \emph{continuous}
if there is a term of type
\begin{equation*}
\prd{x,y:\terms{A}}\terms{\subst{\ctxwk{A}{\mathcal{T}rans_{x,y}}}{\ctxwk{\idtypevar{A}}{\idtypevar{\subst{y}{P}}}}}.
\end{equation*}
\end{defn}

\begin{lem}
If all functions are ``continuous mappings of terms'' then types without terms
are empty, i.e.\ we have
\begin{equation*}
(\terms{A}\to\emptyt)\to\terms{A\to\emptyt^\mfGraph}
\end{equation*}
\end{lem}

\begin{proof}
Any function from a type without terms is vacuously continuous.
\end{proof}

We will see that in the graph model functions are not just continuous mappings
of terms, since there is a graph $\tilde{\emptyt}$ which has no terms but which
differs from the empty graph nevertheless.

\begin{desiderata}
The following may relate to functions being continuous mappings of terms (it
seems to be a very strong condition):
\begin{enumerate}
\item We have the adjunctions $\tfcolim\dashv\Delta\dashv\tflim$ in which the
two monads are idempotent (we don't even have the adjunctions for $\mfGraph$).
\item $\eqv{\terms{\Omega}}{\ctx}$
\end{enumerate}
We also need to check whether all functions in the reflexive graphs have this
property (if not, ditch this section).
\end{desiderata}
\end{comment}