### 2013-12-09 article-models-main.tex

\documentclass{article}

\input{preamble-articles}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Internal models of type theory}
\author{Univalent Foundations Subprogram}
\date{\today}

\begin{document}

\maketitle

\tableofcontents

\input{article-models-intro}

\input{article-models-eg}

\input{article-models-tt}

\input{article-models-idtypes}

\input{article-models-constructors}

\input{article-models-defn}

\bibliographystyle{plain}
%\phantomsection\addcontentsline{toc}{section}{References}
\bibliography{refs}

\end{document}