### 2014-01-08 article-models-main.tex

\documentclass{article}

\input{preamble-articles}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Internal models of type theory and internal higher categories for the
Univalent Foundations}
\author{Egbert Rijke}
\date{\today}

\begin{document}

\maketitle

\tableofcontents

\begin{abstract}
A project is proposed where we investigate notions of internal models of type
theory and internal higher categories. We take the point of view that an
internal model is an internal higher category with extra structure interpreting
the type constructors. We propose to (1) explore possible definitions of both
notions, (2) find a zoo of examples and (3) find extensions of the underlying
theory and models thereof to include higher inductive types.
\end{abstract}

\input{article-models-intro}

%\input{article-models-eg}

\appendix
\input{article-models-tt}

\input{article-models-idtypes}

%\input{article-models-constructors}

\input{article-models-defn}

\bibliographystyle{plain}
%\phantomsection\addcontentsline{toc}{section}{References}
\bibliography{refs}

\end{document}