### 2014-01-08 article-models-stage2.tex

\section{Stage two: extending the basic theory}\label{stage2}
The first aim of this project is thus to give this new description of type
theory with these properties. The second aim is to define a general notion of
an internal model and provide several examples thereof. In the definition of
internal models we envision that all the operations of type theory become
morphisms of type theory: they are compatible with each and every morphism
of type theory. The compatibility of extension, weakening and substitution is
already part of the basic type theory. The compatibility of identity types
with weakening for instance, is a form of function extensionality. The 
compatibility of dependent function types with identity types is the usual
funchtion extensionality principle. The
compatibility of a universe operation with identity types should be univalence.
The descent theorem implies that the colimit operation is also compatible with
all the type theoretic operations (sigma and id are checked).

We also wish to extend the descent property to a more general class of homotopy
colimits. The general recipe has already been implicit in the first descent
property where we have used diagrams over graphs. We have the following in mind:
\begin{enumerate}
\item Give an internal description of a diagram. Probably this is the same thing
      as an internal model of the basic type theory with or without identity types.
\item Define the notion of homotopy colimit thereof.
\item Define the notion if equifibered diagrams.
\item Formulate and prove descent: the map defined by substitution
 from families over the colimit to the equifibered diagrams is an equivalence.
\end{enumerate}

Using the descent property we can view $\tfcolim$ as an operation of type theory
It should preserve all the basic operations, identity types, dependent function
types, itself, etc.

Another thing we must keep in mind is that the compatibility rules in the basic
type theory are strict. We should look to the possibility to weaken them.