### 2014-05-17 article-models-ttuniverses.tex

\section{Universes}\label{sec:universes}

\subsection{Pre-universes}
We would like to investigate the universe operator, which assigns to a family
$A$ over $\Gamma$ and a family $P$ over $\ctxext{\Gamma}{A}$ a universe
$\UU(A,P)$, which is a family over $\Gamma$, and a decoding $\tilde{\UU}(A,P)$,
which is a family over $\ctxext{\Gamma}{\UU(A,P)}$. However, we find it useful
to first look at a pre-universe operator which extends the pair $(A,P)$ to a
quadruple $(V_0(A,P),\tilde{V}_0(A,P),V_1(A,P),\tilde{V}_1(A,P))$ consisting of
a family $V_0(A,P)$ in context $\Gamma$, a family $\tilde{V}_0(A,P)$ over
$\ctxext{\Gamma}{V_0(A,P)}$, a family $V_1(A,P)$ over 
$\ctxext{\Gamma}{V_0(A,P)}$ and a family $\tilde{V}_1(A,P)$ over.... and only
later blur the distinction between $\subst{x}{V_1(A,P)}$ and family morphisms
from $\subst{x}{\tilde{V}_0(A,P)}$ to $V_0(A,P)$ for terms $x:V_0(A,P)$. The
pair $(V_0(A,P),V_1(A,P))$ can be seen as a version of $(A,P)$ which is
expanded to incorporate extension, the empty context, weakening, substitution
and identity terms.

Our initial setup for pre-universes consists just of the two inverence rules
calling for $V_0$ and $V_1$:
\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{\Gamma}{V_0(A,P)}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfam{{\Gamma}{V_0(A,P)}}{V_1(A,P)}
    }
\end{align*}

\subsection{Universe operators}
\label{universes}
Now we investigate universes. When $A$ is a family over context $\Gamma$ and
$P$ is a family over $\ctxext{\Gamma}{A}$, the universe at $(A,P)$ consists,
among other things, of a family $\UU_0(A,P)$ over context $\Gamma$ and a family
$\UU_1(A,P)$

The type operator $\UU$ has as input a pair $(A,P)$ consisting of a family
$A$ in context $\Gamma$ and a family $P$ in context $\ctxext{\Gamma}{A}$. Its
output is a pair $(U(A,P),\tilde{U}(A,P))$ consisting of a family
$U(A,P)$ in context $\Gamma$ and a family $\tilde{U}(A,P)$ over it for which
there are terms representing the empty context, the operations of extension,
weakening and substitution and the identity terms. Moreover there is a morphism
$i$ from $A$ to $U(A,P)$ with the property that 
$\jcomp{}{i}{\tilde{U}(A,P)}\jdeq P$.