### 2014-06-20 article-models-ttdp.tex
The first version of the set of rules for the dependent function constructor is finished

\section{The dependent function constructor}
As we have done with most everything so far, we will take the point of view that
the dependent function constructor is an operation with an action on contexts,
on families and on terms. When $A$ is a family of contexts over $\Gamma$,
$\mprd{A}{\blank}$ takes things in context $\ctxext{\Gamma}{A}$ to the context
$\Gamma$. It's action on terms restricted to the empty context is usually 
denoted by $\lambda$, so we shall denote the entire action by $\lambda$. As
usual, $\lambda$ is reversed by evaluation.

The dependent function type constructor not only acts on families $P$ of
contexts over $\ctxext{\Gamma}{A}$, but it also sends families $Q$ of contexts
over $\ctxext{{\Gamma}{A}}{P}$ to families $\mprd[\famsym]{A}{Q}$ over $\ctxext{\Gamma}
{\mprd{A}{P}}$. Moreover, we will also have a version of $\lambda$-abstraction
for terms of such families $Q$.

\begin{align*}
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfam{\Gamma}{\mprd{A}{P}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{\Gamma}{A}}{P}{P'}
        }
      { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd{A'}{P'}}
        }
  \\
& \inference
    { \jfam{{{\Gamma}{A}}{P}}{Q}
      }
    { \jfam{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jfameq{{{\Gamma}{A}}{P}}{Q}{Q'}
        }
      { \jfameq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\mprd[\famsym]{A'}{Q'}}
        }
  \\
& \inference
    { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
      }
    { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\slam{A}{Q}{g}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jtermeq{{{\Gamma}{A}}{P}}{Q}{g}{g'}
        }
      { \jtermeq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{\slam{A}{Q}{g}}{\slam{A'}{Q}{g'}}
        }
  \\
& \inference
    { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{g}
      }
    { \jterm{{{\Gamma}{A}}{P}}{Q}{\sev{A}{Q}{g}}
      }
& & \inference
      { \jfameq{\Gamma}{A}{A'}
        \jtermeq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{Q}}{g}{g'}
        }
      { \jtermeq{{{\Gamma}{A}}{P}}{Q}{\sev{A}{Q}{g}}{\sev{A'}{Q}{g'}}
        }
\end{align*}

\begin{defn}
Let $A$ and $B$ be families of contexts over $\Gamma$. Then we define
\begin{equation*}
\jfamdefn{\Gamma}{\jfun{A}{B}}{\sprd{A}{\ctxwk{A}{B}}}.
\end{equation*}
\end{defn}

\subsection{The compatibility rules for the dependent product constructor}
In this subsection we lay out the compatibility rules which we will require
for the dependent product constructor.

\subsubsection{Dependent products are compatible with the empty context}
The empty context can appear in the domain and in the codomain of the dependent
function type constructor. We have the following inference rules explaining
what happens when the empty context appears in the domain:
\begin{align}
& \inference
    { \jfam{\Gamma}{P}
      }
    { \jfameq{\Gamma}{\mprd{\emptyf}{P}}{P}
      }
  \\
& \inference
    { \jfam{{\Gamma}{P}}{Q}
      }
    { \jfameq{{\Gamma}{P}}{\mprd[\famsym]{\emptyf}{Q}}{Q}
      }
  \\
& \inference
    { \jterm{{\Gamma}{P}}{Q}{g}
      }
    { \jtermeq{{\Gamma}{P}}{Q}{\slam{\emptyf}{Q}{g}}{g}
      }
  \\
& \inference
    { \jterm{{\Gamma}{P}}{Q}{g}
      }
    { \jtermeq{{\Gamma}{P}}{Q}{\sev{\emptyf}{Q}{g}}{g}
      }
\end{align}
We have the following infernece rules explaining what happens when the empty
context appears in the codomain:
\begin{align}
& \inference
    { \jfam{\Gamma}{A}
      }
    { \jfameq{\Gamma}{\mprd{A}{\emptyf}}{\emptyf}
      }
  \\
& \inference
    { \jfam{{\Gamma}{A}}{P}
      }
    { \jfameq{{\Gamma}{\mprd{A}{P}}}{\mprd[\famsym]{A}{\emptyf}}{\emptyf}
      }
\end{align}
Now we see that we can compare $\mprd{A}{P}$ with $\mprd[\famsym]{A}{P}$ by
seeing $P$ as a family over $\ctxext{{\Gamma}{A}}{\emptyf}$. We will require
the following inference rule to be valid:
\begin{equation}
\inference
  { \jfam{{\Gamma}{A}}{P}
    }
  { \jfameq{\Gamma}{\mprd{A}{P}}{\mprd[\famsym]{A}{P}}
    }
\end{equation}
This rule will allow us to omit the annotation indicating the action on families,
which we will do from now on. We also note that because the empty family over
$\ctxext{\Gamma}{A}$ is mapped to the empty family over $\Gamma$ by the
dependent function constructor, we obtain the important special cases of
lambda abstraction and evaluation that
\begin{align*}
& \inference
  { \jterm{{\Gamma}{A}}{P}{f}
    }
  { \jterm{\Gamma}{\mprd{A}{P}}{\slam{A}{P}{f}}
    }
  \\
& \inference
  { \jterm{\Gamma}{\mprd{A}{P}}{f}
    }
  { \jterm{{\Gamma}{A}}{P}{\sev{A}{P}{f}}
    }
\end{align*}
Thus, we retrieve what is originally meant by lambda abstraction.

\subsubsection{Dependent products are compatible with extension}
The following rules describe what happens when a dependent function constructor
is taken over an extension

\begin{align*}
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq
      {\Gamma}
      {\mprd{\ctxext{A}{P}}{Q}}
      {\mprd{A}{\mprd{P}{Q}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}
      {\mprd{\ctxext{A}{P}}{R}}
      {\mprd{A}{\mprd{P}{R}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jterm{{{{\Gamma}{A}}{P}}{Q}}{R}{h}
    }
  { \jtermeq
      {{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}
      {\mprd{\ctxext{A}{P}}{R}}
      {\slam{\ctxext{A}{P}}{R}{h}}
      {\slam{A}{\mprd{P}{R}}{\slam{P}{R}{h}}}
    }
  \\
& \inference
  { \jfam{{\Gamma}{A}}{P}
    \jterm{{\Gamma}{\mprd{\ctxext{A}{P}}{Q}}}{\mprd{\ctxext{A}{P}}{R}}{h}
    }
  { \jtermeq
      {{{\Gamma}{{A}{P}}}{Q}}
      {R}
      {\sev{\ctxext{A}{P}}{R}{h}}
      {\sev{P}{R}{\sev{A}{\mprd{P}{R}}{h}}}
    }
\end{align*}

When the dependent product constructor is applied to an extension, we get the
following:

\begin{align*}
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jfameq
      {\Gamma}
      {\mprd{A}{\ctxext{P}{Q}}}
      {\ctxext{\mprd{A}{P}}{\mprd{A}{Q}}}
    }
  \\
& \inference
  { \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\mprd{A}{P}}}
      {\mprd{A}{\ctxext{Q}{R}}}
      {\ctxext{\mprd{A}{Q}}{\mprd{A}{R}}}
    }
\end{align*}
We show in \autoref{lem:slam-ext} that the remaining two properties
involving lambda abstraction and evaluation are derivable.

\subsubsection{Weakening is compatible with the dependent product constructor}

\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\mprd{Q}{R}}}
      {\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{{\Gamma}{B}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\mprd{Q}{R}}}}
      {\ctxwk{A}{\mprd{Q}{S}}}
      {\mprd{\ctxwk{A}{Q}}{\ctxwk{A}{S}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{{{\Gamma}{B}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\mprd{Q}{R}}}}
      {\ctxwk{A}{\mprd{Q}{S}}}
      {\ctxwk{A}{\slam{Q}{S}{k}}}
      {\slam{\ctxwk{A}{Q}}{\ctxwk{A}{S}}{\ctxwk{A}{k}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{{\Gamma}{B}}{\mprd{Q}{R}}}{\mprd{Q}{S}}{k}
    }
  { \jtermeq
      {{{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}}{\ctxwk{A}{R}}}
      {\ctxwk{A}{S}}
      {\ctxwk{A}{\sev{Q}{S}{k}}}
      {\sev{\ctxwk{A}{Q}}{\ctxwk{A}{S}}{\ctxwk{A}{k}}}
    }
\end{align*}

\subsubsection{Dependent products are taken fiberwise}
The rules explaining that substitution is compatible with the dependent product
constructor assert that dependent products are taken fiberwise (as is usual).

\begin{align*}
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\subst{x}{P}}}
      {\subst{x}{\mprd{Q}{R}}}
      {\mprd{\subst{x}{Q}}{\subst{x}{R}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\subst{x}{\mprd{Q}{S}}}
      {\mprd{\subst{x}{Q}}{\subst{x}{S}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{\Gamma}{\subst{x}{P}}}{\mprd{\subst{x}{Q}}{\subst{x}{R}}}}
      {\subst{x}{\mprd{Q}{S}}}
      {\subst{x}{\slam{Q}{S}{k}}}
      {\slam{\subst{x}{Q}}{\subst{x}{S}}{\subst{x}{k}}}
    }
  \\
& \inference
  { \jterm{\Gamma}{A}{x}
    \jterm{{{{\Gamma}{A}}{P}}{\mprd{Q}{R}}}{\mprd{Q}{S}}{k}
    }
  { \jtermeq
      {{{{\Gamma}{\subst{x}{P}}}{\subst{x}{Q}}}{\subst{x}{R}}}
      {\subst{x}{S}}
      {\subst{x}{\sev{Q}{S}{k}}}
      {\sev{\subst{x}{Q}}{\subst{x}{S}}{\subst{x}{k}}}
    } 
\end{align*}

\subsubsection{The dependent function constructor is compatible with weakening}
\begin{align*}
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{\Gamma}{Q}}{P}}{R}
    }
  { \jfameq
      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}
      {\mprd{A}{\ctxwk{Q}{R}}}
      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{R}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}{\mprd{A}{\ctxwk{Q}{R}}}}
      {\mprd{A}{\ctxwk{Q}{S}}}
      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{S}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jterm{{{{\Gamma}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}{\mprd{A}{\ctxwk{Q}{R}}}}
      {\mprd{A}{\ctxwk{Q}{S}}}
      {\slam{A}{\ctxwk{Q}{S}}{\ctxwk{Q}{k}}}
      {\ctxwk{\mprd{A}{Q}}{\slam{A}{S}{k}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    \jfam{{{{\Gamma}{P}}{Q}}{R}}{S}
    \jterm{{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{R}}}{\mprd{A}{S}}{k}
    }
  { \jtermeq
      {{{{{\Gamma}{A}}{P}}{Q}}{\ctxwk{Q}{R}}}
      {\ctxwk{Q}{S}}
      {\sev{A}{\ctxwk{Q}{S}}{\ctxwk{\mprd{A}{Q}}{k}}}
      {\ctxwk{Q}{\sev{A}{S}{k}}}
    }
\end{align*}

\subsubsection{The dependent function constructor is compatible with substitution}
\begin{align*}
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    \jfam{{{{\Gamma}{A}}{P}}{Q}}{R}
    }
  { \jfameq
      {{\Gamma}{\mprd{A}{P}}}
      {\mprd{A}{\subst{g}{R}}}
      {\subst{\slam{A}{Q}{g}}{\mprd{A}{R}}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    \jfam{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}
    }
  { \jfameq
      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{\subst{g}{R}}}}
      {\mprd{A}{\subst{g}{S}}}
      {\subst{\slam{A}{Q}{g}}{\mprd{A}{S}}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    \jterm{{{{{\Gamma}{A}}{P}}{Q}}{R}}{S}{k}
    }
  { \jtermeq
      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{\subst{g}{R}}}}
      {\mprd{A}{\subst{g}{S}}}
      {\slam{A}{\subst{g}{S}}{\subst{g}{k}}}
      {\subst{\slam{A}{Q}{g}}{\slam{A}{S}{k}}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{Q}{g}
    \jterm{{\Gamma}{\mprd{A}{\ctxext{{P}{Q}}{R}}}}{\mprd{A}{S}}{k}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{P}}{\subst{g}{R}}}
      {\subst{g}{S}}
      {\sev{A}{\subst{g}{S}}{\subst{\slam{A}{Q}{g}}{k}}}
      {\subst{g}{\sev{A}{S}{k}}}
    }
\end{align*}

\subsubsection{The dependent function constructor is compatible with the identity terms}
\begin{align*}
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jtermeq
      {{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{Q}}}
      {\ctxwk{\mprd{A}{Q}}{\mprd{A}{Q}}}
      {\slam{A}{\ctxwk{Q}{Q}}{\idtm{Q}}}
      {\idtm{\mprd{A}{Q}}}
    }
  \\
& \inference
  { \jfam{{{\Gamma}{A}}{P}}{Q}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{P}}{Q}}
      {\ctxwk{Q}{Q}}
      {\sev{A}{\ctxwk{Q}{Q}}{\idtm{\mprd{A}{Q}}}}
      {\idtm{Q}}
    }
\end{align*}

\subsection{Consequences of the rules for the dependent function constructors}

\begin{lem}\label{lem:slam-ext}
The inference rules asserting that lambda abstraction is compatible with term
extension
\begin{align*}
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{{Q}{R}}{h}
    }
  { \jtermeq
      {{\Gamma}{\mprd{A}{P}}}
      {\mprd{A}{\ctxext{Q}{R}}}
      {\cprojfst{\mprd{A}{Q}}{\mprd{A}{R}}{\slam{A}{\ctxext{Q}{R}}{h}}}
      {\slam{A}{Q}{\cprojfst{Q}{R}{h}}}
    }
  \\
& \inference
  { \jterm{{{\Gamma}{A}}{P}}{{Q}{R}}{h}
    }
  { \jtermeq
      {{\Gamma}{\mprd{A}{P}}}
      {\mprd{A}{\ctxext{Q}{R}}}
      {\cprojfst{\mprd{A}{Q}}{\mprd{A}{R}}{\slam{A}{\ctxext{Q}{R}}{h}}}
      {\slam{A}{R}{\cprojsnd{Q}{R}{h}}}
    }
\intertext{and the inference rules asserting that evaluation is compatible with
term extension}
& \inference
  { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{\ctxext{Q}{R}}}{h}
    }
  { \jtermeq
      {{{\Gamma}{A}}{P}}
      {{Q}{R}}
      {\cprojfst{Q}{R}{\sev{A}{\ctxext{Q}{R}}{h}}}
      {\sev{A}{Q}{\cprojfst{\mprd{A}{Q}}{\mprd{A}{R}}{h}}}
    }
  \\
& \inference
  { \jterm{{\Gamma}{\mprd{A}{P}}}{\mprd{A}{\ctxext{Q}{R}}}{h}
    }
  { \jtermeq
      {{{\Gamma}{A}}{P}}
      {{Q}{R}}
      {\cprojsnd{Q}{R}{\sev{A}{\ctxext{Q}{R}}{h}}}
      {\sev{A}{R}{\cprojsnd{\mprd{A}{Q}}{\mprd{A}{R}}{h}}}
    }
\end{align*}
are valid.
\end{lem}

\begin{proof}
We have the judgmental equalities
\begin{align*}
\cprojfst{\mprd{A}{Q}}{\mprd{A}{R}}{\slam{A}{\ctxext{Q}{R}}{h}}
& \jdeq
  \subst{\slam{A}{\ctxext{Q}{R}}{h}}{\ctxwk{\mprd{A}{R}}{\idtm{\mprd{A}{Q}}}}
  \\
& \jdeq 
  \subst{\slam{A}{\ctxext{Q}{R}}{h}}{\ctxwk{\mprd{A}{R}}{\slam{A}{\ctxwk{Q}{Q}}{\idtm{Q}}}}
  \\
& \jdeq
  \subst{\slam{A}{\ctxext{Q}{R}}{h}}{\slam{A}{\ctxwk{R}{{Q}{Q}}}{\ctxwk{R}{\idtm{Q}}}}
  \\
& \jdeq
  \subst{\slam{A}{\ctxext{Q}{R}}{h}}{\slam{A}{\ctxwk{\ctxext{Q}{R}}{Q}}{\ctxwk{R}{\idtm{Q}}}}
  \\
& \jdeq
  \slam{A}{Q}{\subst{h}{\ctxwk{R}{\idtm{Q}}}}
  \\
& \jdeq
  \slam{A}{Q}{\cprojfst{Q}{R}{h}}
\end{align*}
\end{proof}

\begin{rmk}
It follows from these rules that on the projections (i.e.~on the fibrations),
lambda abstraction and evaluation are each other's inverses even though we
have not even postulated that evaluation is a retraction of lambda 
abstraction. Thus, in combination with univalence this should give that
they are homotopically inverse to each other.
\end{rmk}