### 2014-09-19 article-models-ttinternal.tex

\section{Internalizing the theory of contexts families and terms}
One of the guiding ideas behind the design of the theory of contexts, families
and terms was that it would have to be possible to consider internal versions
of the theory. In this section we aim for this internalization. We stress that
we shall not make any further assumptions in this section, and thus that
it is \emph{by default} possible to consider internal models of the theory
of contexts, families and terms in itself. In particular, we do not assume that
there are universes; this is the subject of a later section in this part.

It would be interesting to write out a weak version of pre-universes, internal
to Martin-L\"of type theory with the function extensionality principle. 
To do this, the empty context needs to
be replaced by a contractible type, extension by dependent pair types,
judgmental equalities of terms by identifications and judgmental equalities
of families by equivalences of types. We conjecture that it is possible to
carry this out (in particular to make sure that all the constructions 
type-check). This could serve as a starting point for investigating internal
models without truncatedness assumptions and for investigating internal higher
categories. 
Moreover, one could then extend the notion of `weak' pre-universes
with the requirement that every internal morphism is weakly anodyne. This
could give an internal theory of weak higher groupoids.

\subsection{Extension algebras}\label{sec:extension-algebras}
In this subsection our goal is to define the notion of extension algebras,
which are internal versions of the extension operation of the theory of
contexts, families and terms. In this article, their use will be mainly in
universes. The theory of extension algebras requires the full power (i.e.~all
of the ingredients) of the theory of contexts, families and terms in its
formulation and it is (perhaps surprisingly) quite involved to formulate it.
The definition of extension algebras is given in \autoref{defn:extension-algebras}.

Let $P$ be a family over an extended context $\ctxext{\Gamma}{A}$. We could
mimic extension by requiring to have terms
\begin{align*}
\jalign\jhom{\Gamma}{{A}{P}}{A}{e_0}\\
\jalign\jhom{{\Gamma}{A}}{{P}{\jcomp{}{e_0}{P}}}{P}{e_1}.
\end{align*}

\begin{rmk}
Instead of looking at $e_1$ as a context morphism from $\ctxext{P}{\jcomp{}{e_0}{P}}$
to $P$ in context $\ctxext{\Gamma}{A}$, one could also look at $e_1$ as a 
morphism \emph{over $\cprojfstf{A}{P}$} in context $\Gamma$, as indicated in the
following diagram:
\begin{equation*}
\begin{tikzcd}
P
  \ar[fib]{d}
& \jcomp{}{e_0}{P}
  \ar[fib]{d}
  \ar{l}[swap]{e_1}
  \ar{r}
& P
  \ar[fib]{d}
  \\
A
& \ctxext{A}{P}
  \ar{l}{\cprojfstf{A}{P}}
  \ar{r}[swap]{e_0}
& A
\end{tikzcd}
\end{equation*}
This makes it clear that $e_1$ takes a family over an extended context as an
argument. The extended context consists of a `base part' and a `family part'. 
The output of $e_1$ is a new (extended) family over that base part. Forgetting 
the family part is what the projection takes care of.
\end{rmk}

Extension also satisfies the properties explained in \autoref{comp-ee}, so we
must find the two judgmental equalities for $e_0$ and $e_1$ interpreting those. 
The first of these judgmental equalities is easy to give: it says that the
following diagram in context $\Gamma$ commutes judgmentally:
\begin{equation}\label{eq:extalg-eq1}
\begin{tikzcd}[column sep=huge]
\ctxext{A}{{P}{\jcomp{}{e_0}{P}}} 
  \ar{d}[swap]{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}
    } 
  \ar{r}{\jvcomp{}{\idtm{A}}{e_1}
    } 
  & \ctxext{A}{P} \ar{d}{e_0}\\
\ctxext{A}{P} \ar{r}[swap]{e_0} & A
\end{tikzcd}
\end{equation}

In the following lemma we find an equivalent reformulation of the condition in 
\autoref{eq:extalg-eq1}.

\begin{lem}\label{lem:extalg-twins}
Suppose we have $\jfam{\Gamma}{A}$, $\jfam{{\Gamma}{A}}{P}$, 
$\jhom{\Gamma}{{A}{P}}{A}{e_0}$ and 
$\jhom{{\Gamma}{A}}{{P}{\jcomp{}{P}{e_0}}}{P}{e_1}$. 
Then the inference rules
\begin{align*}
& \inference
  { }
  { \jhomeq
      {\Gamma}
      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
      {A}
      {\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}}
      {\jcomp{}{e_0}{e_0}}
    }
\intertext{and}
& \inference
  { }
  { \jhomeq
      {\Gamma}
      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
      {A}
      {\jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{e_0}}
      {\jcomp{}{e_1}{e_0}}
    }
\end{align*}
As a consequence, the inference rules
\begin{align*}
& \inference
  { \jhomeq
      {\Gamma}
      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
      {A}
      {\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}}
      {\jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{e_0}}
    }
  { \jhomeq
      {\Gamma}
      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
      {A}
      {\jcomp{}{e_0}{e_0}}
      {\jcomp{}{e_1}{e_0}}
    }
\intertext{and}
& \inference
  { \jhomeq
      {\Gamma}
      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
      {A}
      {\jcomp{}{e_0}{e_0}}
      {\jcomp{}{e_1}{e_0}}
    }
  { \jhomeq
      {\Gamma}
      {\ctxext{A}{{P}{\jcomp{}{e_0}{P}}}}
      {A}
      {\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}}
      {\jcomp{}{\jvcomp{}{\idtm{A}}{e_1}}{e_0}}
    }
\end{align*}
are valid.
\end{lem}

\begin{proof}
To prove the first inference rule, note that we have the judgmental equalities
\begin{align*}
\jcomp{}{\jvcomp{}{e_0}{\idtm{\jcomp{}{e_0}{P}}}}{e_0}
& \jdeq
  \jcomp{}{\idtm{\jcomp{}{e_0}{P}}}{\jcomp{}{e_0}{e_0}}
  \tag{by \autoref{lem:composition-threesome}}
  \\
& \jdeq
  \jcomp{}{e_0}{e_0}
  \tag{by \autoref{precomp-idtm-t}}
\end{align*}
The second inference rule follows immediately from \autoref{lem:edom-edom-comp}.
\end{proof}

The second condition we will put on context and family extension essentially 
asserts the quadruple $(P,\jcomp{}{e_0}{P},e_1,\jcomp{}{e_0}{e_1})$ also satisfies 
\autoref{eq:extalg-eq1}. Here we consider `higher' families, i.e.~families over
families over families, which are described by the family 
$\jcomp{}{e_0}{P}$ over $\ctxext{{\Gamma}{A}}{P}$. The family
extension $e_1$ becomes context extension when we consider the quadruple
$(P,\jcomp{}{e_0}{P},e_1,\jcomp{}{e_0}{e_1})$ and likewise, the morphism
$\jcomp{}{e_0}{e_1}$ becomes family extension. Thus, the second requirement is 
that the diagram
\begin{equation}\label{eq:extalg-eq2}
\begin{tikzcd}[column sep=huge]
\ctxext{P}{{\jcomp{}{e_0}{P}}{\jcomp{}{e_1}{\jcomp{}{e_0}{P}}}} 
  \ar{d}[swap]{\jvcomp{}{e_1}{\idtm{\jcomp{}{e_1}{\jcomp{}{e_0}{P}}}}
    } 
  \ar{r}{\jvcomp{}{\idtm{P}}{\jcomp{}{e_0}{e_1}}
    } 
  & \ctxext{P}{\jcomp{}{e_0}{P}} \ar{d}{e_1}\\
\ctxext{P}{\jcomp{}{e_0}{P}} \ar{r}[swap]{e_1} & P
\end{tikzcd}
\end{equation}

\begin{defn}\label{defn:extension-algebras}
An \emph{extension algebra $\extalg{A}$ in context $\Gamma$} is a quintuple
\begin{equation*}
\unfold{\extalg{A}}
\end{equation*}
consisting of a family $\jextalgctx{\Gamma}{A}$ of \emph{$\extalg{A}$-contexts}, 
a family $\jextalgfam{\Gamma}{A}$ of \emph{$\extalg{A}$-families}, a family 
$\jextalgtm{\Gamma}{A}$ of \emph{$\extalg{A}$-terms}, a term
$\jextalgctxext{\Gamma}{A}$ called the \emph{$\extalg{A}$-context extension}
and a term $\jextalgfamext{\Gamma}{A}$ called the \emph{$\extalg{A}$-family
extension}, satisfying the judgmental equalities of
\autoref{eq:extalg-eq1,eq:extalg-eq2}.
\begin{comment}
We define the judgment
\begin{equation*}
\jextalg{\Gamma}{A}
\end{equation*}
asserting that $\extalg{A}$ is an \emph{extension algebra in context $\Gamma$}
to be the conjunction of the following seven judmgents:
\begin{align*}
\jalign\jextalgctx{\Gamma}{A}
  \\
\jalign\jextalgfam{\Gamma}{A}
  \\
\jalign\jextalgtm{\Gamma}{A}
  \\
\jalign\jextalgctxext{\Gamma}{A}
  \\
\jalign\jextalgfamext{\Gamma}{A}
  \\
\jalign\jhomeq
  { \Gamma}
  { {\cftalgc{\cftalg{A}}}
    { {\cftalgf{\cftalg{A}}}
      {\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
      }
    }
  { \cftalgc{\cftalg{A}}}
  { \jcomp{}
      { \jvcomp{}
          {\cftctxext[\cftalg{A}]}
          {\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}
        }
      { \cftctxext[\cftalg{A}]}
    }
  { \jcomp{}
      { \jvcomp{}
          {\idtm{\cftalgc{\cftalg{A}}}}
          {\cftfamext[\cftalg{A}]}
        }
      { \cftctxext[\cftalg{A}]}
    }
  \\
\jalign\jhomeq
  { {\Gamma}{\cftalgc{\cftalg{A}}}}
  { { \cftalgf{\cftalg{A}}}
    { { \jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
      { \jcomp{}
          {\cftctxext[\cftalg{A}]}
          {{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
        }
      }
    }
  { \cftalgf{\cftalg{A}}}
  { \jcomp{}
      { \cftfamext[\cftalg{A}]}
      { \jvcomp{}
          {\cftfamext[\cftalg{A}]}
          {\idtm{\jcomp{}{\cftfamext[\cftalg{A}]}{{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}
    }
  { \jcomp{}
      { \cftfamext[\cftalg{A}]}
      { \jvcomp{}
          {\idtm{\cftalgf{\cftalg{A}}}}
          {\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext[\cftalg{A}]}}
        }
    }
\end{align*}
In other words, an extension algebra $\extalg{A}$ in context $\Gamma$
is a quintuple $\unfold{\extalg{A}}$ 
satisfying the judgmental equalities displayed in the diagrams in
\autoref{eq:extalg-eq1,eq:extalg-eq2}.
\end{comment}
Extension algebras are judgmentally
equal if they are componentwise judgmentally equal.

When $\extalg{A}$ is an extension algebra in context $\Gamma$, we also refer to
$\cftctxext[\cftalg{A}]$ as the \emph{context extension of $\extalg{A}$} and to
$\cftfamext[\cftalg{A}]$ as the \emph{family extension of $\extalg{A}$}.
\end{defn}

The following lemma both provides intuition behind the judgmental equalities
we have required for context and family extension and it proves that context 
and family extension in fact satisfy the compatibility rules stated in 
\autoref{comp-ee}.

\begin{lem}
For $\gamma:\cftalgc{\cftalg{A}}$, 
$a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and 
$p:\subst{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{\cftalgf{\cftalg{A}}}$ 
we have the judgmental equality
\begin{align*}
\apply\cftctxext[\extalg{A}]{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}
& \jdeq
  \apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}.
  \intertext{%
and when we also have 
$q: \subst
      {\apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}}
      {\cftalgf{\extalg{A}}}$, %
we have the judgmental equality}
\apply\cftfamext[\extalg{A}]{\gamma}{a}{%
  \apply\cftfamext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}{q}}
& \jdeq
  \apply\cftfamext[\extalg{A}]{\gamma}{%
    \apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}{q}.
\end{align*}
\end{lem}

\begin{proof}
Both proofs are simple calculations. For the first judgmental equality we have
\begin{align*}
\apply\cftctxext[\extalg{A}]{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}
& \jdeq
  \subst
    {\tmext{\gamma}{\apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}}
    {\cftctxext[\cftalg{A}]}
  \tag{by \autoref{comp-es}}
  \\
& \jdeq 
  \subst
    { {p}
      { {a}
        { {\gamma}
          {\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}
          }
        }
      }
    { \cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst{p}{{a}{{\gamma}{\jcomp{}{\jvcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext[\cftalg{A}]}}{\cftctxext[\cftalg{A}]}}}}
  \\
& \jdeq
  \subst{p}{{a}{{\gamma}{\jcomp{}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}{\cftctxext[\cftalg{A}]}}}}
  \\
& \jdeq 
  \subst{{p}{{a}{{\gamma}{\jvcomp{}{\cftctxext[\cftalg{A}]}{\idtm{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}}}}}}{\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \subst
    {\tmext{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}}
    {\cftctxext[\cftalg{A}]}
  \\
& \jdeq
  \apply\cftctxext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}.
\end{align*}
and to prove the second judgmental equality we calculate
\begin{align*}
\apply\cftfamext[\extalg{A}]{\gamma}{a}{%
  \apply\cftfamext[\extalg{A}]{\apply\cftctxext[\extalg{A}]{\gamma}{a}}{p}{q}}
& \jdeq
  \apply\cftfamext[\extalg{A}]{\gamma}{%
    \apply\cftfamext[\extalg{A}]{\gamma}{a}{p}}{q}.\qedhere
\end{align*}
\end{proof}

There is a trivial class of examples of extension algebras we can give right
away. More examples will be introduced by universes, later on.

\begin{eg}
Let $A$ be a family in context $\Gamma$. Then the quadruple
\begin{equation*}
(A,\emptyf,\emptyf,\idtm{A},\emptytm)
\end{equation*}
is an extension algebra in context $\Gamma$, as is the quadruple
\begin{equation*}
(\emptyf,A,\emptyf,\emptytm,\idtm{A}).
\end{equation*}
Also, the quadruple
\begin{equation*}
(A,\ctxwk{A}{A},\emptyf,\cprojfstf{A}{\ctxwk{A}{A}},\cprojfstf{\ctxwk{A}}{\ctxwk{A}{{A}{A}}})
\end{equation*}
is an extension algebra in context $\Gamma$.
\end{eg}

As a consequence of the following theorem, every extension algebra gives rise
to infinitely many extension algebras by constructing the extension algebra
of families of $\cftalg{A}$ for each extension algebra $\cftalg{A}$. Notice
that this also explains that the family $\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}$
over $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$ is the family
of families over families. Likewise, higher families are obtained by pulling
back more times along $\cftctxext[\cftalg{A}]$.

\begin{thm}\label{thm:extalg-fam}
Suppose that $\extalg{A}$ is an extension algebra in context
$\Gamma$. Then 
\begin{equation*}
\cftfamalg{\cftalg{A}}
\defeq
( \cftalgf{\cftalg{A}},
  \jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}},
  \jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgt{\cftalg{A}}},
  \cftfamext,
  \jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamext})
\end{equation*}
is an extension algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
\end{thm}

\begin{proof}
In this proof, we shall use the short-hand notations $\cftctxext$ and $\cftfamext$
only to refer to $\cftctxext[\cftalg{A}]$ and $\cftfamext[\cftalg{A}]$, respectively,
and not to $\cftctxext[\cftfamalg{\cftalg{A}}]$ or 
$\cftfamext[\cftfamalg{\cftalg{A}}]$.

We first need to verify that the domain of the morphism 
$\jcomp{}{\cftctxext}{\cftfamext}$ is indeed
$\ctxext{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftfamext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}$. 
This follows from the judgmental equality
$\jcomp{}{\cftfamext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}\jdeq
\jcomp{}{\cftctxext}{{}{\cftctxext}{\cftalgf{\cftalg{A}}}}$, which we have proved in
\autoref{lem:extalg-twins}. Notice how the diagram in \autoref{eq:extalg-eq2} is
of exactly the right sort, so the quadruple
$(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})$
satisfies its version of \autoref{eq:extalg-eq1}. It is left to verify that the diagram
\begin{small}
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxext
  { \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
    }
  { { \jcomp{}{\cftfamext}{%
        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
        }
      }
    { \jcomp{}{\cftfamext}{%
        \jcomp{}{\cftfamext}{%
          \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
          }
        }
      }
    } 
  \ar{r}
    { \jvcomp{}{\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{%
        \jcomp{}{\cftfamext}{%
          \jcomp{}{\cftctxext}{\cftfamext}}}}
  \ar{d}[swap]{
    \jvcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
      \idtm{
        \jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
          \jcomp{}{\cftfamext}{%
            \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
            }
          }
        }
      }
    }
& \ctxext
    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
    {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}} 
  \ar{d}{\jcomp{}{\cftctxext}{\cftfamext}}
  \\
\ctxext
  {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}} 
  \ar{r}[swap]{\jcomp{}{\cftctxext}{\cftfamext}} 
& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
\end{tikzcd}
\end{equation*}
\end{small}%
commutes judgmentally; this diagram is the version of \autoref{eq:extalg-eq2}
for the quadruple
$(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},\cftfamext,\jcomp{}{\cftctxext}{\cftfamext})$. Note
that this follows from \autoref{eq:extalg-eq2} provided that we can show that
\begin{align}
\jvcomp{}{\idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{%
  \jcomp{}{\cftfamext}{%
    \jcomp{}{\cftctxext}{\cftfamext}}}
& \jdeq
  \jcomp{}{\cftctxext}{%
    \jvcomp{}{\idtm{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
    }
  \label{eq:extalg-infty1}
  \\
\jvcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
  \idtm{
    \jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{%
      \jcomp{}{\cftfamext}{%
        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
        }
      }
    }
  }
& \jdeq
\jcomp{}{\cftctxext}{%
  \jvcomp{}{\cftfamext}{%
    \idtm{
      \jcomp{}{\cftfamext}{%
        \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
        }
      }
    }
  }
  \label{eq:extalg-infty2}
\end{align}
Note that \autoref{eq:extalg-infty1} follows if we can show that
\begin{equation*}
\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftfamext}}
  \jdeq
  \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftfamext}}.
\end{equation*}
This is a special case of \autoref{lem:extalg-twins}. The second judgmental
equality, \autoref{eq:extalg-infty2}, is trivial.
\end{proof}

\begin{thm}\label{thm:extalg-wk}
Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
$\jfam{\Gamma}{A}$ be a family of contexts. Then the quintuple
\begin{equation*}
\ctxwk{A}{\extalg{Q}}
  \defeq
  ( \ctxwk{A}{\cftalgc{\cftalg{Q}}},
    \ctxwk{A}{\cftalgf{\cftalg{Q}}},
    \ctxwk{A}{\cftalgt{\cftalg{Q}}},
    \ctxwk{A}{\cftctxext[\cftalg{Q}]},
    \ctxwk{A}{\cftfamext[\cftalg{Q}]})
\end{equation*}
is an extension algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{thm}

\begin{proof}
The proof follows from the fact that weakening by $A$ is compatible with all
the involved operations.
\end{proof}

\begin{thm}\label{thm:extalg-subst}
Let $\extalg{Q}$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$ be a term. Then the quintuple
\begin{equation*}
\subst{x}{\cftalg{Q}}
  \defeq
  ( \subst{x}{\cftalgc{\cftalg{Q}}},
    \subst{x}{\cftalgf{\cftalg{Q}}},
    \subst{x}{\cftalgt{\cftalg{Q}}},
    \subst{x}{\cftctxext[\cftalg{Q}]},
    \subst{x}{\cftfamext[\cftalg{Q}]})
\end{equation*}
is an extension algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{thm}

\begin{proof}
The proof follows from the fact that substitution with $x$ is compatible with
all the involved operations.
\end{proof}

\begin{cor}
Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$
and let $\jhom{\Gamma}{A}{B}{f}$. Then the quintuple
\begin{equation*}
\jcomp{A}{f}{\extalg{Q}}
  \defeq
  ( \jcomp{A}{f}{\cftalgc{\cftalg{Q}}},
    \jcomp{A}{f}{\cftalgf{\cftalg{Q}}},
    \jcomp{A}{f}{\cftalgt{\cftalg{Q}}},
    \jcomp{A}{f}{\cftctxext[\cftalg{Q}]},
    \jcomp{A}{f}{\cftfamext[\cftalg{Q}]})
\end{equation*}
is an extension algebra in context $\ctxext{\Gamma}{A}$.
\end{cor}

\begin{cor}
Let $\extalg{A}$ be an extension algebra in context $\Gamma$. Then
$\extfamalg{\extfamalg{\extalg{A}}}$ and $\jcomp{}{\cftctxext[\extalg{A}]}{\extfamalg{\extalg{A}}}$ are
judgmentally equal extension algebras in context $\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$.
\end{cor}

\begin{lem}
Let $\extalg{Q}$ be an extension algebra in context $\ctxext{\Gamma}{B}$ and let
$\jfam{\Gamma}{A}$. Then we have the judgmental equality
\begin{equation*}
\ctxwk{A}{\extfamalg{\extalg{Q}}}\jdeq\extfamalg{\ctxwk{A}{\extalg{Q}}}
\end{equation*}
of extension algebras in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{lem}

\begin{lem}
Let $\extalg{Q}$ be an extension algebra in context $\ctxext{{\Gamma}{A}}{P}$ 
and let $\jterm{\Gamma}{A}{x}$. Then we have the judgmental equality
\begin{equation*}
\subst{x}{\extfamalg{\extalg{Q}}}\jdeq\extfamalg{\subst{x}{\extalg{Q}}}
\end{equation*}
of extension algebras in context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{lem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Extension homomorphisms}

\begin{defn}
An \emph{extension homomorphism $\cfthom{f}$ from $\extalg{A}$ to
$\extalg{B}$ in context $\Gamma$} is a triple $\unfold{\cfthom{f}}$ consisting of
\begin{align*}
\jalign\jhom
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  \\
\jalign\jfhom
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\cftalgc{\cftalg{B}}}
  {\cfthomc{\cfthom{f}}}
  {\cftalgf{\cftalg{A}}}
  {\cftalgf{\cftalg{B}}}
  {\cfthomf{\cfthom{f}}}
  \\
\jalign\jfhom
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
  {\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  {\cftalgt{\cftalg{A}}}
  {\cftalgt{\cftalg{B}}}
  {\cfthomt{\cfthom{f}}}
\end{align*}
for which the diagrams
\begin{equation}\label{eq:exthom1}
\begin{tikzcd}
\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  \ar{d}[swap]{\cftctxext[\cftalg{A}]}
& \ctxext{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}
  \ar{d}{\cftctxext[\cftalg{B}]}
  \\
\cftalgc{\cftalg{A}}
  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
& \cftalgc{\cftalg{B}}
\end{tikzcd}
\end{equation}
and
\begin{equation}\label{eq:exthom2}
\begin{tikzcd}[column sep=huge]
\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}}
  \ar{r}{\jvcomp{}{\cfthomf{\cfthom{f}}}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}}}}
  \ar{d}[swap]{\cftfamext[\cftalg{A}]}
& \jcomp{}{\cfthomc{\cfthom{f}}}{\ctxext{\cftalgf{\cftalg{B}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
  \ar{d}{\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamext[\cftalg{B}]}}
  \\
\cftalgf{\cftalg{A}}
  \ar{r}[swap]{\cfthomf{\cfthom{f}}}
& \jcomp{}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}
\end{tikzcd}
\end{equation}
commute judgmentally.
\end{defn}

\begin{rmk}
To see that the upper morphism in the diagram of \autoref{eq:exthom2} has
indeed the indicated codomain provided that the diagram of \autoref{eq:exthom1}
commutes judgmentally, note that we have the judgmental equalities
\begin{align*}
\jcomp{}{\cfthomf{\cfthom{f}}}{\jcomp{}{\cfthomc{\cfthom{f}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}}
& \jdeq 
  \jcomp{}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}{\jcomp{}{\cftctxext[\cftalg{B}]}{\cftalgf{\cftalg{B}}}}
  \tag{by \autoref{lem:composition-threesome}}
  \\
& \jdeq
  \jcomp{}{\jcomp{}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}{\cftctxext[\cftalg{B}]}}{\cftalgf{\cftalg{B}}}
  \tag{by \autoref{lem:jcomp-jcomp}}
  \\
& \jdeq
  \jcomp{}{\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomc{\cfthom{f}}}}{\cftalgf{\cftalg{B}}}
  \tag{by \autoref{eq:exthom1}}
  \\
& \jdeq
  \jcomp{}{\cftctxext[\cftalg{A}]}{\jcomp{}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}}.
  \tag{by \autoref{lem:jcomp-jcomp}}
\end{align*}
and we indeed have the morphism $\jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}}$ from 
$\jcomp{}{\cftctxext[\cftalg{A}]}{\cftalgf{\cftalg{A}}}$ to $\jcomp{}{\cftctxext[\cftalg{A}]}{\jcomp{}{\cfthomc{\cfthom{f}}}{\cftalgf{\cftalg{B}}}}$.
\end{rmk}

\begin{thm}
Let $\extalg{A}$ be an extension algebra in context $\Gamma$. Then the triple
\begin{equation*}
\cftidhom{\extalg{A}}\defeq\unfold{\cftidhom{\extalg{A}}}
\end{equation*}
is an extension homomorphism from $\cftalg{A}$ to $\cftalg{A}$ in context
$\Gamma$.
\end{thm}

\begin{thm}
Let $\cfthom{f}$ be an extension homomorphism from $\cftalg{A}$ to $\cftalg{B}$
in context $\Gamma$. Then
\begin{equation*}
\cftfamhom{\cfthom{f}}
  \defeq
  ( \cfthomf{\cfthom{f}},
    \jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomf{\cfthom{f}}},
    \jcomp{}{\cftctxext[\cftalg{A}]}{\cfthomt{\cfthom{f}}})
\end{equation*}
is an extension homomorphism from $\cftfamalg{\cftalg{A}}$ to 
$\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftalg{B}}}$
in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
\end{thm}

\begin{proof}
We have to verify that the diagrams in \autoref{eq:exthom1,eq:exthom2}
commute for $\cftfamhom{\cfthom{f}}$. Unfolding the ingredients of \autoref{eq:exthom1}
for $\cftfamhom{\cfthom{f}}$ gives us quite directly \autoref{eq:exthom2}. The
diagram in \autoref{eq:exthom2} for $\cftfamhom{\cfthom{f}}$ is judgmentally equal
to the pullback of everything in the diagram in \autoref{eq:exthom2} for
$\cfthom{f}$ by $\epsilon_0$, and therefore it commutes too.
\end{proof}

\begin{lem}
Let $\cfthom{f}$ be an extension homomorphism from $\cftalg{A}$ to $\cftalg{B}$
in context $\Gamma$. Then we have the judgmental equality
\begin{equation*}
\cftfamhom{\cftfamhom{\cfthom{f}}}
  \jdeq
  \jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamhom{\cfthom{f}}}
\end{equation*}
of extension homomorphisms from $\cftfamalg{\cftfamalg{\cftalg{A}}}$ to
$\jcomp{}{\cfthomf{\cfthom{f}}}
  {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftfamalg{\cftalg{B}}}}}$ in
context $\ctxext{{\Gamma}{A}}{\cftalgf{\cftalg{A}}}$.
\end{lem}

\begin{defn}
Let $\cfthom{f}$ and $\cfthom{g}$ be extension homomorphisms from
$\extalg{A}$ to $\extalg{B}$ and from $\extalg{B}$ to $\extalg{C}$, respectively.
We define the composition
\begin{equation*}
\cfthomcomp{\cfthom{f}}{\cfthom{g}}
  \defeq
  \unfold{\cfthomcomp{\cfthom{f}}{\cfthom{g}}}.
\end{equation*}
In other words, extension homomorphisms are composed by taking the horizontal 
rectangles in the diagram
\begin{equation*}
\begin{tikzcd}
\extalgt{\extalg{A}} 
  \ar[fib]{d}
  \ar{r}{\cfthomt{\cfthom{f}}}
& \extalgt{\extalg{B}} 
  \ar[fib]{d}
  \ar{r}{\cfthomt{\cfthom{g}}}
& \extalgt{\extalg{C}}
  \ar[fib]{d}
  \\
\extalgf{\extalg{A}} 
  \ar[fib]{d}
  \ar{r}{\cfthomf{\cfthom{f}}}
& \extalgf{\extalg{B}} 
  \ar[fib]{d}
  \ar{r}{\cfthomf{\cfthom{g}}}
& \extalgf{\extalg{C}}
  \ar[fib]{d}
  \\
\extalgc{\extalg{A}}
  \ar{r}{\cfthomc{\cfthom{f}}}
& \extalgc{\extalg{B}}
  \ar{r}{\cfthomc{\cfthom{g}}}
& \extalgc{\extalg{C}}
\end{tikzcd}
\end{equation*}
\end{defn}

\begin{rmk}
It follows from \autoref{lem:jcomp-jcomp,lem:jfcomp-jfcomp} that composition
of extension homomorphisms is associative.
\end{rmk}

\begin{thm}
Let $\cfthom{f}$ and $\cfthom{g}$ be extension homomorphisms from
$\extalg{A}$ to $\extalg{B}$ and from $\extalg{B}$ to $\extalg{C}$, respectively.
Then $\cfthomcomp{\cfthom{f}}{\cfthom{g}}$ is an extension homomorphism from
$\extalg{A}$ to $\extalg{C}$. 
\end{thm}

\begin{proof}
Both judgmental equalities are applications of the interchange law for composition,
\autoref{lem:composition-interchange}. 
\end{proof}

\begin{lem}
Let $\exthom{g}$ be an extension homomorphism from $\extalg{Q}$ to $\extalg{R}$ in
context $\ctxext{\Gamma}{B}$ and let $\jfam{\Gamma}{A}$. Then the triple
\begin{equation*}
\ctxwk{A}{\cfthom{g}}
  \defeq
  ( \ctxwk{A}{\cfthomc{\cfthom{g}}},
    \ctxwk{A}{\cfthomf{\cfthom{g}}},
    \ctxwk{A}{\cfthomt{\cfthom{g}}})
\end{equation*}
is an extension homomorphism from $\ctxwk{A}{\cftalg{Q}}$ to $\ctxwk{A}{\cftalg{R}}$
in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{lem}

\begin{lem}
Let $\cfthom{g}$ be an extension homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
context $\ctxext{{\Gamma}{A}}{P}$ and let $\jterm{\Gamma}{A}{x}$. Then the
triple
\begin{equation*}
\subst{x}{\cfthom{g}}
  \defeq
  ( \subst{x}{\cfthomc{\cfthom{g}}},
    \subst{x}{\cfthomf{\cfthom{g}}},
    \subst{x}{\cfthomt{\cfthom{g}}})
\end{equation*}
is an extension homomorphism from $\subst{x}{\cftalg{Q}}$ to $\subst{x}{\cftalg{R}}$
in context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{lem}

The following theorem explains how context extension can be seen as an extension
homomorphism. Note that in combination with \autoref{thm:extalg-fam}, this also
explains how family extension can be seen as an extension homomorphism.

\begin{thm}
Let $\extalg{A}$ be an extension algebra in context $\Gamma$. Then
\begin{align*}
& ( \ctxext{\extalgc{\extalg{A}}}{\extalgf{\extalg{A}}},
    \jcomp{}{\cftctxext}{\extalgf{\extalg{A}}},
    \jvcomp{}{\ctxwk{\extalgf{\extalg{A}}}{\idtm{\extalgc{\extalg{A}}}}}{\cftfamext},
    \jcomp{}{\cftctxext}{\cftfamext}
    )
\intertext{and}
& ( \ctxext{\extalgc{\extalg{A}}}{\extalgf{\extalg{A}}},
    \jcomp{}{\cftctxext}{\extalgf{\extalg{A}}},
    \jvcomp{}{\cftctxext}{\idtm{\jcomp{}{\cftctxext}{\extalgf{\extalg{A}}}}},
    \jcomp{}{\cftctxext}{\cftfamext}
    )
\end{align*}
are extension algebras in context $\Gamma$ and
\begin{equation*}
\boldsymbol{\cftctxext}\defeq ( \cftctxext,
  \idtm{\jcomp{}{\cftctxext}{\extalgf{\extalg{A}}}}
  )
\end{equation*}
is an extension homomorphism from both of them to $\extalg{A}$. 
\end{thm}

As with ordinary morphisms, extension homomorphism can be considered \emph{over} a
morphism.

\begin{defn}
Let $\jhom{\Gamma}{A}{B}{f}$ be a morphism, let $\extalg{P}$ be an extension
algebra in context $\ctxext{\Gamma}{A}$ and let $\extalg{Q}$ be an extension
algebra in context $\ctxext{\Gamma}{B}$. An \emph{extension homomorphism from
$\cftalg{P}$ to $\cftalg{Q}$ over $f$ in context $\Gamma$} is defined to be
an extension homomorphism from $\extalg{P}$ to $\jcomp{A}{f}{\extalg{Q}}$ in
context $\ctxext{\Gamma}{A}$. 
\end{defn}

\begin{eg}
Let $\cfthom{f}$ be an extension homomorphism from $\extalg{A}$ to $\extalg{B}$
in context $\Gamma$. Then $\cftfamhom{\cfthom{f}}$ is an
extension homomorphism from $\extfamalg{\extalg{A}}$ to $\extfamalg{\extalg{B}}$ over
$\cfthomc{\cfthom{f}}$ in context $\Gamma$.
\end{eg}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Empty family algebras}
The notion of empty family algebras that we will
study in this subsection will be extension algebras which also have an empty 
context $\cftempc{\cftalg{A}}$ and an empty family $\cftempf{\cftalg{A}}$ 
satisfying (among other conditions) that families over the empty context in
$\cftalg{A}$ are just the contexts of $\cftalg{A}$.

\begin{defn}
An \emph{empty family algebra $\cftalg{A}$ in context $\Gamma$} is a septuple
\begin{equation*}
\unfold{\cftalg{A}}
\end{equation*}
where the quintuple $\unfold{\extalg{A}}$ is an extension
algebra in context $\Gamma$, and where
\begin{align*}
\jalign\jterm{\Gamma}{\cftalgc{\cftalg{A}}}{\cftempc{\cftalg{A}}}
  \\
\jalign\jterm{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}{\cftempf{\cftalg{A}}}
\end{align*}
which satisfy the following judgmental equalities
\begin{enumerate}
\item Families over the empty context in $\cftalg{A}$ are contexts of $\cftalg{A}$:
\begin{align}
\jalign\jfameq
  {\Gamma}
  {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\cftalgc{\cftalg{A}}}
  \label{empalg-eq1}
\end{align}
\item The empty family over the empty context of $\cftalg{A}$ is the empty context of $\cftalg{A}$:
\begin{align}
\jalign\jtermeq
  {\Gamma}
  {\cftalgc{\cftalg{A}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftempf{\cftalg{A}}}}
  {\cftempc{\cftalg{A}}}.
  \label{empalg-eq2}
\end{align}
\item Context extension of $\cftalg{A}$ is compatible with the empty context and
family of $\cftalg{A}$:
\begin{align}
\jalign\jtermeq
  {{\Gamma}{\cftalgc{\cftalg{A}}}}
  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
  {\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq1}
  \\
\jalign\jtermeq
  {{\Gamma}{\cftalgc{\cftalg{A}}}}
  {\ctxwk{\cftalgc{\cftalg{A}}}{\cftalgc{\cftalg{A}}}}
  {\subst{\cftempf{\cftalg{A}}}{\cftctxext[\cftalg{A}]}}
  {\idtm{\cftalgc{\cftalg{A}}}}
  \label{cftalg-eq2}
\end{align}
\item Family extension of $\cftalg{A}$ is compatible with the empty family and the empty family
over families of $\cftalg{A}$:
\begin{align}
\jalign\jtermeq
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
  {\idtm{\cftalgf{\cftalg{A}}}}
  \label{cftalg-eq3}
  \\
\jalign\jtermeq
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
  {\idtm{\cftalgf{\cftalg{A}}}}
  \label{cftalg-eq4}
\end{align}
\item Family extension over the empty context of $\cftalg{A}$ is context extension
of $\cftalg{A}$:
\begin{align}
\jalign\jhomeq
  {\Gamma}
  {{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  {\cftalgc{\cftalg{A}}}
  {\subst{\cftempc{\cftalg{A}}}{\cftfamext}}
  {\cftctxext}
  \label{cftalg-eq5}
\end{align}
\end{enumerate}
\end{defn}

\begin{rmk}\label{rmk:cftalg-defn}
We need to verify that the judgmental equalities
\autoref{cftalg-eq1,cftalg-eq2,cftalg-eq3,cftalg-eq4,cftalg-eq5}
are indeed well-typed (i.e.~compare two terms of the same type).
\begin{enumerate}
\item \autoref{lem:ehom-subst} implies that the morphisms 
$\subst{\cftempc{\cftalg{A}}}{\cftctxext}$ and
$\subst{\cftempf{\cftalg{A}}}{\cftctxext}$ both go from $\cftalgc{\cftalg{A}}$ 
to $\cftalgc{\cftalg{A}}$.
\item \label{rmkenum:famfamempf-jdeq-fam}
Now we have the judgmental equalities
\begin{align*}
\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
& \jdeq
  \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftalgf{\cftalg{A}}}
  \tag{by \autoref{lem:ehom-subst}}
  \\
& \jdeq
  \jcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}
  \tag{by \autoref{cftalg-eq1}}
  \\
& \jdeq
  \cftalgf{\cftalg{A}}
  \tag{by \autoref{precomp-idtm-c}}
\end{align*}
Therefore we can apply \autoref{lem:ehom-subst} with the triple
$(\cftalgf{\cftalg{A}},\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},
\cftempf{\cftalg{A}})$ and the morphism $\cftfamext$ to see that both 
$\subst{\cftempf{\cftalg{A}}}{\cftfamext}$ and
$\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}$
are morphisms from $\cftalgf{\cftalg{A}}$ to $\cftalgf{\cftalg{A}}$.
\item The morphism $\subst{\cftempc{\cftalg{A}}}{\cftfamext}$ goes from
$ \subst
    {\cftempc{\cftalg{A}}}
    {\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}
  $
to $\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}$. For the domain we
have the judgmental equalities
\begin{align*}
  \subst
    {\cftempc{\cftalg{A}}}
    {\ctxext{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}
& \jdeq
  \ctxext
    {\subst{\cftempc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
    {\subst{\cftempc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}
  \tag{by \autoref{comp-se-c}}
  \\
& \jdeq
  \ctxext
    {\cftalgc{\cftalg{A}}}
    {\subst{\cftempc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}
  \tag{by \autoref{empalg-eq1}}
  \\
& \jdeq
  \ctxext
    {\cftalgc{\cftalg{A}}}
    {\cftalgf{\cftalg{A}}}
  \tag{by \autoref{rmkenum:famfamempf-jdeq-fam} above}
\end{align*}
The codomain of $\subst{\cftempc{\cftalg{A}}}{\cftfamext}$
is judgmentally equal to $\cftalgc{\cftalg{A}}$ by
\autoref{empalg-eq1}.
\end{enumerate}
\end{rmk}

We can extend the results of 
\autoref{thm:extalg-fam,thm:extalg-wk,thm:extalg-subst} to empty family algebras.

\begin{thm}\label{thm:cftalg-fam}
Let $\cftalg{A}$ be an empty family algebra
in context $\Gamma$. Then the septuple
\begin{equation*}
\cftfamalg{\cftalg{A}}
  \defeq
  ( \cftalgf{\cftalg{A}},
    \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}},
    \jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}},
    \cftfamext,
    \jcomp{}{\cftctxext}{\cftfamext},
    \cftempf{\cftalg{A}},
    \jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}})
\end{equation*}
is an empty family algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
\end{thm}

\begin{proof}
\begin{enumerate}
\item The judgmental equality
\begin{equation*}
\jfameq{{\Gamma}{\cftalgc{\cftalg{A}}}}{\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}}{\cftalgf{\cftalg{A}}}
\end{equation*}
was verified in \autoref{rmkenum:famfamempf-jdeq-fam} of \autoref{rmk:cftalg-defn}.
\item Now we verify the judgmental equality
\begin{equation*}
\jtermeq
  {{\Gamma}{\cftalgc{\cftalg{A}}}}
  {\cftalgf{\cftalg{A}}}
  {\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}}
  {\cftempf{\cftalg{A}}}.
\end{equation*}
Note that we can apply \autoref{lem:ehom-subst}, so we get the judgmental
equalities
\begin{align*}
\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}
& \jdeq
  \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftempf{\cftalg{A}}}
  \tag{by \autoref{lem:ehom-subst}}
  \\
& \jdeq
  \jcomp{}{\idtm{\cftalgf{\cftalg{A}}}}{\cftempf{\cftalg{A}}}
  \tag{by \autoref{cftalg-eq3}}
  \\
& \jdeq
  \cftempf{\cftalg{A}}.
  \tag{by \autoref{precomp-idtm-t}}
\end{align*}
\item The judgmental equalities
\begin{align*}
\subst{\cftempf{\cftalg{A}}}{\cftfamext}
& \jdeq 
  \idtm{\cftalgf{\cftalg{A}}}
  \\
\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}
& \jdeq 
  \idtm{\cftalgf{\cftalg{A}}}
\end{align*}
are given by assumption.
\item The judgmental equalities
\begin{align*}
\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
& \jdeq
  \idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  \\
\subst{\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
& \jdeq
  \idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
\end{align*}
follow from the judgmental equalities
\begin{align*}
\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}
& \jdeq
  \jcomp{}{\cftctxext}{\subst{\cftempf{\cftalg{A}}}{\cftfamext}}
  \tag{by \autoref{lem:jcomp-subst}}
  \\
& \jdeq
  \jcomp{}{\cftctxext}{\idtm{\cftalgf{\cftalg{A}}}}
  \tag{by \autoref{cftalg-eq3}}
  \\
& \jdeq
  \idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  \tag{by \autoref{comp-wi-t}}
\end{align*}
and
\begin{align*}
\subst
  {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}}
  {\jcomp{}{\cftctxext}{\cftfamext}}
& \jdeq
  \subst
    {\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}}
    {\jcomp{}{\cftctxext}{\cftfamext}}
  \tag{by \autoref{lem:extalg-twins}}
  \\
& \jdeq
  \jcomp
    {}
    {\cftctxext}
    {\subst{\jcomp{}{\cftctxext}{\cftempf{\cftalg{A}}}}{\cftfamext}}
  \tag{by \autoref{comp-sw-t}}
  \\
& \jdeq
  \jcomp
    {}
    {\cftctxext}
    {\idtm{\cftalgf{\cftalg{A}}}}
  \tag{by \autoref{cftalg-eq4}}
  \\
& \jdeq
  \idtm{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}.
  \tag{by \autoref{comp-wi-t}}
\end{align*}
\item Finally, the judgmental equality
\begin{equation*}
\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftfamext}}
  \jdeq
  \cftfamext
\end{equation*}
is verified as follows:
\begin{align*}
\subst{\cftempf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftfamext}}
& \jdeq
  \jcomp{}{\subst{\cftempf{\cftalg{A}}}{\cftctxext}}{\cftfamext}
  \tag{by \autoref{lem:ehom-subst}}
  \\
& \jdeq
  \jcomp{}{\idtm{\cftalgc{\cftalg{A}}}}{\cftfamext}
  \tag{by \autoref{cftalg-eq1}}
  \\
& \jdeq
  \cftfamext.
  \tag{by \autoref{precomp-idtm-t}}
\end{align*}
\end{enumerate}
\end{proof}

\begin{thm}\label{thm:cftalg-wk}
Let $\cftalg{Q}$ be an empty family algebra in context $\ctxext{\Gamma}{B}$ and let
$\jfam{\Gamma}{A}$ be a family of contexts. Then the septuple
\begin{equation*}
\ctxwk{A}{\cftalg{Q}}
  \defeq
  ( \ctxwk{A}{\cftalgc{\cftalg{Q}}},
    \ctxwk{A}{\cftalgf{\cftalg{Q}}},
    \ctxwk{A}{\cftalgt{\cftalg{Q}}},
    \ctxwk{A}{\cftctxext[\cftalg{Q}]},
    \ctxwk{A}{\cftfamext[\cftalg{Q}]},
    \ctxwk{A}{\cftempc{\cftalg{Q}}},
    \ctxwk{A}{\cftempf{\cftalg{Q}}})
\end{equation*}
is an empty family algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{thm}

\begin{proof}
The proof follows from the fact that weakening by $A$ is compatible with all
the involved operations.
\end{proof}

\begin{thm}\label{thm:cftalg-subst}
Let $\extalg{Q}$ be an empty family algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$ be a term. Then the septuple
\begin{equation*}
\subst{x}{\cftalg{Q}}
  \defeq
  ( \subst{x}{\cftalgc{\cftalg{Q}}},
    \subst{x}{\cftalgf{\cftalg{Q}}},
    \subst{x}{\cftalgt{\cftalg{Q}}},
    \subst{x}{\cftctxext[\cftalg{Q}]},
    \subst{x}{\cftfamext[\cftalg{Q}]},
    \subst{x}{\cftempc{\cftalg{Q}}},
    \subst{x}{\cftempf{\cftalg{Q}}})
\end{equation*}
is an empty family algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{thm}

\begin{proof}
The proof follows from the fact that substitution with $x$ is compatible with
all the involved operations.
\end{proof}

\begin{cor}
Let $\cftalg{Q}$ be an empty family algebra in context $\ctxext{\Gamma}{B}$
and let $\jhom{\Gamma}{A}{B}{f}$. Then the septuple
\begin{equation*}
\jcomp{A}{f}{\cftalg{Q}}
  \defeq
  ( \jcomp{A}{f}{\cftalgc{\cftalg{Q}}},
    \jcomp{A}{f}{\cftalgf{\cftalg{Q}}},
    \jcomp{A}{f}{\cftalgt{\cftalg{Q}}},
    \jcomp{A}{f}{\cftctxext[\cftalg{Q}]},
    \jcomp{A}{f}{\cftfamext[\cftalg{Q}]},
    \jcomp{A}{f}{\cftempc{\cftalg{Q}}},
    \jcomp{A}{f}{\cftempf{\cftalg{Q}}})
\end{equation*}
is an empty family algebra in context $\ctxext{\Gamma}{A}$.
\end{cor}

\begin{cor}
Let $\cftalg{A}$ be an empty family algebra in context $\Gamma$. Then
$\cftfamalg{\cftfamalg{\cftalg{A}}}$ and 
$\jcomp{}{\cftctxext[\cftalg{A}]}{\cftfamalg{\cftalg{A}}}$ are
judgmentally equal empty family algebras in context 
$\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$.
\end{cor}

\begin{lem}
Let $\cftalg{Q}$ be an empty family algebra in context $\ctxext{\Gamma}{B}$ and let
$\jfam{\Gamma}{A}$. Then we have the judgmental equality
\begin{equation*}
\ctxwk{A}{\cftfamalg{\cftalg{Q}}}\jdeq\cftalg{F}_{\ctxwk{A}{\cftalg{Q}}}
\end{equation*}
of empty family algebras in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{lem}

\begin{lem}
Let $\cftalg{Q}$ be an empty family algebra in context $\ctxext{{\Gamma}{A}}{P}$ 
and let $\jterm{\Gamma}{A}{x}$. Then we have the judgmental equality
\begin{equation*}
\subst{x}{\cftfamalg{\cftalg{Q}}}\jdeq\cftalg{F}_{\subst{x}{\cftalg{Q}}}
\end{equation*}
of empty family algebras in context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{lem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Empty family homomorphisms}
\begin{defn}
An empty family homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in context
$\Gamma$ is an extension homomorphism from $\unfold{\extalg{A}}$ to
$\unfold{\extalg{B}}$ in context $\Gamma$ for which we have the judgmental
equalities
\begin{align*}
\subst{\cftempc{\cftalg{A}}}{\cfthomc{\cfthom{f}}}
& \jdeq
  \cftempc{\cftalg{B}}
  \\
\subst{\cftempc{\cftalg{A}}}{\cfthomf{\cfthom{f}}}
& \jdeq
  \cfthomc{\cfthom{f}}
  \\
\subst{\cftempf{\cftalg{A}}}{\cfthomf{\cfthom{f}}}
& \jdeq
  \jcomp{}{\cfthomc{\cfthom{f}}}{\cftempf{\cftalg{B}}}
\end{align*}

Judgmental equality of empty family homomorphisms from $\cftalg{A}$ to $\cftalg{B}$ in
context $\Gamma$ is taken to be judgmental equality of extension homomorphisms
from $\extalg{A}$ to $\extalg{B}$ in context $\Gamma$.
\end{defn}

\begin{lem}
For any empty family algebra $\cftalg{A}$, the extension homomorphism $\cftidhom{\cftalg{A}}$
is an empty family homomorphism.
\end{lem}

\begin{lem}
Suppose that $\cfthom{f}$ is an empty family homomorphism from $\cftalg{A}$ to $\cftalg{B}$
in context $\Gamma$ and that $\cfthom{g}$ is an empty family homomorphism from $\cftalg{B}$
to $\cftalg{C}$ in context $\Gamma$. Then $\jcomp{}{\cfthom{f}}{\cfthom{g}}$
is an empty family homomorphism from $\cftalg{A}$ to $\cftalg{C}$ in context $\Gamma$.
\end{lem}

\begin{lem}
Let $\cfthom{f}$ be an empty family homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in
context $\Gamma$. Then $\cftfamhom{\exthom{f}}$ is an empty family homomorphism from
$\cftfamalg{\cftalg{A}}$ to $\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftalg{B}}}$
in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$. 
\end{lem}

\begin{lem}
Let $\cfthom{g}$ be an empty family homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
context $\ctxext{\Gamma}{B}$ and let $\jfam{\Gamma}{A}$. Then 
$\ctxwk{A}{\cfthom{g}}$
is an empty family homomorphism from $\ctxwk{A}{\cftalg{Q}}$ to $\ctxwk{A}{\cftalg{R}}$
in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{lem}

\begin{lem}
Let $\cfthom{g}$ be an empty family homomorphism from $\cftalg{Q}$ to $\cftalg{R}$ in
context $\ctxext{{\Gamma}{A}}{P}$ and let $\jterm{\Gamma}{A}{x}$. Then 
$\subst{x}{\cfthom{g}}$
is an empty family homomorphism from $\subst{x}{\cftalg{Q}}$ to $\subst{x}{\cftalg{R}}$
in context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{lem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Weakening algebras}
Weakening algebras will be extension algebras with certain added structure.
We shall first formulate the notion of a pre-weakening algrebra, which is an
extension algebra together with an extension homomorphism describing weakening,
but without the additional judgmental equalities which will have to hold for
weakening algebras. The advantage of this approach is that this allows us to
consider pre-weakening homomorphisms and require that the weakening of the
pre-weakening algebra is a pre-weakening homomorphism. This is the condition
of \autoref{comp-ww} that weakening is compatible with itself. We could also
formulate this condition directly, but it is complicated to state and it would
not be very insightful if we did that.

\begin{defn}
A pre-weakening algebra $\cftalg{A}$ in context $\Gamma$ is an extension algebra
$\cftalg{A}$ in context $\Gamma$ together with an extension homomorphism
\begin{equation}\label{cftwk}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
\end{equation}
called the \emph{weakening of $\cftalg{A}$}.
\end{defn}

Weakening algebras will be pre-weakening algebras for which the weakening
satisfies judgmental equalities which express abstracted analogues of
the rules in \autoref{comp-ww,comp-ew,comp-0w}. The
definition of weakening algebras can be found in \autoref{sec:cftwkalg-defn}.

We begin with some trivial remarks to familiarize ourselves with the situation
of pre-weakening algebras.
The extension homomorphism $\cftwk{\cftalg{A}}$ of \autoref{cftwk} is a triple
$(\cftwkc{\cftalg{A}},\cftwkf{\cftalg{A}},\cftwkt{\cftalg{A}})$ consisting of
the morphisms displayed in the following diagram:
\begin{equation*}
\begin{tikzcd}
\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}}}
  \ar[fib]{d}
  \ar{r}{\cftwkt{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftalgt{\cftalg{A}}}}
  \ar[fib]{d}
  \\
\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  \ar[fib]{d}
  \ar{r}{\cftwkf{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
  \ar[fib]{d}
  \\
\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{r}[swap]{\cftwkc{\cftalg{A}}}
& \jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}
\end{tikzcd}
\end{equation*}
In particular, we see that $\cftwk{\cftalg{A}}$ is an extension homomorphism from
$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to $\cftfamalg{\cftalg{A}}$ over
$\cftctxext$ in context $\Gamma$.

\begin{rmk}
Let $\cftalg{A}$ be an extension algebra and let $\cftwk{\cftalg{A}}$ be as in
\autoref{cftwk}. Then $\apply\cftwk{\cftalg{A}}{\gamma}{a}$ is an extension homomorphism
from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
$\subst{a}{{\gamma}{\cftfamalg{\cftfamalg{\cftalg{A}}}}}$. Furthermore, we have
\begin{equation*}
\jterm
  {\Gamma}
  {\subst{\apply\cftctxext{\gamma}{a}}{\cftalgf{\cftalg{A}}}}
  {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
\end{equation*}
for $\gamma:A$ and $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$; we have
\begin{equation*}
\jterm
  { \Gamma}
  { \subst
      { \apply\cftctxext
          {\apply\cftctxext{\gamma}{a}}
          {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
        }
      { \cftalgf{\cftalg{A}}}}
  { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}}
\end{equation*}
for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
$q:\subst{\apply\cftctxext{\gamma}{b}}{\cftalgf{\cftalg{A}}}$; and finally, we have
\begin{equation*}
\jterm
  { \Gamma}
  { \subst
      { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}}
      { { \apply\cftctxext
            {\apply\cftctxext{\gamma}{a}}
            {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
          }
        {\cftalgt{\cftalg{A}}}
        }
    }
  {\subst{g}{{q}{{b}{{a}{{\gamma}{\cftwkt{\cftalg{A}}}}}}}}
\end{equation*}
for $\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$,
$q:\subst{\apply\cftctxext{\gamma}{b}}{\cftalgf{\cftalg{A}}}$ and
$g:\subst{q}{{\apply\cftctxext{\gamma}{b}}{\cftalgt{\cftalg{A}}}}$.
These observations show that the weakening morphism $\cftwk{\cftalg{A}}$ is
indeed an abstraction of the introduction rules for the weakening operation.
\end{rmk}

\begin{defn}
Suppose $\cftalg{A}$ is a pre-weakening algebra in context $\Gamma$. Then we
define the pre-weakening algebra $\cftfamalg{\cftalg{A}}$ in context 
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$ by taking
\begin{equation*}
\cftwk{\cftfamalg{\cftalg{A}}}
  \defeq \jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{Q}$ is a pre-weakening algebra in context $\ctxext{\Gamma}{B}$ and
let $\jfam{\Gamma}{A}$. Then we define the pre-weakening algebra 
$\ctxwk{A}{\cftalg{Q}}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ by 
taking
\begin{equation*}
\cftwk{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftwk{\cftalg{Q}}.
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{Q}$ is a pre-weakening algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$. Then we define the pre-weakening algebra 
$\subst{x}{\cftalg{Q}}$ in context $\ctxext{\Gamma}{\subst{x}{P}}$ by taking
\begin{equation*}
\cftwk{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftwk{\cftalg{Q}}}.
\end{equation*}
\end{defn}

\begin{lem}
Suppose $\cftalg{A}$ is a pre-weakening algebra in context $\Gamma$. Then
$\jcomp{}{\cftctxext}{\cftfamalg{\cftalg{A}}}$ and $\cftfamalg{\cftfamalg{\cftalg{A}}}$
are judgmentally equal pre-weakening algebras.
\end{lem}

\begin{proof}
This follows immediately from the facts that
$\jcomp{}{\cftctxext}{\cftfamalg{\cftalg{A}}}$ and $\cftfamalg{\cftfamalg{\cftalg{A}}}$
are judgmentally equal extension algebras in context $\Gamma$ and that
\begin{equation*}
\jcomp{}{\cftctxext}{\cftwk{\cftfamalg{\cftalg{A}}}}
  \jdeq
\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \jdeq
\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \jdeq
\cftwk{\cftfamalg{\cftfamalg{\cftalg{A}}}}.
  \qedhere
\end{equation*}
\end{proof}

\begin{lem}
Suppose $\cftalg{Q}$ is a pre-weakening algebra in context $\ctxext{\Gamma}{B}$
and that $\jfam{\Gamma}{A}$. Then
\end{lem}

\subsubsection{Pre-weakening homomorphisms}
It is useful to start with the definition of pre-weakening homomorphisms. Those
are extension homomorphisms which also preserve weakening. Before we give the
definition, recall from \autoref{lem:jvcomp-cprojfstf} that we have a
commutative square
\begin{equation*}
\begin{tikzcd}
\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{r}{\jvcomp{}{\cfthomc{\cfthom{f}}}{\cfthomf{\cfthom{f}}}}
  \ar{d}[swap]{\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  &
\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{B}}}
  \ar{d}{\cprojfstf{\cftalgc{\cftalg{B}}}{\cftalgf{\cftalg{B}}}}
  \\
\cftalgc{\cftalg{A}}
  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
  &
\cftalgc{\cftalg{B}}
\end{tikzcd}
\end{equation*}
of morphisms in context $\Gamma$ for every extension homomorphism $\cfthom{f}$ from
$\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$. Because weakening is the
same thing as pulling back along projection, we therefore see that the
judgmental equality
\begin{equation*}
\jfameq
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  { \ctxwk
      {\cftalgf{\cftalg{A}}}
      {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftalg{B}}}}
    }
  { \jcomp{}
      { \cfthomf{\cfthom{f}}}
      { \jcomp{}
        {\cfthomc{\cfthom{f}}}
        {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftalg{B}}}}
        }
    }
\end{equation*}
is valid for every extension homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
in context $\Gamma$. This judgmental equality helps us seeing that the condition
on pre-weakening homomorphisms is well-typed.

\begin{defn}
Let $\cftalg{A}$ and $\cftalg{B}$ be pre-weakening algebras in context $\Gamma$.
A \emph{pre-weakening homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context
$\Gamma$} is an extension homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$
in context $\Gamma$ for which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}
  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamhom{\cfthom{f}}}}
  \ar{d}[swap]{\cftwk{\cftalg{A}}}
  &
\jcomp{}
  { \cfthomf{\cfthom{f}}}
  { \jcomp{}
    {\cfthomc{\cfthom{f}}}
    {\ctxwk{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftalg{B}}}}
    }
  \ar{d}
    { \jcomp{}
        {\cfthomf{\cfthom{f}}}
        {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftwk{\cftalg{B}}}}
      }
  \\
\cftfamalg{\cftfamalg{\cftalg{A}}}
  \ar{r}[swap]{\cftfamhom{\cftfamhom{\cfthom{f}}}}
  &
\jcomp{}
  {\cfthomf{\cfthom{f}}}
  {\jcomp{}{\cfthomc{\cfthom{f}}}{\cftfamalg{\cftfamalg{\cftalg{B}}}}}
\end{tikzcd}
\end{equation*}
of extension algebras in context 
$\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$ commutes
judgmentally.
\end{defn}

\begin{rmk}\label{rmk:cftwk-prewk}
In this remark we explain what it means that weakening is a pre-weakening
homomorphism. Let $\cftalg{A}$ be a pre-weakening algebra. Then its weakening
$\cftwk{\cftalg{A}}$ is an extension homomorphism from
$\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}$ to
$\cftfamalg{\cftfamalg{\cftalg{A}}}$. Therefore, $\cftwk{\cftalg{A}}$ is a
pre-weakening homomorphism if the diagram
\begin{equation*}
\begin{tikzcd}[column sep=8em]
\ctxwk
  {\cftalgf{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  {\cftfamalg{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  \ar{r}
    { \ctxwk
        {\cftalgf{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
  \ar{d}[swap]{\cftwk{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  &
\jcomp{}
  { \cftwkf{\cftalg{A}}}
  { \jcomp{}
    {\cftwkc{\cftalg{A}}}
    {\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
    }
  \ar{d}
    { \jcomp{}
        { \cftwkf{\cftalg{A}}}
        { \jcomp{}
            { \cftwkc{\cftalg{A}}}
            { \cftwk{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
      }
  \\
\cftfamalg{\cftfamalg{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  \ar{r}[swap]
    { \cftfamhom
        { \cftfamhom
            { \cftwk{\cftalg{A}}}
          }
      }
  &
\jcomp{}
  { \cftwkf{\cftalg{A}}}
  { \jcomp{}
      {\cftwkc{\cftalg{A}}}
      {\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
    }
\end{tikzcd}
\end{equation*}
of pre-weakening algebras in context
$ \ctxext
    { { { { \Gamma}
          { \cftalgc{\cftalg{A}}}
          }
        { \cftalgf{\cftalg{A}}}
        }
      { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
      }
    { \cftalgf{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
      }$
commutes judgmentally.
\end{rmk}

\begin{comment}
\subsubsection{Currying for weakening}
Let $\cftalg{A}$ be an extension algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be an extension homomorphism. The abstraction of the rules in \autoref{comp-ew} is the
condition that the diagram
\begin{equation}\label{diag:cftwk-cftextcurry}
\begin{tikzcd}[column sep=large]
\ctxwk
  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
  {\cftfamalg{\cftalg{A}}}
  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \\
& \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}
\end{tikzcd}
\end{equation}
of extension algebras in context 
$\ctxext
    {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
    {\cftalgf{\cftfamalg{\cftalg{A}}}}
$ %
commutes.

\begin{lem}
Let $\cftalg{A}$ be an extension algebra in context gamma with an extension homomorphism 
$ \jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\jcomp{}{\cftctxext}{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
$
satisfying the judgmental equality of \autoref{diag:cftwk-cftextcurry}. Then we
have the judgmental equality
\begin{equation*}
\apply\cftwk{\cftalg{A}}{\gamma}{\apply\cftfamext{\gamma}{a}{p}}
\jdeq
\jcomp{}
    {\apply\cftwk{\cftalg{A}}{\gamma}{a}}
    {\apply\cftwk{\cftalg{A}}{\apply\cftctxext{\gamma}{a}}{p}}
\end{equation*}
of extension homomorphisms from $\subst{\gamma}{\cftfamalg{\cftalg{A}}}$ to 
$\subst{p}{{a}{{\gamma}{\cftalg{F_{F_{F_A}}}}}}$ in context $\Gamma$,
for every $\gamma:A$, $a:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
$p:\subst{a}{{\gamma}{\cftalgf{\cftfamalg{\cftalg{A}}}}}$. 
\end{lem}

\subsubsection{Weakening by the empty family}
\begin{equation}\label{eq:cftwk-cftempcurry}
\jhomeq
  { {\Gamma}{\cftalgc{\cftalg{A}}}
    }
  { \cftfamalg{\cftalg{A}}
    }
  { \cftfamalg{\cftalg{A}}
    }
  { \subst{\cftempf{\cftalg{A}}}{\cftwk{\cftalg{A}}}
    }
  { \cftidhom{\cftfamalg{\cftalg{A}}}
    }
\end{equation}

\subsubsection{Weakening preserves itself}
Let $\cftalg{A}$ be an extension algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be an extension homomorphism. The abstraction of the rules in \autoref{comp-ww} is the
condition that the diagram
\begin{equation}\label{diag:cftwk-cftwk}
\begin{tikzcd}[column sep=10em]
\ctxwk{\cftalgf{\cftalg{A}}}{{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
  \ar{d}
    [swap]{ \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
  \ar{r}{\ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
& \ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\cftfamhom{\cftwk{\cftalg{A}}}}}
  \\
\ctxwk{{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}{\jcomp{}{\cftwkc{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  \ar{r}[swap]{%
    \jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cftwk{\cftalg{A}}}
                }
            }
        }
    }
& \jcomp{}
    {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
    {\jcomp{}{\cftwkc{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
\end{tikzcd}
\end{equation}
of extension algebras in context 
$ \ctxext
    { { { {\Gamma}
          {\cftalgc{\cftalg{A}}}
          }
        { \cftalgf{\cftalg{A}}}
        }
      { \ctxwk
          {\cftalgf{\cftalg{A}}}
          {\cftalgf{\cftalg{A}}}
        }
      }
    { \ctxwk
        {\cftalgf{\cftalg{A}}}
        {\cftalgf{\cftfamalg{\cftalg{A}}}}
      }$ %
commutes. In this diagram, it is not immediately obvious that\ldots
\begin{enumerate}
\item \ldots the domain of the extension homomorphism
$\jcomp{}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}{\cftfamhom{\cftwk{\cftalg{A}}}}$
is the indicated domain. Note that we have the judgmental equalities
\begin{align*}
\jcomp{}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\jcomp{}{\cftfamext}{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
  \\
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
  \\
& \jdeq
\ctxwk
  {\cftalgf{\cftalg{A}}}
  {\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}.
\end{align*}
\item \ldots the domain and codomain of the extension homomorphism
\begin{equation*}
\jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cftwk{\cftalg{A}}}
                }
            }
        }
\end{equation*}
from
$ \jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
                }
            }
        }
$ to
$ \jcomp{}
      { \cftwkf{\cftalg{A}}}
      { \jcomp{}
          {\cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
                }
            }
        }
$ are indeed judgmentally equal to the indicated domain and codomain.

To see
that the two domains are equal, note that $\cftwkf{\cftalg{A}}$ may be seen as
a morphism from $\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}$
to $\cftalgf{\cftalg{A}}$ over 
$\jcomp{}{\cftwkc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftctxext}}$ in context
$\Gamma$, i.e. we have the following commuting diagram in context $\Gamma$:
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}
  \ar{r}
    {\cftwkf{\cftalg{A}}}
  \ar[fib]{d}
  &
\cftalgf{\cftalg{A}}
  \ar[fib]{d}
  \\
\ctxext{ {\cftalgc{\cftalg{A}}}
         {\cftalgf{\cftalg{A}}}
         }
       { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
         }
  \ar{r}[swap]
    {\jcomp{}{\cftwkc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftctxext}}}
  &
\cftalgc{\cftalg{A}}
\end{tikzcd}
\end{equation*}
Using the associativity of composition, we see that we have the judgmental 
equality
\begin{equation*}
\jcomp{}
  { \cftwkf{\cftalg{A}}}
  { \jcomp{}
      {\cftwkc{\cftalg{A}}}
      { \jcomp{}
          {\cftctxext}
          { \jcomp{}
              {\cftctxext}
              {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
            }
        }
    }
\jdeq
\jcomp{}
  { \jvcomp{}
      { \jcomp{}
          { \cftwkc{\cftalg{A}}}
          { \jcomp{}
              {\cftctxext}
              {\cftctxext}
            }
        }
      { \cftwkf{\cftalg{A}}
        }
    }
  { \ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}  
\end{equation*}
Recall that weakening is the same as pulling
back along a projection. Therefore, we see that
it suffices to show that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=10em]
\ctxext{ { { \cftalgc{\cftalg{A}}}
           { \cftalgf{\cftalg{A}}}
           }
         { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
         }
       { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}
         }
  \ar{r}
    { \jvcomp{}
        { \jcomp{}
            {\cftwkc{\cftalg{A}}}
            { \jcomp{}
                {\cftctxext}
                {\cftctxext}
              }
          }
        { \cftwkf{\cftalg{A}}}
      }
  \ar{d}[swap]
    { \cprojfstf
        { { { \cftalgc{\cftalg{A}}}
            { \cftalgf{\cftalg{A}}}
            }
          { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
          }
        { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}
          }
      }
  &
\ctxext{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
  \ar{d}
    {\cprojfstf{\cftalgc{\cftalg{A}}}{\cftalgf{\cftalg{A}}}}
  \\
\ctxext{ {\cftalgc{\cftalg{A}}}
         {\cftalgf{\cftalg{A}}}
         }
       { \ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftalg{A}}}
         }
  \ar{r}[swap]
    {\jcomp{}{\cftwkc{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftctxext}}}
  &
\cftalgc{\cftalg{A}}
\end{tikzcd}
\end{equation*}
commutes judgmentally. This follows from \autoref{lem:jvcomp-cprojfstf}.

To see that the codomains are equal, use the fact that $\cftwk{\cftalg{A}}$ is
an extension homomorphism.
\end{enumerate}

\begin{lem}
Let $\cftalg{A}$ be an extension algebra in context $\Gamma$ and let
\begin{equation*}
\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
\end{equation*}
be a homomorphism of extension algebras for which the diagram in
\autoref{diag:cftwk-cftwk} commutes judgmentally. 
Then the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\subst
  {\apply\cftctxext{\gamma}{b}}
  {\cftfamalg{\cftalg{A}}}
  \ar{r}{\apply\cftwk{\cftalg{A}}{\apply\cftctxext{\gamma}{b}}{q}}
  \ar{d}[swap]{\subst{b}{{a}{{\gamma}{\cftfamhom{\cftwk{\cftalg{A}}}}}}}
& \subst
    { q}
    { {\apply\cftctxext{\gamma}{b}}
      {\cftfamalg{\cftfamalg{\cftalg{A}}}}
      }
  \ar{d}
    { \subst
        {\apply\cftfamext{\gamma}{b}{q}}
        {{a}{{\gamma}{\cftfamhom{\cftwk{\cftalg{A}}}}}}
      }
  \\
\subst
  { \apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}
    }
  { { \apply\cftctxext{\gamma}{a}}
    { \cftfamalg{\cftfamalg{\cftalg{A}}}}
    }
  \ar{r}[swap,yshift=-1em]
    { \subst
        { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}}
        { {\apply\cftctxext{\apply\cftctxext{\gamma}{a}}{\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}}
          {\cftwk{\cftalg{A}}}
          }
      }
& \subst
    { \apply\cftwkf{\cftalg{A}}{\gamma}{a}{b}{q}
      }
    { {\apply\cftwkc{\cftalg{A}}{\gamma}{a}{b}}
      {{a}{{\gamma}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}}}
      }
\end{tikzcd}
\end{equation*}
of extension algebras in context $\Gamma$ commutes judgmentally for
$\gamma:A$, $a,b:\subst{\gamma}{\cftalgf{\cftalg{A}}}$ and
$q:\subst{b}{{\gamma}{\cftalgf{\cftfamalg{\cftalg{A}}}}}$.
\end{lem}

\begin{proof}
All one has to do to prove this is to substitute all of the ingredients of
the diagram in \autoref{diag:cftwk-cftwk} consecutively by
$\gamma$, by $a$, by $b$ and finally by $q$.
\end{proof}
\end{comment}

\subsubsection{The definition of weakening algebras}\label{sec:cftwkalg-defn}

\begin{defn}
A weakening algebra in context $\Gamma$ is a pre-weakening algebra $\cftalg{A}$
in context $\Gamma$ for which the weakening $\cftwk{\cftalg{A}}$ is a
pre-weakening homomorphism as explained in \autoref{rmk:cftwk-prewk},
satisfying the two additional judgmental equalities
expressing that: (1) the diagram
\begin{equation}\label{diag:cftwk-cftextcurry}
\begin{tikzcd}[column sep=large]
\ctxwk
  {\ctxext{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
  {\cftfamalg{\cftalg{A}}}
  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
  \ar{dr}[swap]{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
  \ar{d}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \\
& \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}
\end{tikzcd}
\end{equation}
of pre-weakening algebras in context $\ctxext{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}$ commutes judgmentally and (2) that
\begin{equation}\label{eq:cftwk-cftempcurry}
\jhomeq
  { {\Gamma}{\cftalgc{\cftalg{A}}}
    }
  { \cftfamalg{\cftalg{A}}
    }
  { \cftfamalg{\cftalg{A}}
    }
  { \subst{\cftempf{\cftalg{A}}}{\cftwk{\cftalg{A}}}
    }
  { \cftidhom{\cftfamalg{\cftalg{A}}}
    }
\end{equation}
\end{defn}

\begin{rmk}
The condition in \autoref{diag:cftwk-cftextcurry} implements the rules of
\autoref{comp-ew} and the condition in \autoref{eq:cftwk-cftempcurry} implements
the rules of \autoref{comp-0w}.
\end{rmk}

\begin{defn}
A \emph{weakening homomorphism} is a pre-weakening homomorphism between
weakening algebras.
\end{defn}

\subsubsection{Derivable properties of weakening algebras}

\begin{thm}
Suppose $\cftalg{A}$ is a weakening algebra in context $\Gamma$. Then
$\cftfamalg{\cftalg{A}}$ together with the extension homomorphism
\begin{equation*}
\cftwk{\cftfamalg{\cftalg{A}}}
  \defeq \jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}
\end{equation*}
forms a weakening algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
\end{thm}

\begin{proof}
To show that $\cftwk{\cftfamalg{\cftalg{A}}}$ satisfies the Currying rule, we
must verify that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxwk
  {\ctxext{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftwk{\cftfamalg{\cftalg{A}}}}}
  \ar{dr}[swap]{\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
  \ar{d}{\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
  \\
& \cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
\end{tikzcd}
\end{equation*}
commutes judgmentally. We do this by showing that each of the ingredients of this
diagram is judgmentally equal to the corresponding ingredient of the diagram in
\autoref{diag:cftwk-cftextcurry} pulled back along $\cftctxext$. Thus, we are going
to show that
\begin{align*}
\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
  \tag{1}
  \\
\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \tag{2}
  \\
\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
\jcomp{}{\cftctxext}{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
  \tag{3}
\end{align*}
For the first, we have the judgmental equalities
\begin{align*}
\ctxwk{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
\ctxwk
  {\jcomp{}{\cftctxext}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \\
& \jdeq
\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftwk{\cftalg{A}}}}
\end{align*}
For the second, we have the judgmental equalities
\begin{align*}
\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
  \jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \\
& \jdeq
  \jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
\end{align*}
For the third, note that we have the judgmental equalities
\begin{align*}
\jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\cftwk{\cftfamalg{\cftalg{A}}}}
& \jdeq
  \jcomp{}{\jcomp{}{\cftctxext}{\cftfamext}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
  \\
& \jdeq
\jcomp{}{\cftctxext}{\jcomp{}{\cftfamext}{\cftwk{\cftalg{A}}}}
\end{align*}
Thus we may conclude that the diagram above a pullback of the diagram in 
\autoref{diag:cftwk-cftextcurry} by $\cftctxext$ and therefore it commutes as desired.

To show that $\cftwk{\cftfamalg{\cftalg{A}}}$ preserves itself, we must verify
that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  \ar{d}
    [swap]{ \ctxwk
        {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
        {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
      }
  \ar{r}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}}
& \ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  \ar{d}{\jcomp{}{\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}{\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}}
  \\
\ctxwk{{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}{\jcomp{}{\cftwkc{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}}
  \ar{r}[swap,yshift=-1em]{%
    \jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          {\cftwkc{\cftfamalg{\cftalg{A}}}}
          { \jcomp{}
              {\cftfamext}
              { \jcomp{}
                  {\cftfamext}
                  {\cftwk{\cftfamalg{\cftalg{A}}}}
                }
            }
        }
    }
& \jcomp{}
    {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
    {\jcomp{}{\cftwkc{\cftfamalg{\cftalg{A}}}}{\cftfamalg{\cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}}
\end{tikzcd}
\end{equation*}
commutes judgmentally. We do this by showing that each of the ingredients of this
diagram is judgmentally equal to the corresponding ingredient of the diagram in
\autoref{diag:cftwk-cftwk} pulled back along $\cftctxext$. Thus, we are going
to show that
\begin{align*}
\ctxwk
  {\cftalgf{\cftfamalg{\cftalg{A}}}}
  {\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \tag{1}
  \\
\jcomp{}
  {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \jcomp{}
        {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
  \tag{2}
  \\
\ctxwk
  {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
  \tag{3}
  \\
\jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          {\cftwkc{\cftfamalg{\cftalg{A}}}}
          { \jcomp{}
              {\cftfamext}
              { \jcomp{}
                  {\cftfamext}
                  {\cftwk{\cftfamalg{\cftalg{A}}}}
                }
            }
        }
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \jcomp{}
        { \cftwkf{\cftalg{A}}}
        { \jcomp{}
            {\cftwkc{\cftalg{A}}}
            { \jcomp{}
                {\cftctxext}
                { \jcomp{}
                    {\cftctxext}
                    {\cftwk{\cftalg{A}}}
                  }
              }
          }
      }
  \tag{4}
\end{align*}
For the first, note that we have the judgmental equalities
\begin{align*}
\ctxwk
  {\cftalgf{\cftfamalg{\cftalg{A}}}}
  {\jcomp{}{\cftfamext}{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq 
  \ctxwk
    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
    {\jcomp{}{\cftfamext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \ctxwk
    {\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}
    {\jcomp{}{\cftctxext}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \ctxwk{\cftalgf{\cftalg{A}}}{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
\end{align*}
For the second, note that we have the judgmental equalities
\begin{align*}
\jcomp{}
  {\ctxwk{\cftalgf{\cftfamalg{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \jcomp{}
    {\ctxwk{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftfamext}}}
    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}}
    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}}
    {\jcomp{}{\cftctxext}{\cftfamhom{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \jcomp{}
        {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamext}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
\end{align*}
For the third, note that we have the judgmental equalities
\begin{align*}
\ctxwk
  {{\cftalgf{\cftfamalg{\cftalg{A}}}}{\cftalgf{\cftfamalg{\cftfamalg{\cftalg{A}}}}}}
  {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
& \jdeq
  \ctxwk
    {{\jcomp{}{\cftctxext}{\cftalgf{\cftalg{A}}}}{\jcomp{}{\cftctxext}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
    {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
  \\
& \jdeq
  \ctxwk
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
    {\cftfamhom{\cftwk{\cftfamalg{\cftalg{A}}}}}
  \\
& \jdeq
  \ctxwk
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
    {\cftfamhom{\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \ctxwk
    {\jcomp{}{\cftctxext}{\ctxwk{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}}
    {\jcomp{}{\cftctxext}{\cftfamhom{\cftwk{\cftalg{A}}}}}
  \\
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \ctxwk
        {{\cftalgf{\cftalg{A}}}{\cftalgf{\cftfamalg{\cftalg{A}}}}}
        {\cftfamhom{\cftwk{\cftalg{A}}}}
      }
\end{align*}
For the fourth, note that we have the judgmental equalities
\begin{align*}
\jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          {\cftwkc{\cftfamalg{\cftalg{A}}}}
          { \jcomp{}
              {\cftfamext}
              { \jcomp{}
                  {\cftfamext}
                  {\cftwk{\cftfamalg{\cftalg{A}}}}
                }
            }
        }
& \jdeq
  \jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          { \jcomp{}{\cftctxext}{\cftwkc{\cftalg{A}}}}
          { \jcomp{}
              {\cftfamext}
              { \jcomp{}
                  {\cftfamext}
                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
                }
            }
        }
  \\
& \jdeq
  \jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          { \jcomp{}{\cftctxext}{\cftwkc{\cftalg{A}}}}
          { \jcomp{}
              {\cftctxext}
              { \jcomp{}
                  {\cftctxext}
                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
                }
            }
        }
  \\
& \jdeq
  \jcomp{}
      { \cftwkf{\cftfamalg{\cftalg{A}}}}
      { \jcomp{}
          { \cftctxext}
          { \jcomp{}
              { \cftwkc{\cftalg{A}}}
              { \jcomp{}
                  {\cftctxext}
                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
                }
            }
        }
  \\
& \jdeq
  \jcomp{}
      { \jcomp{}{\cftctxext}{\cftwkf{\cftalg{A}}}}
      { \jcomp{}
          { \cftctxext}
          { \jcomp{}
              { \cftwkc{\cftalg{A}}}
              { \jcomp{}
                  {\cftctxext}
                  {\jcomp{}{\cftctxext}{\cftwk{\cftalg{A}}}}
                }
            }
        }
  \\
& \jdeq
  \jcomp{}
    { \cftctxext}
    { \jcomp{}
        { \cftwkf{\cftalg{A}}}
        { \jcomp{}
            {\cftwkc{\cftalg{A}}}
            { \jcomp{}
                {\cftctxext}
                { \jcomp{}
                    {\cftctxext}
                    {\cftwk{\cftalg{A}}}
                  }
              }
          }
      }
\end{align*}
Thus we may conclude that the diagram above a pullback of the diagram in \autoref{diag:cftwk-cftwk}
by $\cftctxext$ and therefore it commutes as desired.
\end{proof}

\begin{thm}
Suppose $\cftalg{Q}$ is a weakening algebra in context $\ctxext{\Gamma}{B}$ and
let $\jfam{\Gamma}{A}$. Then $\ctxwk{A}{\cftalg{Q}}$ together with the
extension homomorphism
\begin{equation*}
\cftwk{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftwk{\cftalg{Q}}
\end{equation*}
forms a weakening algebra.
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is a weakening algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$. Then $\subst{x}{\cftalg{Q}}$ together with the
extension homomorphism
\begin{equation*}
\cftwk{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftwk{\cftalg{Q}}}
\end{equation*}
forms a weakening algebra.
\end{thm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Projection algebras}
Projection algebras will be weakening algebras with additional terms implementing
the identity terms. We call these algebras projection algebras because the
identity terms can only be formulated in weakening algebras and together with
weakening, the identity terms provide for all the projections.

\begin{defn}
A \emph{pre-projection algebra $\cftalg{A}$ in context $\Gamma$} consists of a 
weakening algebra $\cftalg{A}$ in context $\Gamma$ together with an 
\emph{internal identity term}
\begin{equation*}
\jterm
  { { {\Gamma}
      {\cftalgc{\cftalg{A}}}
      }
    { \cftalgf{\cftalg{A}}
      }
    }
  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftfamalg{\cftalg{A}}}}}
  {\cftidtm{\cftalg{A}}}.
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{A}$ is a pre-projection algebra in context $\Gamma$. Then we
define the pre-projection algebra $\cftfamalg{\cftalg{A}}$ in context 
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$ by taking
\begin{equation*}
\cftidtm{\cftfamalg{\cftalg{A}}}
  \defeq \jcomp{}{\cftctxext}{\cftidtm{\cftalg{A}}}.
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{Q}$ is a pre-projection algebra in context $\ctxext{\Gamma}{B}$ and
let $\jfam{\Gamma}{A}$. Then we define the pre-projection algebra 
$\ctxwk{A}{\cftalg{Q}}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ by 
taking
\begin{equation*}
\cftidtm{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftidtm{\cftalg{Q}}.
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{Q}$ is a pre-projection algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$. Then we define the pre-projection algebra 
$\subst{x}{\cftalg{Q}}$ in context $\ctxext{\Gamma}{\subst{x}{P}}$ by taking
\begin{equation*}
\cftidtm{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftidtm{\cftalg{Q}}}.
\end{equation*}
\end{defn}

\begin{defn}
A pre-projection homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context
$\Gamma$ is a weakening homomorphism $\cfthom{f}$ from $\cftalg{A}$ to
$\cftalg{B}$ in context $\Gamma$ for which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=10em]
\subst
  {{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}
  {\cftalgt{\cftfamalg{\cftalg{A}}}}
  \ar{r}{\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cfthomt{\cftfamhom{\cfthom{f}}}}}
  \ar[fib,xshift=-.7ex]{d}
  &
\subst
  {{\idtm{\cftalgf{\cftalg{B}}}}{\cftwkc{\cftalg{B}}}}
  {\cftalgt{\cftfamalg{\cftalg{B}}}}
  \ar[fib,xshift=-.7ex]{d}
  \\
\cftalgf{\cftalg{A}}
  \ar{r}[swap]{\cfthomf{\cfthom{f}}}
  \ar[fib]{d}
  \ar[densely dotted,xshift=.7ex]{u}[swap]{\cftidtm{\cftalg{A}}}
  &
\cftalgf{\cftalg{B}}
  \ar[fib]{d}
  \ar[densely dotted,xshift=.7ex]{u}[swap]{\cftidtm{\cftalg{B}}}
  \\
\cftalgc{\cftalg{A}}
  \ar{r}[swap]{\cfthomc{\cfthom{f}}}
  &
\cftalgc{\cftalg{B}}
\end{tikzcd}
\end{equation*}
\end{defn}

\begin{defn}
A projection algebra in context $\Gamma$ is a pre-projection algebra in context
$\Gamma$ for which weakening is a pre-projection homomorphism.
\end{defn}

\begin{defn}
A projection homomorphism is a pre-projection homomorphism between projection
algebras.
\end{defn}

\begin{thm}
Suppose $\cftalg{A}$ is a projection algebra in context $\Gamma$. Then the 
pre-projection algebra $\cftfamalg{\cftalg{A}}$ is a projection algebra in context 
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is a projection algebra in context $\ctxext{\Gamma}{B}$ and
let $\jfam{\Gamma}{A}$. Then the pre-projection algebra 
$\ctxwk{A}{\cftalg{Q}}$ is a
projection algebra in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is a projection algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$. Then the pre-projection algebra 
$\subst{x}{\cftalg{Q}}$ is a
projection algebra in context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{thm}

\begin{cor}
In a projection algebra $\cftalg{A}$ in context $\Gamma$, weakening is a
projection homomorphism.
\end{cor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Substitution algebras}
\begin{defn}
A \emph{pre-substitution algebra $\cftalg{A}$ in context $\Gamma$} consists of an extension
algebra $\cftalg{A}$ in context $\Gamma$ and an extension homomorphism
\begin{equation*}
\jhom
  {{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}}
  {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
  {\ctxwk{\cftalgt{\cftalg{A}}}{{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  {\cftsubst{\cftalg{A}}}.
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{A}$ is a pre-substitution algebra in context $\Gamma$. Then we
define the pre-substitution algebra $\cftfamalg{\cftalg{A}}$ in context 
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$ by taking
\begin{equation*}
\cftsubst{\cftfamalg{\cftalg{A}}}
  \defeq \jcomp{}{\cftctxext}{\cftsubst{\cftalg{A}}}.
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{Q}$ is a pre-substitution algebra in context $\ctxext{\Gamma}{B}$ and
let $\jfam{\Gamma}{A}$. Then we define the pre-substitution algebra 
$\ctxwk{A}{\cftalg{Q}}$ in context $\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$ by 
taking
\begin{equation*}
\cftsubst{\ctxwk{A}{\cftalg{Q}}}\defeq \ctxwk{A}\cftsubst{\cftalg{Q}}.
\end{equation*}
\end{defn}

\begin{defn}
Suppose $\cftalg{Q}$ is a pre-substitution algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$. Then we define the pre-substitution algebra 
$\subst{x}{\cftalg{Q}}$ in context $\ctxext{\Gamma}{\subst{x}{P}}$ by taking
\begin{equation*}
\cftsubst{\subst{x}{\cftalg{Q}}} \defeq \subst{x}{\cftsubst{\cftalg{Q}}}.
\end{equation*}
\end{defn}

\begin{defn}
A \emph{pre-substitution homomorphism $\cfthom{f}$ from $\cftalg{A}$ to $\cftalg{B}$ in
context $\Gamma$} is an extension homomorphism $\cfthom{f}$ for which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=8em]
\ctxwk{\cftalgt{\cftalg{A}}}{\cftfamalg{\cftfamalg{\cftalg{A}}}}
  \ar{r}{\ctxwk{\cftalgt{\cftalg{A}}}{\cftfamhom{\cftfamhom{\cfthom{f}}}}}
  \ar{d}[swap]{\cftsubst{\cftalg{A}}}
  &
\jcomp{}
  { \cfthomt{\cfthom{f}}}
  { \jcomp{}
      { \cfthomf{\cfthom{f}}}
      { \jcomp{}
          { \cfthomc{\cfthom{f}}}
          { \ctxwk{\cftalgt{\cftalg{B}}}{\cftfamalg{\cftfamalg{\cftalg{B}}}}}
        }
    }
  \ar{d}
    { \jcomp{}
        { \cfthomt{\cfthom{f}}}
        { \jcomp{}
            { \cfthomf{\cfthom{f}}}
            { \jcomp{}
                { \cfthomc{\cfthom{f}}}
                {\cftsubst{\cftalg{B}}}
              }
          }
      }   
  \\
\ctxwk{\cftalgt{\cftalg{A}}}{{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  \ar{r}[swap]
    {\ctxwk{\cftalgt{\cftalg{A}}}{{\cftalgf{\cftalg{A}}}{\cftfamhom{\cfthom{f}}}}}
  &
\jcomp{}
  { \cfthomt{\cfthom{f}}}
  { \jcomp{}
      { \cfthomf{\cfthom{f}}}
      { \jcomp{}
          { \cfthomc{\cfthom{f}}}
          { \ctxwk
              {\cftalgt{\cftalg{B}}}
              {{\cftalgf{\cftalg{B}}}{\cftfamalg{\cftalg{B}}}}
            }
        }
    }
\end{tikzcd}
\end{equation*}
of pre-substitution algebras in context
$\ctxext{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}$
commutes judgmentally.
\end{defn}

\begin{defn}
A \emph{substitution algebra in context $\Gamma$} is a pre-substitution algebra
in context $\Gamma$ for which substitution is a pre-substitution homomorphism.
\end{defn}

\begin{rmk}
The condition that the substitution of a pre-substitution algebra is a
pre-substitution homomorphism assert that the diagram
\begin{small}
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxwk
  {\cftalgt{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
  {\cftfamalg{\cftfamalg{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}}
  \ar{r}
    { \ctxwk
        {\cftalgt{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
        {\cftfamhom{\cftfamhom{\cftsubst{\cftalg{A}}}}}
      }
  \ar{d}[swap]
    {\cftsubst{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
  &
\jcomp{}
  { \cftsubstt{\cftalg{A}}}
  { \jcomp{}
      { \cftsubstf{\cftalg{A}}}
      { \jcomp{}
          { \cftsubstc{\cftalg{A}}}
          { \ctxwk
              { \cftalgt
                  { \ctxwk
                      { \cftalgt{\cftalg{A}}}
                      { {\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
                    }
                }
              { \cftfamalg
                  { \cftfamalg
                      { \ctxwk
                          { \cftalgt{\cftalg{A}}}
                          { {\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
                        }
                    }
                }
            }
        }
    }
  \ar{d}
    { \jcomp{}
        { \cftsubstt{\cftalg{A}}}
        { \jcomp{}
            { \cftsubstf{\cftalg{A}}}
            { \jcomp{}
                { \cftsubstc{\cftalg{A}}}
                { \cftsubst
                    { \ctxwk
                        {\cftalgt{\cftalg{A}}}
                        {{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
                      }
                  }
              }
          }
      }
  \\
\ctxwk
  { \cftalgt{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
  { {\cftalgf{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
    {\cftfamalg{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
    }
  \ar{r}[swap,yshift=-1em]
    { \ctxwk
        { \cftalgt{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
        { {\cftalgf{\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}}
          {\cftfamhom{\cftsubst{\cftalg{A}}}}}
          }
  &
\jcomp{}
  { \cftsubstt{\cftalg{A}}}
  { \jcomp{}
      { \cftsubstf{\cftalg{A}}}
      { \jcomp{}
          { \cftsubstc{\cftalg{A}}}
          { \ctxwk
              { \cftalgt
                  { \ctxwk
                      {\cftalgt{\cftalg{A}}}
                      {{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
                    }
                }
              { { \cftalgf
                    { \ctxwk
                        { \cftalgt{\cftalg{A}}}
                        {{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
                      }
                  }
                { \cftfamalg
                    { \ctxwk
                        {\cftalgt{\cftalg{A}}}
                        {{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
                      }
                  }
                }
            }
        }
    }
\end{tikzcd}
\end{equation*}
\end{small}
\end{rmk}

\begin{defn}
A substitution homomorphism is a pre-substitution homomorphism between
substitution algebras.
\end{defn}

\begin{thm}
Suppose $\cftalg{A}$ is a substitution algebra in context $\Gamma$. Then
the pre-substitution algebra $\cftfamalg{\cftalg{A}}$ is a substitution algebra
in context 
$\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is a substitution algebra in context $\ctxext{\Gamma}{B}$ 
and let $\jfam{\Gamma}{A}$. Then the pre-substitution algebra 
$\ctxwk{A}{\cftalg{Q}}$ is a substitution algebra in context 
$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$.
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is a substitution algebra in context $\ctxext{{\Gamma}{A}}{P}$
and let $\jterm{\Gamma}{A}{x}$. Then the pre-substitution algebra 
$\subst{x}{\cftalg{Q}}$ is a substitution algebra in context 
$\ctxext{\Gamma}{\subst{x}{P}}$.
\end{thm}

\begin{cor}
In a substitution algebra $\cftalg{A}$ in context $\Gamma$, substitution is a
substitution homomorphism.
\end{cor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{E-algebras}
\begin{defn}
An extension algebra $\cftalg{A}$ in context $\Gamma$ with internal empty context,
empty family, weakening, identity terms and
substitution
\begin{align*}
\jalign\jhom
  {{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}
  {\ctxwk{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}
  {\cftfamalg{\cftfamalg{\cftalg{A}}}}
  {\cftwk{\cftalg{A}}}
  \\
\jalign\jterm
  { { {\Gamma}
      {\cftalgc{\cftalg{A}}}
      }
    { \cftalgf{\cftalg{A}}
      }
    }
  {\subst{{\idtm{\cftalgf{\cftalg{A}}}}{\cftwkc{\cftalg{A}}}}{\cftalgt{\cftfamalg{\cftalg{A}}}}}
  {\cftidtm{\cftalg{A}}}
  \\
\jalign\jhom
  {{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}}
  {\ctxwk{\cftalgt{\cftalg{A}}}{\cftalg{F_{F_A}}}}
  {\ctxwk{\cftalgt{\cftalg{A}}}{{\cftalgf{\cftalg{A}}}{\cftfamalg{\cftalg{A}}}}}
  {\cftsubst{\cftalg{A}}}
\end{align*}
is said to be an \emph{E-algebra in context $\Gamma$} if both weakening and substitution are both
projection and substitution homomorphisms and if (1) the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxwk
  { \cftalgt{\cftalg{A}}}
  { { \cftalgf{\cftalg{A}}}
    { \cftfamalg{\cftalg{A}}}
    }
  \ar{r}{\ctxwk{\cftalgt{\cftalg{A}}}{\cftwk{\cftalg{A}}}}
  \ar{dr}[swap]{\cftidhom{}}
  &
\ctxwk
  { \cftalgt{\cftalg{A}}}
  { \cftfamalg{\cftfamalg{\cftalg{A}}}}
  \ar{d}{\cftsubst{\cftalg{A}}}
  \\
  &
\ctxwk
  { \cftalgt{\cftalg{A}}}
  { { \cftalgf{\cftalg{A}}}
    { \cftfamalg{\cftalg{A}}}
    }
\end{tikzcd}
\end{equation*}
of extension algebras in context 
$\ctxext{{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}}{\cftalgt{\cftalg{A}}}$
commutes judgmentally
and (2) the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cftfamalg{\cftfamalg{\cftalg{A}}}
  \ar{r}{\subst{\idtm{\cftalgf{\cftalg{A}}}}{\cftfamhom{\cftwk{\cftalg{A}}}}}
  \ar{dr}[swap]{\cftidhom{}}
  &
\subst
  { \idtm{\cftalgf{\cftalg{A}}}}
  { \jcomp{}
      { \cftwk{\cftalg{A}}}
      { \cftfamalg{\cftfamalg{\cftfamalg{\cftalg{A}}}}}
    }
  \ar{d}
    { \subst
        { \cftidtm{\cftalg{A}}}
        { { { \idtm{\cftalgf{\cftalg{A}}}}
            { \cftwkc{\cftalg{A}}}
            }
          { \cftsubst{\cftalgf{\cftalg{A}}}}
          }
      }
  \\
  &
    { \subst
        { \cftidtm{\cftalg{A}}}
        { { { \idtm{\cftalgf{\cftalg{A}}}}
            { \cftwkc{\cftalg{A}}}
            }
          { \ctxwk
              { \cftalgt{\cftfamalg{\cftalg{A}}}}
              { { \cftalgf{\cftfamalg{\cftalg{A}}}}
                { \cftfamalg{\cftfamalg{\cftalg{A}}}}
                }
            }
          }
      }
\end{tikzcd}
\end{equation*}
of extension algebras in context 
$\ctxext{{\Gamma}{\cftalgc{\cftalg{A}}}}{\cftalgf{\cftalg{A}}}$ commutes
judgmentally.

E-algebras are considered judgmentally equal if their underlying extension algebras
and their internal weakening, identity terms and substitution are judgmentally
equal.
\end{defn}

\begin{thm}
Suppose $\cftalg{A}$ is an E-algebra in context $\Gamma$. Then
$\cftfamalg{\cftalg{A}}$ is an E-algebra in context $\ctxext{\Gamma}{\cftalgc{\cftalg{A}}}$.
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is an E-algebra in context $\ctxext{\Gamma}{B}$ and that
$\jfam{\Gamma}{A}$. Then $\ctxwk{A}{\cftalg{Q}}$ is an E-algebra in context
$\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}$. 
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is an E-algebra in context $\ctxext{{\Gamma}{A}}{P}$ and
that $\jterm{\Gamma}{A}{x}$. Then $\subst{x}{\cftalg{Q}}$ is an E-algebra in
context $\ctxext{\Gamma}{\subst{x}{P}}$.
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is an E-algebra in context $\ctxext{\Gamma}{B}$ and that
$\jfam{\Gamma}{A}$. Then we have the
judgmental equality $\cftfamalg{\ctxwk{A}{\cftalg{Q}}}\jdeq\ctxwk{A}{\cftfamalg{\cftalg{Q}}}$
of E-algebras in context $\ctxext{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{\cftalgc{\cftalg{Q}}}}$.
\end{thm}

\begin{thm}
Suppose $\cftalg{Q}$ is an E-algebra in context $\ctxext{{\Gamma}{A}}{P}$ and
that $\jterm{\Gamma}{A}{x}$. Then we have the judgmental equality
$\cftfamalg{\subst{x}{\cftalg{Q}}}\jdeq\subst{x}{\cftfamalg{\cftalg{Q}}}$
of E-algebras in context $\ctxext{{\Gamma}{\subst{x}{P}}}{\subst{x}{\cftalgc{\cftalg{Q}}}}$.
\end{thm}

\subsection{E-homomorphisms}
\begin{defn}
An \emph{E-homomorphism from $\cftalg{A}$ to $\cftalg{B}$ in context $\Gamma$}
is a extension homomorphism which is both a projection homomorphism and a
substitution homomorphism.
\end{defn}

\subsection{E-transformations}
E-transformations are going to be the analogues of natural transformations.

\subsection{E-adjunctions}

\subsection{Pre-universes}
Pre-universes are internal versions of the theory of contexts, families and
terms. They interpret extension, the empty context, weakening, substitution
and identity terms all at once in a compatible way. Besides the compatibility
properties there will be judgmental equalities analoguous to the cancellation
properties of \autoref{cancellation-ws,cancellation-i}. Pre-universes are to
the theory of contexts, families and terms what internal categories to a
category.