### 2014-09-22 article-models-intro.tex
intro

As of yet, there is no theory of internal models of type theory of sufficient
generality. The problem of internalizing type theory has two facets. On the one
hand one may choose to model the definitional equalities of type theory by
identifications. While this approach would have the benefint of being completely
internal (allowing one to use the internal logic of type theory to prove things
about internal models), complications in this approach do arise when the 
underlying types are not sets. This immediately leaves univalent universes beyond
the scope of that program. On the other hand, one may want to interpret the
judgmental equalities of type theory with new judgmental equalities. The problem
here is that the theory of judgmental equality in many variations of dependent
type theory is not developed systematically, leaving anyone essentially in the
dark on this matter.