### 2014-10-28 Rules/rules_projections.tex

\subsection{The theory of projections}
\label{identityterms}

We formulate the theory of projections on top of the theory of weakening
described in \autoref{weakening}. In the theory of projections we introduce
the identity terms. The identity term at a family $A$ in context $\Gamma$ is
a term $\jterm{{\Gamma}{A}}{\ctxwk{A}{A}}{\idtm{A}}$, so weakening is necessary.
Together with weakening, the identity terms will provide for the projections.
The projection from $\cftext{A}{P}$ to $A$ will be the term
$\jterm{{\Gamma}{{A}{P}}}{\ctxwk{\ctxext{A}{P}}{A}}{\ctxwk{P}{\idtm{A}}}$ and
will be studied extensively in \autoref{extension-on-terms}.

The following inference rules are required to be valid:
\begin{align}
& \inference
  { \jfam{\Gamma}{A}
    }
  { \unfoldall{\jhom{\Gamma}{A}{A}{\idtm{A}}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      }
    { \unfoldall{\jhomeq{\Gamma}{A}{A}{\idtm{A}}{\idtm{A'}}}
      }
\end{align}

\subsubsection{The identity term of an extended family}
The fact that identity terms are compatible with extensions is derived in
\autoref{comp-ie}.

\subsubsection{The identity term of a weakened family}
\label{comp-wi}
The identity term of a weakened family is the weakened identity term:
\begin{equation}
\inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \unfoldall{\jhomeq
      {{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{Q}}
      {\ctxwk{A}{Q}}
      {\ctxwk{A}{\idtm{Q}}}
      {\idtm{\ctxwk{A}{Q}}}
      }
    }
  \label{comp-wi-t}
\end{equation}

\begin{rmk}
An important special case is the judgmental equality
\begin{equation*}
\unfoldall{\jhomeq
      {{\Gamma}{A}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{B}}
      {\ctxwk{A}{\idtm{B}}}
      {\idtm{\ctxwk{A}{B}}}
      }
\end{equation*}
for a family $\jfam{\Gamma}{B}$.
\end{rmk}