### 2015-02-10 Algebras/empty.tex
large reorganisation

\subsection{Extension algebras with empty context and families}

\begin{defn}
An extension algebra $\stesys$ is said to have \emph{empty families} if there
is a section
\begin{equation*}
\phi_1(\stesys):\stesysc\to\stesysf
\end{equation*}
of $\eft$, satisfying the following additional properties:
\begin{enumerate}
\item $\phi_1(\stesys)$ is also a section of $\ectxext$.
\item The unique morphism $\phi_2$ rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\stesysf
  \ar[bend left=20]{drr}{\phi_1(\stesys)\circ\ectxext}
  \ar[equals,bend right=20]{ddr}
  \ar[dotted]{dr}{\phi_2}
  \\
  {}&
\stesysf_2
  \ar{d}{\eft[2]}
  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
  &
\stesysf
  \ar{d}{\eft}
  \\
  {}&
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutative, is a section of $\efamext$.
\item The unique morphism $\iota_1$ rendering the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\stesysf
  \ar[equals,bend left=20]{drr}
  \ar[bend right=20]{ddr}[swap]{\phi_1(\stesys)\circ\eft}
  \ar[dotted]{dr}{\iota_1}
  \\
  {}&
\stesysf_2
  \ar{d}{\eft[2]}
  \ar{r}[swap]{\pullbackpr{2}{\ectxext}{\eft}}
  &
\stesysf
  \ar{d}{\eft}
  \\
  {}&
\stesysf
  \ar{r}[swap]{\ectxext}
  &
\stesysc
\end{tikzcd}
\end{equation*}
commutative, is a section of $\efamext$.
\end{enumerate}
\end{defn}

\begin{defn}
A homomorphism of extension algebras with empty families is an extension
homomorphism $f:\stesys'\to\stesys$ for which the diagram
\begin{equation*}
\begin{tikzcd}
\stesysf'
  \ar{r}{f_1}
  &
\stesysf
  \\
\stesysc'
  \ar{u}{\phi_1(\stesys')}
  \ar{r}[swap]{f_0}
  &
\stesysc
  \ar{u}[swap]{\phi_1(\stesys)}
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{lem}
Suppose $\stesys$ is an extension algebra with empty families. Then
$\famesys{\stesys}$ is an extension algebra with empty families, with
$\phi_1(\famesys{\stesys}):=\phi_2$. 
\end{lem}

\begin{lem}
Suppose $\stesys$ is an extension algebra with empty families and consider
$p:\stesysc\rightarrow X\leftarrow Y:p$. Then
$\cobesys{Y}{\stesys}{g}{p}$ is an extension algebra with empty families
with $\phi_1(\cobesys{Y}{\stesys}{g}{p}):=g^\ast(\phi_1)$.
\end{lem}