### 2015-03-12 weakeningcats.tex
The new E-systems

\section{Weakening systems}

\begin{defn}
A \emph{pre-weakening structure on a category $\cat{C}$} consists of a functor
$W_f:\cat{C}/Y\to \cat{C}/X$ for every $f:X\to Y$ in $\cat{C}$ such 
that
\begin{enumerate}
\item $W_{\catid{X}}=\catid{\cat{C}/X}$ for every object $X\in\cat{C}$.
\item $W_{f\circ f'}=W_{f'}\circ W_f$ for every $f':X'\to X$ and $f:X\to Y$ in
$\cat{C}$.
\item $W_f$ preserves the final object, i.e.~$W_f(\catid{Y})=\catid{X}$.
%\item $W^1_f\circ W^0_g= W^0_{W^1_f(g)}\circ W^1_f$.
\end{enumerate}
A \emph{pre-weakening system} is a category together with a pre-weakening structure.
\end{defn}

\begin{defn}
A \emph{pre-weakening homomorphism} $F:\cat{C}\to\cat{D}$ between pre-weakening systems
is a functor $F:\cat{C}\to\cat{D}$ such that the square
\begin{equation*}
\begin{tikzcd}
\cat{C}/X
  \arrow[r,"F/X"]
  &
\cat{D}/F(X)
  \\
\cat{C}/Y
  \arrow[u,"W_f"]
  \arrow[r,swap,"F/Y"]
  &
\cat{D}/F(Y)
  \arrow[u,swap,"W_{F(f)}"]
\end{tikzcd}
\end{equation*}
commutes for any $f:X\to Y$ in $\cat{C}$.
\end{defn}

\begin{lem}
Let $\cat{C}$ be a pre-weakening system. Then for any object $X$ of $\cat{C}$,
the slice category $\cat{C}/X$ is also a pre-weakening system.
\end{lem}

\begin{defn}
A \emph{weakening system} is a pre-weakening system for which every $W_f$ is a
pre-weakening homomorphism. A \emph{weakening homomorphism} is a pre-weakening 
homomorphism between weakening systems.
\end{defn}

\begin{cor}
For any object $X$ of a weakening system $\cat{C}$, the slice category
$\cat{C}/X$ is a weakening system.
\end{cor}

\begin{rmk}
If the underlying category is a group, the axioms of a weakening system
obtain the following form: There is a group homomorphism $W_x:G\to G$ for
every $x\in G$, in such a way that $W_e=\catid{G}$ and $W_{xy}=W_y\circ W_x$
for every $x,y\in G$, and for any $x,y\in G$ the diagram
\begin{equation*}
\begin{tikzcd}
G \arrow[r,"W_y"]
  &
G 
  \\
G \arrow[u,"W_x"]
  \arrow[r,swap,"W_y"]
  &
G \arrow[u,swap,"W_{W_y(x)}"]
\end{tikzcd}
\end{equation*}
is required to commute.
\end{rmk}

\begin{lem}
For any group, conjugation is a weakening structure. Thus, any group is
canonically a weakening system.
\end{lem}

\begin{proof}
Note that $e^{-1}xe=x$ and $(xy)^{-1}z(xy)=y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
Also, the diagram
\begin{equation*}
\begin{tikzcd}
G \arrow[r,"\mathrm{conj}_y"]
  &
G 
  \\
G \arrow[u,"\mathrm{conj}_x"]
  \arrow[r,swap,"\mathrm{conj}_y"]
  &
G \arrow[u,swap,"\mathrm{conj}_{y^{-1}xy}"]
\end{tikzcd}
\end{equation*}
commutes.
\end{proof}

\begin{rmk}
Similarly, every groupoid has a canonical weakening structure given by
conjugation.
\end{rmk}

\section{Weakening systems with term structure}
\begin{defn}
A \emph{category with a term structure} is a category $\cat{C}$ with a set $T(f)$ for every
morphism $f$ of $\cat{C}$. A \emph{functor with term structure} from $\cat{C}$ to $\cat{D}$
is a functor $F:\cat{C}\to\cat{D}$ with a function $T(f)\to T(F(f))$ for
every $f:X\to Y$ in $\cat{C}$.
\end{defn}

\begin{defn}
Suppose $\cat{C}$ is a category with term structure, and $X$ is an object of $\cat{C}$. 
Then $\cat{C}/X$ is a category with term structure, where $T_{\cat{C}/X}(g)=T_\cat{C}(g)$.
\end{defn}

\begin{defn}
A \emph{weakening system with term term structure} is the same thing as a weakening system,
except that the underlying category and the functors $W_f$ are required to
be functors with term term structure. The equalities are also required to hold on the term structure.
\end{defn}

\begin{cor}
For any object $X$ of a weakening system with term structure, the slice category
$\cat{C}/X$ is a weakening system with term structure.
\end{cor}

\section{Projection systems}
\begin{defn}
A \emph{pre-projection system} is a weakening system with term structure with an element
$\tfid{f}\in T(W_f(f))$ for every $f:X\to Y$ in $\cat{C}$. 
\end{defn}

\begin{defn}
A \emph{pre-projection homomorphism} $F:\cat{C}\to\cat{D}$ is a weakening homomorphism for which
\begin{equation*}
F(\tfid{f})=\tfid{F(f)}
\end{equation*}
for every $f:X\to Y$ in $\cat{C}$.
\end{defn}

\begin{lem}
For any object $X$ in a pre-projection system $\cat{C}$, the slice category
$\cat{C}/X$ is a pre-projection system.
\end{lem}

\begin{defn}
A \emph{projection system} is a pre-projection system for which every $W_f$ is
a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
between projection systems.
\end{defn}

\begin{cor}
For any object $X$ of a projection system $\cat{C}$, the slice category $\cat{C}/X$
is a projection system.
\end{cor}

\section{Substitution systems}

\begin{defn}
A \emph{pre-substitution structure} on a category with term structure $\cat{C}$ consists
of a functor with term structure $S_t:\cat{C}/X\to\cat{C}/Y$ for every $t\in T(f)$ and
$f:X\to Y$, such that $S_t(\catid{X})=\catid{Y}$.

A \emph{pre-substitution system} is a category with term structure together with a
pre-substitution structure. 
\end{defn}

\begin{defn}
A \emph{pre-substitution homomorphism} $F:\cat{C}\to\cat{D}$ is a functor with term structure for
which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{C}/X
  \arrow[r,"F/X"]
  \arrow[d,swap,"S_t"]
  &
\cat{D}/F(X)
  \arrow[d,"S_{F(t)}"]
  \\
\cat{C}/Y
  \arrow[r,swap,"F/Y"]
  &
\cat{D}/F(Y)
\end{tikzcd}
\end{equation*}
commutes.
\end{defn}

\begin{lem}
For any object $X$ of a pre-substitution system $\cat{C}$, the slice category
$\cat{C}/X$ is a pre-substitution system.
\end{lem}

\begin{defn}
A \emph{substitution system} is a pre-substitution system for which each
$S_t$ is a pre-substitution homomorphism. A \emph{substitution homomorphism} is a
pre-substitution homomorphism between substitution systems.
\end{defn}

\begin{cor}
For any object $X$ of a substitution system, the slice category $\cat{C}/X$
is a substitution system.
\end{cor}

\section{E-systems}
\begin{defn}
An \emph{E-system} is a category with term structure with
\begin{itemize}
\item a weakening structure $W$,
\item the structure $\tfid{}$ of a projection system,
\item a substitution structure $S$,
\end{itemize}
such that
\begin{enumerate}
\item each $W_f$ is a substitution homomorphism;
\item each $S_t$ is a projection homomorphism;
\item $S_t(\tfid{f})=t$ for any $t\in T(f)$ and $f:X\to Y$;
\item $S_t\circ W_f=\catid{\cat{C}/Y}$ for any $t\in T(f)$ and $f:X\to Y$;
\item $S_{\tfid{f}}\circ W_{W_f(f)}=\catid{\cat{C}/X}$ for any $f:X\to Y$.
\end{enumerate}
\end{defn}

\begin{cor}
For any object $X$ of an E-system $\cat{C}$, the slice category $\cat{C}/X$
is an E-system.
\end{cor}

\begin{defn}
An \emph{E-homomorphism} $F:\cat{C}\to\cat{D}$ is a functor which is both a projection
homomorphism and a substitution homomorphism.
\end{defn}

\begin{comment}
\begin{lem}
Suppose $\cat{D}$ is a retract of a category $\cat{C}$, i.e. there are functors
$S:\cat{D}\to\cat{C}$ and $R:\cat{C}\to\cat{D}$ with a natural isomorphism
$\tau:R\circ S \simeq \catid{\cat{D}}$, and suppose that $\cat{C}$ is a weakening
category. Then $\cat{D}$ is a weakening system too.
\end{lem}

\begin{proof}
Let $f:X\to Y$ be a morphism in $\cat{D}$. Then we define $W_f$ to be the
composition
\begin{equation*}
\begin{tikzcd}[column sep=3.2em]
\cat{D}/Y
  \arrow[r,"S/Y"]
  &
\cat{C}/S(Y)
  \arrow[r,"W_{S(f)}"]
  &
\cat{C}/S(X)
  \arrow[r,"R/S(X)"]
  &
\cat{D}/R(S(X))
  \arrow[r,"\tau_X\circ{-}"]
  &
\cat{D}/X
\end{tikzcd}
\end{equation*}
We first verify whether this assembles a pre-weakening structure on $\cat{D}$.
\begin{enumerate}
\item Let $X$ be an object of $\cat{D}$. Then $W_{\catid{X}}$ is the composition
\begin{equation*}
\begin{tikzcd}[column sep=3.2em]
\cat{D}/X
  \arrow[r,"S/X"]
  &
\cat{C}/S(X)
  \arrow[r,"R/S(X)"]
  &
\cat{D}/R(S(X))
  \arrow[r,"{-}\circ\tau_X"]
  &
\cat{D}/X
\end{tikzcd}
\end{equation*}
This functor maps $g:X'\to X$ consecutively to $S(g)$, to $R(S(g))$, to
$\tau_X\circ R(S(g))$. Note that $\tau_X\circ R(S(g))=g\circ \tau_{X'}$ by
naturality.
\end{enumerate}
\end{proof}
\end{comment}