### 2015-04-24 weakeningcats.tex
Theorems about E-systems

\begin{rmk}
Recall that the slice pre-category $\cat{C}/\Gamma/(\Delta,f)$ is isomorphic to the slice pre-category
$\cat{C}/\Delta$, in the essentially algebraic theory of pre-categories. In the following,
we shall usually identify these two categories. Recall also that for every functor
$F:\cat{C}\to\cat{D}$ and every object $\Delta$ of $\cat{C}$, there is a functor
$F/\Delta:\cat{C}/\Delta\to \cat{D}/F(\Delta)$ given by the action on morphisms of $F$. 
\end{rmk}

\begin{rmk}
It will be convenient to denote the objects of the category $\cat{C}$ with capital
greek letters $\Gamma$, $\Delta$, \ldots, and the morphisms by capital latin letters
$A$, $B$, $P$, $Q$, etcetera. When $A$ is a morphism into $\Gamma$, we will denote
its codomain by $\ctxext{\Gamma}{A}$. When $A$ is a morphism into $\Gamma$ and
$P$ is a morphism into $\ctxext{\Gamma}{A}$, it will be convenient to denote the
composition of $P$ with $A$ by $\ctxext{A}{P}$ as well. Note that $\ctxext{A}{P}$
is indeed the domain of $P$ in the slice category $\cat{C}/\Gamma$. Also, objects
of $\cat{C}$ will always have a capital greek letter in their notation, e.g.~as
in $\ctxext{{\Gamma}{A}}{P}$. 
\end{rmk}

\section{Weakening systems}

\begin{defn}
A \emph{pre-weakening structure on a pre-category $\cat{C}$} consists of a functor
$W_A:\cat{C}/\Gamma\to \cat{C}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{C}$ such 
that
\begin{enumerate}
\item $W_{\catid{\Gamma}}=\catid{\cat{C}/\Gamma}$ for every object $\Gamma\in\cat{C}$.
\item $W_{\ctxext{A}{P}}=W_P\circ W_A$ for every $P\in\cat{C}/\ctxext{\Gamma}{A}$ and $A\in\cat{C}/\Gamma$.
\item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})=\catid{\ctxext{\Gamma}{A}}$.
\end{enumerate}
A \emph{pre-weakening system} is a pre-category together with a pre-weakening structure.
\end{defn}

\begin{defn}
A \emph{pre-weakening homomorphism} $F:\cat{C}\to\cat{D}$ between pre-weakening systems
is a functor $F:\cat{C}\to\cat{D}$ such that the square
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{C}/\ctxext{\Gamma}{A}
  \arrow[r,"F/\ctxext{\Gamma}{A}"]
  &
\cat{D}/F(\ctxext{\Gamma}{A})
  \\
\cat{C}/\Gamma
  \arrow[u,"W_A"]
  \arrow[r,swap,"F/\Gamma"]
  &
\cat{D}/F(\Gamma)
  \arrow[u,swap,"W_{F(A)}"]
\end{tikzcd}
\end{equation*}
commutes for any $A\in\cat{C}/\Gamma$.
\end{defn}

\begin{lem}
Let $\cat{C}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{C}$,
the slice pre-category $\cat{C}/\Gamma$ is also a pre-weakening system, with the weakening
structure given by $W(\cat{C}/\Gamma)_P=W(\cat{C})_P$ for each $P\in\cat{C}/\ctxext{\Gamma}{A}$
and each $A\in\cat{C}/\Gamma$.
\end{lem}

\begin{defn}
A \emph{weakening system} is a pre-weakening system for which every $W_f$ is a
pre-weakening homomorphism. A \emph{weakening homomorphism} is a pre-weakening 
homomorphism between weakening systems.
\end{defn}

\begin{rmk}
In other words, a weakening system is a pre-weakening system with a weakening
structure $W$ for which the square
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{C}/\ctxext{{\Gamma}{B}}{Q}
  \arrow[r,"W_A/\ctxext{B}{Q}"]
  &
\cat{C}/\ctxext{{\Gamma}{A}}{W_A(\ctxext{B}{Q})}
  \\
\cat{C}/\ctxext{\Gamma}{B}
  \arrow[u,"W_Q"]
  \arrow[r,swap,"W_A/B"]
  &
\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
  \arrow[u,swap,"W_{W_A(Q)}"]
\end{tikzcd}
\end{equation*}
for each $A,B\in\cat{C}/\Gamma$ and $Q:\cat{C}/\ctxext{\Gamma}{B}$. The objects part of this 
property asserts that for any $k\in \cat{C}/\greek{E}$, the dotted arrows in the diagram
\begin{equation*}
\begin{tikzcd}[column sep=tiny]
\bullet \arrow[dr,densely dotted,swap,near start,"(W_A/\ctxext{B}{Q})(W_Q(R))"] & & \bullet \arrow[dl,densely dotted,near start,"W_{(W_A/B)(Q)}((W_A/B)(R))"] & & & & & & & & & & \bullet \arrow[dl,"W_Q(R)"] \\
& \bullet \arrow[dr,swap,"(W_A/B)(Q)"] & & \bullet \arrow[dl,"(W_A/B)(R)"] & & & & & & & & \bullet \arrow[dr,swap,"Q"] & & \bullet \arrow[dl,"R"] \\
& & \bullet \arrow[d,swap,"W_A(B)"] & & & & & & & & & & \ctxext{\Gamma}{B} \arrow[d,"B"] \\
& & \ctxext{\Gamma}{A} \arrow[rrrrrrrrrr,swap,"A"] & & & & & & & & & & \Gamma
\end{tikzcd}
\end{equation*}
are equal.

A useful special case of this property is where $B=\catid{\Gamma}$. Thus, if $W$ is
a weakening system, then the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{C}/\ctxext{\Gamma}{Q}
  \arrow[r,"W_A/Q"]
  &
\cat{C}/\ctxext{{\Gamma}{A}}{W_A(Q)}
  \\
\cat{C}/\Gamma
  \arrow[u,"W_Q"]
  \arrow[r,swap,"W_A"]
  &
\cat{C}/\ctxext{\Gamma}{A}
  \arrow[u,swap,"W_{W_A(Q)}"]
\end{tikzcd}
\end{equation*}
commutes for every $A\in\cat{C}/\Gamma$ and $Q\in\cat{C}/\Gamma$. 
\end{rmk}

\begin{cor}
For any object $\Gamma$ of a weakening system $\cat{C}$, the slice pre-category
$\cat{C}/\Gamma$ is a weakening system.
\end{cor}

\begin{eg}
If the underlying pre-category is a group, the axioms of a weakening system
obtain the following form: There is a group homomorphism $W_x:G\to G$ for
every $x\in G$, in such a way that $W_e=\catid{G}$ and $W_{xy}=W_y\circ W_x$
for every $x,y\in G$, and for any $x,y\in G$ the diagram
\begin{equation*}
\begin{tikzcd}
G \arrow[r,"W_y"]
  &
G 
  \\
G \arrow[u,"W_x"]
  \arrow[r,swap,"W_y"]
  &
G \arrow[u,swap,"W_{W_y(x)}"]
\end{tikzcd}
\end{equation*}
is required to commute. Thus, a weakening system on a group $G$ consists of
a group homomorphism $W:\op G\to\mathrm{Aut}(G)$ such that $W_a\circ W_b=
W_{W_a(b)}\circ W_a$. The latter requirement is satisfied if $b\cdot a=a\cdot W_a(b)$
for every $a,b\in G$, or equivalently if $[b,a]=b^{-1}\cdot W_a(b)$.
\end{eg}

\begin{lem}
For any group, conjugation is a weakening structure. Thus, any group is
canonically a weakening system.
\end{lem}

\begin{proof}
Note that $e^{-1}xe=x$ and $(xy)^{-1}z(xy)=y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
Also, the diagram
\begin{equation*}
\begin{tikzcd}
G \arrow[r,"\mathrm{conj}_y"]
  &
G 
  \\
G \arrow[u,"\mathrm{conj}_x"]
  \arrow[r,swap,"\mathrm{conj}_y"]
  &
G \arrow[u,swap,"\mathrm{conj}_{y^{-1}xy}"]
\end{tikzcd}
\end{equation*}
commutes.
\end{proof}

\begin{comment}
\begin{rmk}
Similarly, every groupoid has a canonical weakening structure given by
conjugation.
\end{rmk}
\end{comment}

\begin{lem}
Let $G$ be a graph, and let $\cat{C}_G$ be the free pre-category generated by
$G$. A pre-weakening structure on $\cat{C}_G$ is determined by functors
$W_e:\cat{C}_G/\Gamma\to\cat{C}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
property that $W_e(\catid{\Gamma})=\catid{\Delta}$.

A weakening structure on $\cat{C}_G$ is
a pre-weakening structure on $\cat{C}_G$ such that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{C}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{C}_G/\mathrm{dom}(W_e(g\circ h)) \\
\cat{C}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{C}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
\end{tikzcd}
\end{equation*}
commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{C}_G$ and every two edges $e:\Delta\to \Gamma$ and
$h:\greek{Z}\to \greek{E}$ in $G$.
\end{lem}

\begin{eg}
Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
Then $\cat{C}_G/\ast=\N$. Then the only weakening structure on $\cat{C}/G$ is
given by $W_e=\catid{\N}$. The reason is that 
\end{eg}

\begin{lem}
Let $P$ be a pre-order. A pre-weakening structure on $P$ is determined by order
preserving maps $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$ satisfying
$W_{a,b}(b)=a$, for every $a\leq b$, where
${\downarrow}(x)$ is the set $\{y\in P\mid y\leq x\}$.

A weakening structure on $P$ is a pre-weakening structure on $P$ such that the
diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
{\downarrow}(c') \arrow[r,"W_{a,b}|_{{\downarrow}(c')}"] & {\downarrow}(W_{a,b}(c')) \\
{\downarrow}(c) \arrow[u,"W_{c',c}"] \arrow[r,swap,"W_{a,b}|_{{\downarrow}(c)}"] & {\downarrow}(W_{a,b}(c)) \arrow[u,swap,"W_{W_{a,b}(c'),W_{a,b}(c)}"]
\end{tikzcd}
\end{equation*}
for any $a\leq b$ and $c'\leq c\leq b$.
\end{lem}

\begin{eg}
In a poset with meets, the functions $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$
given by $W_{a,b}(x):= a\land x$ assemble a weakening structure, because for
any $a\leq b$ and $x,c'\leq c\leq b$ we have $a\land(c'\land x)=(a\land c')\land
(a\land x)$.  
\end{eg}

\begin{eg}
Consider the poset $(\N,\geq)$, which has a morphism from $a$ to $b$ if $a\geq b$.
On this poset, we may consider the weakening structure defined by $W_{a,b}(x)=
a+(x-b)$.
\end{eg}

\section{Weakening systems with term structure}
\begin{defn}
A \emph{category with a term structure} is a pre-category $\cat{C}$ with a set $T(A)$ for every
$A\in\cat{C}/\Gamma$ for every $\Gamma\in\cat{C}$. A \emph{functor with term structure} from $\cat{C}$ to $\cat{D}$
is a functor $F:\cat{C}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
every $A\in\cat{C}/\Gamma$, for every $\Gamma\in\cat{C}$.
\end{defn}

\begin{defn}
Suppose $\cat{C}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{C}$. 
Then $\cat{C}/\Gamma$ is a pre-category with term structure, where $T_{\cat{C}/\Gamma}(A)=T_\cat{C}(A)$.
\end{defn}

\begin{defn}
A \emph{weakening system with term term structure} is the same thing as a weakening system,
except that the underlying pre-category and the functors $W_A$ are required to
be functors with term term structure. The equalities are also required to hold on the term structure.
\end{defn}

\begin{cor}
For any object $\Gamma$ of a weakening system with term structure, the slice pre-category
$\cat{C}/\Gamma$ is a weakening system with term structure.
\end{cor}

\begin{eg}
%For any group $G$ with weakening structure $W:\op{G}\to\mathrm{Aut}(G)$, a term 
%structure is determined by a presheaf on $G/\ker(W)$.

Let $G$ be a group, with the weakening structure given by conjugation. A term
structure on $G$ is given by a set $X$ with a right $G$-action $\mu:\op G\times X\to X$.
We take $T(a):= X$ for every $a\in G$.
\end{eg}

\section{Projection systems}
\begin{defn}
A \emph{pre-projection system} is a weakening system with term structure with an element
$\tfid{A}\in T(W_A(A))$ for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$. 
\end{defn}

\begin{defn}
A \emph{pre-projection homomorphism} $F:\cat{C}\to\cat{D}$ is a weakening homomorphism for which
\begin{equation*}
F(\tfid{A})=\tfid{F(A)}
\end{equation*}
for every $A\in\cat{C}/\Gamma$ and $\Gamma\in\cat{C}$.
\end{defn}

\begin{lem}
For any object $\Gamma$ in a pre-projection system $\cat{C}$, the slice pre-category
$\cat{C}/\Gamma$ is a pre-projection system.
\end{lem}

\begin{defn}
A \emph{projection system} is a pre-projection system for which every $W_A$ is
a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
between projection systems.
\end{defn}

\begin{cor}
For any object $\Gamma$ of a projection system $\cat{C}$, the slice pre-category $\cat{C}/\Gamma$
is a projection system.
\end{cor}

\begin{eg}
Consider a group $G$ with a right $G$-set $X$. A projection structure on $(G,X)$
consists of an element $\tfid{a}\in X$ for every $a\in G$, so that $\tfid{a}\cdot b=
\tfid{b^{-1}ab}$ for every $a,b\in G$.

%For instance, this is the case if we take $X$ to be a pointed set with $\ast\in X$,
%and take $\tfid{a}:=\ast$ for all $a\in G$.
\end{eg}

\section{Substitution systems}

\begin{defn}
A \emph{pre-substitution structure} on a pre-category with term structure $\cat{C}$ consists
of a functor with term structure $S_x:\cat{C}/\ctxext{\Gamma}{A}\to\cat{C}/\Gamma$ for every $x\in T(A)$ and
$A\in\cat{C}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})=\catid{\Gamma}$.

A \emph{pre-substitution system} is a pre-category with term structure together with a
pre-substitution structure. 
\end{defn}

\begin{defn}
A \emph{pre-substitution homomorphism} $F:\cat{C}\to\cat{D}$ is a functor with term structure for
which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{C}/\ctxext{\Gamma}{A}
  \arrow[r,"F/\ctxext{\Gamma}{A}"]
  \arrow[d,swap,"S_x"]
  &
\cat{D}/F(\ctxext{\Gamma}{A})
  \arrow[d,"S_{F(x)}"]
  \\
\cat{C}/\Gamma
  \arrow[r,swap,"F/\Gamma"]
  &
\cat{D}/F(\Gamma)
\end{tikzcd}
\end{equation*}
commutes for every $x\in T(A)$ and $A\in\cat{C}/\Gamma$.
\end{defn}

\begin{lem}
For any object $\Delta$ of a pre-substitution system $\cat{C}$, the slice pre-category
$\cat{C}/\Delta$ is a pre-substitution system.
\end{lem}

\begin{defn}
A \emph{substitution system} is a pre-substitution system for which each
$S_t$ is a pre-substitution homomorphism. A \emph{substitution homomorphism} is a
pre-substitution homomorphism between substitution systems.
\end{defn}

\begin{cor}
For any object $\Delta$ of a substitution system, the slice pre-category $\cat{C}/\Delta$
is a substitution system.
\end{cor}

\begin{rmk}
The condition that every $S_x$ is a substitution homomorphism, asserts that
the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
  \arrow[r,"S_x/\ctxext{P}{Q}"]
  \arrow[d,swap,"S_y"]
  &
\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
  \arrow[d,"S_{S_x(y)}"]
  \\
\cat{C}/\ctxext{{\Gamma}{A}}{P}
  \arrow[r,swap,"S_x/P"]
  &
\cat{C}/\ctxext{\Gamma}{S_x(P)}
\end{tikzcd}
\end{equation*}
commutes for every $y\in T(Q)$.
\end{rmk}

\begin{eg}
Let $G$ be a group and let $H$ be a set, forming a category with term 
structure of which $T(a)=H$ for every $a\in G$. A substitution
system on $(G,H)$ consists of a function $\sigma_G : H\times G\to G$ and
a function $\sigma_H:H\times H\to H$, for which the diagrams
\begin{equation*}
\begin{tikzcd}[column sep=large]
G \arrow[r,"\sigma_G(h)"]
  \arrow[d,swap,"\sigma_G(k)"]
  &
G \arrow[d,"{\sigma_G(\sigma_H(h,k))}"]
  \\
G \arrow[r,swap,"\sigma_G(h)"] 
& G
\end{tikzcd}
\end{equation*}
and
\begin{equation*}
\begin{tikzcd}[column sep=large]
H \arrow[r,"\sigma_H(h)"]
  \arrow[d,swap,"\sigma_H(k)"]
  &
H \arrow[d,"{\sigma_H(\sigma_H(h,k))}"]
  \\
H \arrow[r,swap,"\sigma_H(h)"]
  &
H
\end{tikzcd}
\end{equation*}
commute for every $h,k\in H$. 
\end{eg}

\section{E-systems}
\begin{defn}
\label{defn:esystems}
An \emph{E-system} is a pre-category $\cat{C}$ with term structure with
\begin{itemize}
\item a weakening structure $W$,
\item the structure $\tfid{}$ of a projection system,
\item a substitution structure $S$,
\end{itemize}
such that
\begin{enumerate}
\item each $W_A$ is a substitution homomorphism, i.e.~the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{C}/\ctxext{{\Gamma}{B}}{Q}
  \arrow[r,"W_A/\ctxext{B}{Q}"]
  \arrow[d,swap,"S_y"]
  &
\cat{C}/\ctxext{{{\Gamma}{A}}{W_A(B)}}{W_A(Q)}
  \arrow[d,"S_{W_A(y)}"]
  \\
\cat{C}/\ctxext{\Gamma}{B}
  \arrow[r,swap,"W_A/B"]
  &
\cat{C}/\ctxext{{\Gamma}{A}}{W_A(B)}
\end{tikzcd}
\end{equation*}
commutes for every $Q\in\cat{C}/\ctxext{\Gamma}{B}$, $B\in\cat{C}/\Gamma$ and each 
$y\in T(Q)$; 
\item each $S_x$ is a projection homomorphism, i.e.~the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{C}/\ctxext{{\Gamma}{A}}{P}
  \arrow[r,"S_x/P"]
  \arrow[d,swap,"W_Q"]
  &
\cat{C}/\ctxext{\Gamma}{S_x(P)}
  \arrow[d,"W_{S_x(Q)}"]
  \\
\cat{C}/\ctxext{{{\Gamma}{A}}{P}}{Q}
  \arrow[r,swap,"S_x/\ctxext{P}{Q}"]
  &
\cat{C}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
\end{tikzcd}
\end{equation*}
commutes for every $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$, and we have
\begin{equation*}
S_x(\tfid{Q})=\tfid{S_x(Q)}
\end{equation*}
for every $Q\in\cat{C}/\ctxext{{\Gamma}{A}}{P}$;
\item $S_x(\tfid{A})=x$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$;
\item \label{tTf:StWf_id} $S_x\circ W_A=\catid{\cat{C}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{C}/\Gamma$;
\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A=\catid{\cat{C}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{C}/\Gamma$.
\end{enumerate}
An \emph{E-homomorphism} $F:\cat{C}\to\cat{D}$ is a functor which is both a projection
homomorphism and a substitution homomorphism.
\end{defn}

\begin{cor}
For any object $\Delta$ of an E-system $\cat{C}$, the slice pre-category $\cat{C}/\Delta$
is an E-system.
\end{cor}

\begin{rmk}[Needs more precision]
Consider the category $E$ generated by the graph consisting of the objects of 
the underlying category of an $E$-system $\mathcal{E}$ as edges, and the morphisms
$S_t:X\to Y$ and $W_A:Y\to X$ as edges, subject to the axioms of an $E$-systems.

Any morphism of $E$ can be rewritten in a unique way
as a composition $\tilde S\circ \tilde W$, where $\tilde S$ is a morphism in
substitution-normal form, $\tilde W$ is in weakening normal form, every term
expression is in normal form, and where
each occurence of $S_x\circ W_A$ and $S_{\tfid{A}}\circ W_A/A$ is eliminated.
\end{rmk}

\begin{defn}
An E-system of which the underlying category has a terminal object $\emptyc$, is said to
be \emph{E-system with empty context}. An E-homomorphism is said to preserve the
empty context when it preserves the terminal object $\emptyc$ on the nose.
\end{defn}

\begin{rmk}
When there is a terminal object, we get a weakening functor $W_\Gamma:\cat{C}\to
\cat{C}/\Gamma$ for every object $\Gamma$ of $\cat{C}$; we get a set $T(\Gamma)$;
and for $i\in T(\Gamma)$ we get substitutions $S_i:\cat{C}/\Gamma\to\cat{C}$. 
\end{rmk}

\section{Properties of E-systems}

\begin{defn}
For every $A,B\in\cat{C}/\Gamma$ we define the set 
\begin{equation*}
\thom{A}{B}:=T(W_A(B)).
\end{equation*} 
\end{defn}

\begin{rmk}
Note that $\thom{\catid{\Gamma}}{A}=T(A)$ for any $A\in\cat{C}/\Gamma$, 
because we have $W_{\catid{\ctxext{\Gamma}{A}}}=\catid{\cat{C}/\ctxext{\Gamma}{A}}$.

Note also that $\thom{\ctxext{A}{P}}{B}=\thom{P}{W_A(B)}$ for any $P\in\cat{C}/\ctxext{\Gamma}{A}$
and $B\in\cat{C}/\Gamma$, because $W_{\ctxext{A}{P}}=
W_P\circ W_A$.
\end{rmk}

\begin{defn}
Let $A,B\in\cat{C}/\Gamma$.
For any $f\in\thom{A}{B}$ we define the pre-composition E-homomorphism
\begin{equation*}
f_\ast := S_f\circ W_A/B : \cat{C}/\ctxext{\Gamma}{B}\to\cat{C}/\ctxext{\Gamma}{A}.
\end{equation*}
Furthermore, for $g\in\thom{B}{C}$ we may consider $W_B(C)\in\cat{C}/\ctxext{\Gamma}{B}$ as a morphism from $W_B(C)$ to $\catid{\ctxext{\Gamma}{B}}$. Therefore, $f_\ast$ acts on the terms of
$W_B(C)$, so it makes sense to define $g\circ f:=f_\ast(g)$. 
\end{defn}

Note that $g\circ f\in T(S_f(W_A/B(W_B(C))))$, whereas we would like that
$g\circ f\in\thom{A}{C}$. The following lemma confirms this:

\begin{lem}
Suppose $A$ and $B$ are morphisms of the underlying pre-category of $\mathcal{E}$
with the same codomain, and let $f\in\thom{A}{B}$. Then we have 
\begin{equation*}
S_f\circ (W_A/B)\circ W_B=W_A.
\end{equation*}
\end{lem}

\begin{proof}
Since weakening is a weakening homomorphism, we have
\begin{equation*}
S_f\circ (W_A/B)\circ W_B=S_f\circ W_{W_A(B)}\circ W_A
\end{equation*}
By \autoref{tTf:StWf_id} of \autoref{defn:esystems} it we get that
\begin{equation*}
S_f\circ W_{W_A(B)}\circ W_A=W_A.\qedhere
\end{equation*}
\end{proof}

\begin{rmk}
Note that \autoref{f:SidfWff_id} of \autoref{defn:esystems} asserts precisely
that ${(\tfid{A})}_\ast=\catid{\cat{C}/\Delta}$ for any $A:\Delta\to \Gamma$ in $\cat{C}$. In
particular, it follows that $g\circ\tfid{A}=g$ for any $g\in\thom{A}{B}$
\end{rmk}

\begin{lem}
For any $f\in\thom{A}{B}$ we have $\tfid{B}\circ f=f$.
\end{lem}

\begin{proof}
We have
\begin{equation*}
\tfid{B}\circ f=S_f(W_A(\tfid{B}))=S_f(\tfid{W_A(B)})=f.\qedhere
\end{equation*}
\end{proof}

\begin{lem}
For any $f\in\thom{A}{B}$ and $g\in\thom{B}{C}$ we have $f_\ast\circ g_\ast=(g\circ f)_\ast$. 
\end{lem}

\begin{proof}
\begin{align*}
f_\ast\circ g_\ast & = S_f\circ (W_A/B) \circ S_g \circ (W_B/C)\\
& = S_f\circ S_{W_A(g)}\circ(W_A/\ctxext{B}{W_B(C)})\circ (W_B/C) \\
& = S_{S_f(W_A(g))}\circ (S_f/W_A(W_B(C)))\circ (W_A/\ctxext{B}{W_B(C)}) \circ W_B/C \\
& = S_{S_f(W_A(g))}\circ ((S_f\circ (W_A/B) \circ W_B)/C) \\
& = S_{S_f(W_A(g))}\circ ((S_f\circ W_{W_A(B)}\circ W_A)/C) \\
& = S_{S_f(W_A(g))}\circ W_A/C \\
& = (g\circ f)_\ast.\qedhere
\end{align*}
\end{proof}

\begin{cor}
Composition is associative. Thus, we have constructed a pre-category for each
object $\Gamma$, with the same objects as $\cat{C}/\Gamma$, of which the hom-sets
are the sets $\thom{A}{B}$. 
\end{cor}

\begin{rmk}
Since any E-homomorphism is both a substitution homomorphism and a weakening
homomorphism, it follows immediately that any E-homomorphism preserves any
$f_\ast$, i.e.~for any $f\in\thom{A}{B}$ in $\mathcal{E}$ and any E-homomorphism
$F:\mathcal{E}\to\mathcal{D}$, the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\mathcal{E}/\ctxext{\Gamma}{A} \arrow[r,"F/\ctxext{\Gamma}{A}"] & \mathcal{D}/F(\ctxext{\Gamma}{A})
  \\
\mathcal{E}/\ctxext{\Gamma}{B} \arrow[u,"f_\ast"] \arrow[r,swap,"F/\ctxext{\Gamma}{B}"] & \mathcal{D}/F(\ctxext{\Gamma}{B}) \arrow[u,swap,"F(f)_\ast"]
\end{tikzcd}
\end{equation*}
commutes. This applies to weakening, substitution and pre-composition.
\end{rmk}

\begin{cor}
Suppose $F:\mathcal{E}\to\mathcal{D}$ is an E-homomorphism. Then $F$ induces
a functor $\mathbb{T}_F:\mathbb{T}_\mathcal{E}\to \mathbb{T}_\mathcal{D}$. 
\end{cor}

\subsection{Pairing and the projections}

\begin{defn}
Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. 
We define
\begin{equation*}
\tmext{x}{u}\defeq S_u(S_x(\tfid{\ctxext{A}{P}}))\in T(\ctxext{A}{P}).
\end{equation*}
\end{defn}

\begin{thm}
Let $x\in T(A)$ and $u\in T(S_x(P))$ for $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$.
Then we have
\begin{equation*}
S_{\tmext{x}{u}}=S_u\circ (S_x/P):\cat{C}/\ctxext{\Gamma}{{A}{P}}\to\cat{C}/\Gamma
\end{equation*}
\end{thm}

\begin{defn}
Let $A\in\cat{C}/\Gamma$ and $P\in\cat{C}/\ctxext{\Gamma}{A}$. We define
\begin{align*}
\cprojfstf{A}{P} & \defeq W_P(\tfid{A})\in\thom{\ctxext{A}{P}}{A}\\
\cprojsndf{A}{P} & \defeq \tfid{P}\in\thomd{\cprojfstf{A}{P}}{\ctxext{A}{P}}{P}
\end{align*}
\end{defn}

\begin{lem}
Let $x\in T(A)$ and $u\in T(S_x(P))$. then we have
\begin{align*}
S_{\tmext{x}{u}}(\cprojfstf{A}{P}) & \jdeq x\\
S_{\tmext{x}{u}}(\cprojsndf{A}{P}) & \jdeq u.
\end{align*}
Also, for $w\in T(\ctxext{A}{P})$ we have
\begin{equation*}
\tmext{S_w(\cprojfstf{A}{P})}{S_w(\cprojsndf{A}{P})} \jdeq w.
\end{equation*}
\end{lem}

\begin{lem}
For any $x\in T(A)$, $u\in T(S_x(P))$ and $v\in T(S_{\tmext{x}{u}}(Q))$ we have
\begin{equation*}
\tmext{{x}{u}}{v}\jdeq \tmext{x}{{u}{v}}\in T(\ctxext{{A}{P}}{Q}).
\end{equation*}
Also, we have
\begin{align*}
\cprojfstf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq\cprojfstf{A}{\ctxext{P}{Q}} \\
\cprojsndf{A}{P}\circ\cprojfstf{\ctxext{A}{P}}{Q} & \jdeq \cprojfstf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}} \\
\cprojsndf{\ctxext{A}{P}}{Q} & \jdeq \cprojsndf{P}{Q}\circ\cprojsndf{A}{\ctxext{P}{Q}}.
\end{align*}
\end{lem}

\begin{thm}
Let $F:\mathcal{E}\to\mathcal{D}$ be an E-homomorphism and let $x\in T(A)$ and
$u\in T(S_x(P))$. Then we have
\begin{equation*}
F(\tmext{x}{u})\jdeq \tmext{F(x)}{F(u)}.
\end{equation*}
Also, we have
\begin{align*}
F(\cprojfstf{A}{P}) & \jdeq \cprojfstf{F(A)}{F(P)} \\
F(\cprojsndf{A}{P}) & \jdeq \cprojsndf{F(A)}{F(P)}.
\end{align*}
\end{thm}

\begin{lem}
Let $f\in\thom{A}{\ctxext{B}{Q}}$ Then we have
\begin{equation*}
f \jdeq \tmext{\jcomp{A}{f}{\cprojfstf{B}{Q}}}{\jcomp{A}{f}{\cprojsndf{B}{Q}}}.
\end{equation*}
Alternatively, when $f_0\in\thom{A}{B}$ and $f_1\in\thom{A}{\jcomp{A}{f}{Q}}$,
then we have $\tmext{f_0}{f_1}\in\thom{A}{\ctxext{B}{Q}}$ and
\begin{align*}
\jcomp{A}{\tmext{f_0}{f_1}}{\cprojfstf{B}{Q}} & \jdeq f_0 \\
\jcomp{A}{\tmext{f_0}{f_1}}{\cprojsndf{B}{Q}} & \jdeq f_1.
\end{align*}
\end{lem}

\begin{thm}
Precomposition with $\cprojfstf{A}{P}$ is weakening by $P$, i.e.
\begin{equation*}
S_{\cprojfstf{A}{P}}\circ W_{\ctxext{A}{P}}\jdeq W_P.
\end{equation*}
\end{thm}

\section{Natural transformations of E-homomorphisms}
\begin{rmk}
When $\mathcal{E}$ is an E-system with empty context, then we have the category
structure on the contexts given by $\thom{\Gamma}{\Delta}= T(W_\Gamma(\Delta))$.
\end{rmk}

\begin{defn}
A \emph{natural transformation} of E-homomorphisms from $F:\mathcal{E}\to\mathcal{D}$
to $G:\mathcal{E}\to\mathcal{D}$ preserving the empty context consists of
\begin{enumerate}
\item A t-morphism $\tau0_A\in\thom{F(\Gamma)}{G(\Gamma)}$ for every $\Gamma\in\mathcal{E}$.
\item A commuting square square
\begin{equation*}
\begin{tikzcd}
F(A) \arrow[r,"\tau1_A"] \arrow[d,twoheadrightarrow] & G(A) \arrow[d,twoheadrightarrow] \\
F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] & G(\Gamma)
\end{tikzcd}
\end{equation*}
for every family $P$ over $A$ of solid arrows, such that the diagram
\begin{equation*}
\begin{tikzcd}
F(A) \arrow[r,"\tau1_A"] & G(A) \\
F(\Gamma) \arrow[r,swap,"\tau0_\Gamma"] \arrow[densely dotted,u,"F(x)"] & G(\Gamma) \arrow[densely dotted,u,swap,"G(x)"]
\end{tikzcd}
\end{equation*}
commutes for every $x\in T(A)$.
\end{enumerate}
such that
\begin{enumerate}
\item the square
\begin{equation*}
\begin{tikzcd}[column sep=huge]
W_{F(A)}(F(B)) \arrow[r,"W_{F(A)}(\tau1_B)"] \arrow[d,->>] & W_{G(A)}(G(B)) \arrow[d,->>] \\
F(A) \arrow[r,"\tau1_A"] & G(A)
\end{tikzcd}
\end{equation*}
commutes for every $A,B\in\mathcal{E}/\Gamma$. Note that 
\begin{align*}
W_{F(A)}(\tau1_B)
  & \in 
T(W_{F(A)}(W_{F(B)}(G(B))))
  \\
  & =
T(W_{W_{F(A)}(F(B))}(W_{F(A)}(G(B))))
\end{align*}
\end{enumerate}
\end{defn}

\section{Indexed E-systems}
\begin{defn}
Let $\cat{A}$ be a category. An $\cat{A}$-indexed E-system is an internal
E-system in the category $Psh(\cat{A})$.
\end{defn}

\section{Examples of adjunctions}

\begin{lem}
For each $A\in\mathcal{E}/\Gamma$, $W_A$ is left adjoint to $\Pi_A$. 
\end{lem}

\begin{proof}
The unit $\eta:1_{\cat{C}/\ctxext{\Gamma}{A}}\Rightarrow W_A\circ\Pi_A$ is given by
\begin{equation*}
\eta_P:\lam{f}{x} ev(\lambda(f),x)
\end{equation*}
$\eta_P\in\thom{P}{W_A(\Pi_A(P))}$
\end{proof}

\begin{comment}
\begin{lem}
Suppose $\cat{D}$ is a retract of a pre-category $\cat{C}$, i.e. there are functors
$S:\cat{D}\to\cat{C}$ and $R:\cat{C}\to\cat{D}$ with a natural isomorphism
$\tau:R\circ S \simeq \catid{\cat{D}}$, and suppose that $\cat{C}$ is a weakening
category. Then $\cat{D}$ is a weakening system too.
\end{lem}

\begin{proof}
Let $f:\Delta\to \Gamma$ be a morphism in $\cat{D}$. Then we define $W_f$ to be the
composition
\begin{equation*}
\begin{tikzcd}[column sep=3.2em]
\cat{D}/\Gamma
  \arrow[r,"S/\Gamma"]
  &
\cat{C}/S(\Gamma)
  \arrow[r,"W_{S(f)}"]
  &
\cat{C}/S(\Delta)
  \arrow[r,"R/S(\Delta)"]
  &
\cat{D}/R(S(\Delta))
  \arrow[r,"\tau_\Delta\circ{-}"]
  &
\cat{D}/\Delta
\end{tikzcd}
\end{equation*}
We first verify whether this assembles a pre-weakening structure on $\cat{D}$.
\begin{enumerate}
\item Let $\Delta$ be an object of $\cat{D}$. Then $W_{\catid{\Delta}}$ is the composition
\begin{equation*}
\begin{tikzcd}[column sep=3.2em]
\cat{D}/\Delta
  \arrow[r,"S/\Delta"]
  &
\cat{C}/S(\Delta)
  \arrow[r,"R/S(\Delta)"]
  &
\cat{D}/R(S(\Delta))
  \arrow[r,"{-}\circ\tau_\Delta"]
  &
\cat{D}/\Delta
\end{tikzcd}
\end{equation*}
This functor maps $g:\Delta'\to \Delta$ consecutively to $S(g)$, to $R(S(g))$, to
$\tau_\Delta\circ R(S(g))$. Note that $\tau_\Delta\circ R(S(g))=g\circ \tau_{\Delta'}$ by
naturality.
\end{enumerate}
\end{proof}
\end{comment}