### 2015-07-29 e-systems.tex
separation of sections into different files

\section{E-systems}\label{sec:esys_defn}

\subsection{Weakening systems}

\begin{defn}
A \emph{pre-weakening structure on a pre-category $\cat{F}$} consists of a functor
$W_A:\cat{F}/\Gamma\to \cat{F}/\ctxext{\Gamma}{A}$ for every $A:\ctxext{\Gamma}{A}\to \Gamma$ in $\cat{F}$ such 
that
\begin{enumerate}
\item $W_{\catid{\Gamma}}\jdeq \catid{\cat{F}/\Gamma}$ for every object $\Gamma\in\cat{F}$.
\item $W_{\ctxext{A}{P}}\jdeq W_P\circ W_A$ for every $P\in\cat{F}/\ctxext{\Gamma}{A}$ and $A\in\cat{F}/\Gamma$.
\item $W_A$ preserves the final object, i.e.~$W_A(\catid{\Gamma})\jdeq \catid{\ctxext{\Gamma}{A}}$.
\end{enumerate}
A \emph{pre-weakening system} is a pre-category together with a pre-weakening structure.
\end{defn}

\begin{defn}
A \emph{pre-weakening homomorphism} $F:\cat{F}\to\cat{D}$ between pre-weakening systems
is a functor $F:\cat{F}\to\cat{D}$ such that the square
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{F}/\ctxext{\Gamma}{A}
  \arrow[r,"F/\ctxext{\Gamma}{A}"]
  &
\cat{D}/F(\ctxext{\Gamma}{A})
  \\
\cat{F}/\Gamma
  \arrow[u,"W_A"]
  \arrow[r,swap,"F/\Gamma"]
  &
\cat{D}/F(\Gamma)
  \arrow[u,swap,"W_{F(A)}"]
\end{tikzcd}
\end{equation*}
commutes for any $A\in\cat{F}/\Gamma$.
\end{defn}

\begin{lem}
Let $\cat{F}$ be a pre-weakening system. Then for any object $\Gamma$ of $\cat{F}$,
the slice pre-category $\cat{F}/\Gamma$ is also a pre-weakening system, with the weakening
structure given by $W(\cat{F}/\Gamma)_P\jdeq W(\cat{F})_P$ for each $P\in\cat{F}/\ctxext{\Gamma}{A}$
and each $A\in\cat{F}/\Gamma$.
\end{lem}

\begin{defn}
A \emph{weakening system} is a pre-weakening system for which every $W_f$ is a
pre-weakening homomorphism. A \emph{weakening homomorphism} is a pre-weakening 
homomorphism between weakening systems.
\end{defn}

\begin{rmk}
In other words, a weakening system is a pre-weakening system with a weakening
structure $W$ for which the square
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{F}/\ctxext{{\Gamma}{B}}{Q}
  \arrow[r,"W_A/\ctxext{B}{Q}"]
  &
\cat{F}/\ctxext{{\Gamma}{A}}{W_A(\ctxext{B}{Q})}
  \\
\cat{F}/\ctxext{\Gamma}{B}
  \arrow[u,"W_Q"]
  \arrow[r,swap,"W_A/B"]
  &
\cat{F}/\ctxext{{\Gamma}{A}}{W_A(B)}
  \arrow[u,swap,"W_{W_A(Q)}"]
\end{tikzcd}
\end{equation*}
for each $A,B\in\cat{F}/\Gamma$ and $Q:\cat{F}/\ctxext{\Gamma}{B}$. The objects part of this 
property asserts that for any $k\in \cat{F}/\greek{E}$, the dotted arrows in the diagram
\begin{equation*}
\begin{tikzcd}[column sep=tiny]
\bullet \arrow[dr,densely dotted,swap,near start,"(W_A/\ctxext{B}{Q})(W_Q(R))"] & & \bullet \arrow[dl,densely dotted,near start,"W_{(W_A/B)(Q)}((W_A/B)(R))"] & & & & & & & & & & \bullet \arrow[dl,"W_Q(R)"] \\
& \bullet \arrow[dr,swap,"(W_A/B)(Q)"] & & \bullet \arrow[dl,"(W_A/B)(R)"] & & & & & & & & \bullet \arrow[dr,swap,"Q"] & & \bullet \arrow[dl,"R"] \\
& & \bullet \arrow[d,swap,"W_A(B)"] & & & & & & & & & & \ctxext{\Gamma}{B} \arrow[d,"B"] \\
& & \ctxext{\Gamma}{A} \arrow[rrrrrrrrrr,swap,"A"] & & & & & & & & & & \Gamma
\end{tikzcd}
\end{equation*}
are equal.

A useful special case of this property is where $B\jdeq \catid{\Gamma}$. Thus, if $W$ is
a weakening system, then the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{F}/\ctxext{\Gamma}{C}
  \arrow[r,"W_A/C"]
  &
\cat{F}/\ctxext{{\Gamma}{A}}{W_A(C)}
  \\
\cat{F}/\Gamma
  \arrow[u,"W_C"]
  \arrow[r,swap,"W_A"]
  &
\cat{F}/\ctxext{\Gamma}{A}
  \arrow[u,swap,"W_{W_A(C)}"]
\end{tikzcd}
\end{equation*}
commutes for every $A,C\in\cat{F}/\Gamma$. In particular, we see
that $W_A(W_C(D))\jdeq W_{W_A(C)}(W_A(D))$ for any $D\in\cat{F}/\Gamma$,
i.e.~that weakening is a self-distributive operation. 
\end{rmk}

\begin{cor}
For any object $\Gamma$ of a weakening system $\cat{F}$, the slice pre-category
$\cat{F}/\Gamma$ is a weakening system.
\end{cor}

\begin{eg}
If the underlying pre-category is a monoid $M$, the axioms of a weakening system
obtain the following form: There is a group homomorphism $W_x:M\to M$ for
every $x\in M$, in such a way that $W_e\jdeq \catid{M}$ and $W_{xy}\jdeq W_y\circ W_x$
for every $x,y\in M$, and for any $x,y\in M$ the diagram
\begin{equation*}
\begin{tikzcd}
M \arrow[r,"W_y"]
  &
M 
  \\
M \arrow[u,"W_x"]
  \arrow[r,swap,"W_y"]
  &
M \arrow[u,swap,"W_{W_y(x)}"]
\end{tikzcd}
\end{equation*}
is required to commute. Thus, a weakening system on a monoid $M$ consists of
a homomorphism $W:\op M\to\mathrm{Aut}(M)$ such that $W_x\circ W_y\jdeq 
W_{W_x(y)}\circ W_x$. The latter requirement is satisfied if $y\cdot x\jdeq x\cdot W_x(y)$
for every $x,y\in M$. Group conjugation forms a weakening structure on any group.
\end{eg}

\begin{comment}
\begin{lem}
For any group, conjugation is a weakening structure. Thus, any group is
canonically a weakening system.
\end{lem}

\begin{proof}
Note that $e^{-1}xe\jdeq x$ and $(xy)^{-1}z(xy)\jdeq y^{-1}(x^{-1}zx)y$ for any $x,y,z\in G$.
Also, the diagram
\begin{equation*}
\begin{tikzcd}
G \arrow[r,"\mathrm{conj}_y"]
  &
G 
  \\
G \arrow[u,"\mathrm{conj}_x"]
  \arrow[r,swap,"\mathrm{conj}_y"]
  &
G \arrow[u,swap,"\mathrm{conj}_{y^{-1}xy}"]
\end{tikzcd}
\end{equation*}
commutes.
\end{proof}
\end{comment}

\begin{comment}
\begin{rmk}
Similarly, every groupoid has a canonical weakening structure given by
conjugation.
\end{rmk}
\end{comment}

\begin{comment}
\begin{lem}
Let $G$ be a graph, and let $\cat{F}_G$ be the free pre-category generated by
$G$. A pre-weakening structure on $\cat{F}_G$ is determined by functors
$W_e:\cat{F}_G/\Gamma\to\cat{F}_G/\Delta$ for every edge $e:\Delta\to \Gamma$ in $G$, with the
property that $W_e(\catid{\Gamma})\jdeq \catid{\Delta}$.

A weakening structure on $\cat{F}_G$ is
a pre-weakening structure on $\cat{F}_G$ such that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{F}_G/\greek{Z} \arrow[r,"W_e/(g\circ h)"] & \cat{F}_G/\mathrm{dom}(W_e(g\circ h)) \\
\cat{F}_G/\greek{E} \arrow[u,"W_h"] \arrow[r,swap,"W_e/g"] & \cat{F}_G/\mathrm{dom}(W_e(g)) \arrow[u,swap,"W_{W_e(h)}"]
\end{tikzcd}
\end{equation*}
commutes for every arrow $g:\greek{E}\to \Gamma$ in $\cat{F}_G$ and every two edges $e:\Delta\to \Gamma$ and
$h:\greek{Z}\to \greek{E}$ in $G$.
\end{lem}

\begin{eg}
Let $G$ be the graph consisting of one vertex $\ast$ and one edge $e:\ast\to\ast$.
Then $\cat{F}_G/\ast\jdeq \N$. Then the only weakening structure on $\cat{F}/G$ is
given by $W_e\jdeq \catid{\N}$. The reason is that 
\end{eg}

\begin{lem}
Let $P$ be a pre-order. A pre-weakening structure on $P$ is determined by order
preserving maps $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$ satisfying
$W_{a,b}(b)\jdeq a$, for every $a\leq b$, where
${\downarrow}(x)$ is the set $\{y\in P\mid y\leq x\}$.

A weakening structure on $P$ is a pre-weakening structure on $P$ such that the
diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
{\downarrow}(c') \arrow[r,"W_{a,b}|_{{\downarrow}(c')}"] & {\downarrow}(W_{a,b}(c')) \\
{\downarrow}(c) \arrow[u,"W_{c',c}"] \arrow[r,swap,"W_{a,b}|_{{\downarrow}(c)}"] & {\downarrow}(W_{a,b}(c)) \arrow[u,swap,"W_{W_{a,b}(c'),W_{a,b}(c)}"]
\end{tikzcd}
\end{equation*}
for any $a\leq b$ and $c'\leq c\leq b$.
\end{lem}

\begin{eg}
In a poset with meets, the functions $W_{a,b}:{\downarrow}(b)\to{\downarrow}(a)$
given by $W_{a,b}(x)\defeq a\land x$ assemble a weakening structure, because for
any $a\leq b$ and $x,c'\leq c\leq b$ we have $a\land(c'\land x)\jdeq (a\land c')\land
(a\land x)$.  
\end{eg}

\begin{eg}
Consider the poset $(\N,\geq)$, which has a morphism from $a$ to $b$ if $a\geq b$.
On this poset, we may consider the weakening structure defined by $W_{a,b}(x)\jdeq
a+(x-b)$.
\end{eg}
\end{comment}

\subsection{Weakening systems with term structure}
\begin{defn}
A \emph{category with a term structure} is a pre-category $\cat{F}$ with a set $T(A)$ for every
$A\in\cat{F}/\Gamma$ for every $\Gamma\in\cat{F}$. A \emph{functor with term structure} from $\cat{F}$ to $\cat{D}$
is a functor $F:\cat{F}\to\cat{D}$ with a function $T(A)\to T(F(A))$ for
every $A\in\cat{F}/\Gamma$, for every $\Gamma\in\cat{F}$.
\end{defn}

\begin{defn}
Suppose $\cat{F}$ is a pre-category with term structure, and $\Gamma$ is an object of $\cat{F}$. 
Then $\cat{F}/\Gamma$ is a pre-category with term structure, where $T_{\cat{F}/\Gamma}(A)\jdeq T_\cat{F}(A)$.
\end{defn}

\begin{defn}
A \emph{weakening system with term term structure} is the same thing as a weakening system,
except that the underlying pre-category and the functors $W_A$ are required to
be functors with term term structure. The equalities are also required to hold on the term structure.
\end{defn}

\begin{cor}
For any object $\Gamma$ of a weakening system with term structure, the slice pre-category
$\cat{F}/\Gamma$ is a weakening system with term structure.
\end{cor}

\begin{eg}
%For any group $G$ with weakening structure $W:\op{G}\to\mathrm{Aut}(G)$, a term 
%structure is determined by a presheaf on $G/\ker(W)$.

Let $G$ be a group, with the weakening structure given by conjugation. A term
structure on $G$ is given by a set $X$ with a right $G$-action $\mu:\op G\times X\to X$.
We take $T(a)\defeq X$ for every $a\in G$.
\end{eg}

\subsection{Projection systems}
\begin{defn}
A \emph{pre-projection system} is a pre-weakening system with term structure with an element
$\tfid{A}\in T(W_A(A))$ for every $A\in\cat{F}/\Gamma$ and $\Gamma\in\cat{F}$. 
\end{defn}

\begin{defn}
A \emph{pre-projection homomorphism} $F:\cat{F}\to\cat{D}$ is a pre-weakening homomorphism for which
\begin{equation*}
F(\tfid{A})\jdeq \tfid{F(A)}
\end{equation*}
for every $A\in\cat{F}/\Gamma$ and $\Gamma\in\cat{F}$.
\end{defn}

\begin{lem}
For any object $\Gamma$ in a pre-projection system $\cat{F}$, the slice pre-category
$\cat{F}/\Gamma$ is a pre-projection system.
\end{lem}

\begin{proof}
Straightforward.
\end{proof}

\begin{defn}
A \emph{projection system} is a pre-projection system for which every $W_A$ is
a pre-projection homomorphism. A \emph{projection homomorphism} is a pre-projection homomorphism
between projection systems.
\end{defn}

\begin{cor}
For any object $\Gamma$ of a projection system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
is a projection system.
\end{cor}

%\begin{eg}
%Consider a group $G$ with a right $G$-set $X$. A projection structure on $(G,X)$
%consists of an element $\tfid{a}\in X$ for every $a\in G$, so that $\tfid{a}\cdot b=
%\tfid{b^{-1}ab}$ for every $a,b\in G$. 
%For instance, this is the case if we take $X$ to be a pointed set with $\ast\in X$,
%and take $\tfid{a}:=\ast$ for all $a\in G$.
%\end{eg}

\subsection{Substitution systems}

\begin{defn}
A \emph{pre-substitution structure} on a pre-category with term structure $\cat{F}$ consists
of a functor with term structure $S_x:\cat{F}/\ctxext{\Gamma}{A}\to\cat{F}/\Gamma$ for every $x\in T(A)$ and
$A\in\cat{F}/\Gamma$, such that $S_x(\catid{\ctxext{\Gamma}{A}})\jdeq\catid{\Gamma}$.

A \emph{pre-substitution system} is a pre-category with term structure together with a
pre-substitution structure. 
\end{defn}

\begin{defn}
A \emph{pre-substitution homomorphism} $F:\cat{F}\to\cat{D}$ is a functor with term structure for
which the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{F}/\ctxext{\Gamma}{A}
  \arrow[r,"F/\ctxext{\Gamma}{A}"]
  \arrow[d,swap,"S_x"]
  &
\cat{D}/F(\ctxext{\Gamma}{A})
  \arrow[d,"S_{F(x)}"]
  \\
\cat{F}/\Gamma
  \arrow[r,swap,"F/\Gamma"]
  &
\cat{D}/F(\Gamma)
\end{tikzcd}
\end{equation*}
commutes for every $x\in T(A)$ and $A\in\cat{F}/\Gamma$.
\end{defn}

\begin{lem}
For any object $\Gamma$ of a pre-substitution system $\cat{F}$, the slice pre-category
$\cat{F}/\Gamma	$ is a pre-substitution system.
\end{lem}

\begin{defn}
A \emph{substitution system} is a pre-substitution system for which each
$S_t$ is a pre-substitution homomorphism. A \emph{substitution homomorphism} is a
pre-substitution homomorphism between substitution systems.
\end{defn}

\begin{cor}
For any object $\Gamma$ of a substitution system, the slice pre-category $\cat{F}/\Gamma$
is a substitution system.
\end{cor}

\begin{rmk}
The condition that every $S_x$ is a substitution homomorphism, asserts that
the diagram
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{F}/\ctxext{{{\Gamma}{A}}{P}}{Q}
  \arrow[r,"S_x/\ctxext{P}{Q}"]
  \arrow[d,swap,"S_y"]
  &
\cat{F}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
  \arrow[d,"S_{S_x(y)}"]
  \\
\cat{F}/\ctxext{{\Gamma}{A}}{P}
  \arrow[r,swap,"S_x/P"]
  &
\cat{F}/\ctxext{\Gamma}{S_x(P)}
\end{tikzcd}
\end{equation*}
commutes for every $y\in T(Q)$.
\end{rmk}

\begin{eg}
Let $G$ be a group and let $H$ be a set, forming a category with term 
structure of which $T(a)\jdeq H$ for every $a\in G$. A substitution
system on $(G,H)$ consists of a function $\sigma_G : H\times G\to G$ and
a function $\sigma_H:H\times H\to H$, for which the diagrams
\begin{equation*}
\begin{tikzcd}[column sep=large]
G \arrow[r,"\sigma_G(h)"]
  \arrow[d,swap,"\sigma_G(k)"]
  &
G \arrow[d,"{\sigma_G(\sigma_H(h,k))}"]
  \\
G \arrow[r,swap,"\sigma_G(h)"] 
& G
\end{tikzcd}
\end{equation*}
and
\begin{equation*}
\begin{tikzcd}[column sep=large]
H \arrow[r,"\sigma_H(h)"]
  \arrow[d,swap,"\sigma_H(k)"]
  &
H \arrow[d,"{\sigma_H(\sigma_H(h,k))}"]
  \\
H \arrow[r,swap,"\sigma_H(h)"]
  &
H
\end{tikzcd}
\end{equation*}
commute for every $h,k\in H$. 
\end{eg}

\subsection{The definition of E-systems}

Although it is now possible to state the definition of an E-system without
any difficulty, it shall be convenient later to also have a category of
pre-E-systems at our disposal. Pre-E-systems shall by default be non-unital,
i.e.~they do not come with a pre-projection system, while E-systems shall
always be assumed to be unital. In the present context, we shall not need
unital pre-E-systems or non-unital E-systems.

\begin{defn}
A \emph{pre-E-system} is a pre-category $\cat{F}$ with term structure
with a pre-weakening structure $W$, a pre-substitution structure $S$ and a terminal
object. A \emph{unital pre-E-system}, or simply \emph{pre-E-system}, is a non-unital pre-E-system
which also has the structure of a pre-projection system. 
\end{defn}

\begin{defn}
A \emph{pre-E-homomorphism} from $\mathbb{E}\to\mathbb{D}$ is a functor
$H:\mathbb{C}_\mathbb{E}\to\mathbb{C}_\mathbb{D}$ between the underlying categories
with term structure, which is a pre-weakening homomorphism, a pre-substitution
homomorphism, and which preserves the terminal object.

A \emph{unital} pre-E-homomorphism is a non-unital pre-E-homomorphism which is 
also a pre-projection homomorphism. 
\end{defn}

\begin{defn}
\label{defn:esystems}
An \emph{E-system} is a unital pre-E-system in which 
\begin{enumerate}
\item each $W_A$ is a unital pre-E-homomorphism,
\item each $S_x$ is a unital pre-E-homomorphism,
\item \label{tTf:StWf_id} weakened families are constant, i.e.~$S_x\circ W_A\jdeq \catid{\cat{F}/\Gamma}$ for any $x\in T(A)$ and $A\in\cat{F}/\Gamma$,
\item $S_x(\tfid{A})\jdeq x$ for any $x\in T(A)$ and $A\in\cat{F}/\Gamma$, and
\item \label{f:SidfWff_id} $S_{\tfid{A}}\circ W_A/A\jdeq \catid{\cat{F}/\ctxext{\Gamma}{A}}$ for any $A\in\cat{F}/\Gamma$.
\end{enumerate}
An \emph{E-homomorphism} $H:\mathbb{E}\to\mathbb{D}$ is a functor from
$\cat{F}_{\mathbb{E}}$ to $\cat{F}_{\mathbb{D}}$ which is both a projection
homomorphism and a substitution homomorphism.
\end{defn}

\begin{rmk}
The condition that each $W_A$ is a substitution homomorphism asserts that
the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{F}/\ctxext{{\Gamma}{B}}{Q}
  \arrow[r,"W_A/\ctxext{B}{Q}"]
  \arrow[d,swap,"S_y"]
  &
\cat{F}/\ctxext{{{\Gamma}{A}}{W_A(B)}}{W_A(Q)}
  \arrow[d,"S_{W_A(y)}"]
  \\
\cat{F}/\ctxext{\Gamma}{B}
  \arrow[r,swap,"W_A/B"]
  &
\cat{F}/\ctxext{{\Gamma}{A}}{W_A(B)}
\end{tikzcd}
\end{equation*}
of functors with term structure commutes for every 
$Q\in\cat{F}/\ctxext{\Gamma}{B}$, $B\in\cat{F}/\Gamma$ and each $y\in T(Q)$.

Likewise, the condition that each $S_x$ is a weakening homomorphism
asserts that the diagram
\begin{equation*}
\begin{tikzcd}[column sep=huge]
\cat{F}/\ctxext{{\Gamma}{A}}{P}
  \arrow[r,"S_x/P"]
  \arrow[d,swap,"W_Q"]
  &
\cat{F}/\ctxext{\Gamma}{S_x(P)}
  \arrow[d,"W_{S_x(Q)}"]
  \\
\cat{F}/\ctxext{{{\Gamma}{A}}{P}}{Q}
  \arrow[r,swap,"S_x/\ctxext{P}{Q}"]
  &
\cat{F}/\ctxext{{\Gamma}{S_x(P)}}{S_x(Q)}
\end{tikzcd}
\end{equation*}
of functors with term structure commutes for every 
$Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$.
\end{rmk}

\begin{cor}
For any object $\Gamma$ of an E-system $\cat{F}$, the slice pre-category $\cat{F}/\Gamma$
is an E-system.
\end{cor}

\begin{comment}
\begin{rmk}[Needs more precision]
Consider the category $E$ generated by the graph consisting of the objects of 
the underlying category of an $E$-system $\mathcal{E}$ as edges, and the morphisms
$S_t:X\to Y$ and $W_A:Y\to X$ as edges, subject to the axioms of an $E$-systems.

Any morphism of $E$ can be rewritten in a unique way
as a composition $\tilde S\circ \tilde W$, where $\tilde S$ is a morphism in
substitution-normal form, $\tilde W$ is in weakening normal form, every term
expression is in normal form, and where
each occurence of $S_x\circ W_A$ and $S_{\tfid{A}}\circ W_A/A$ is eliminated.
\end{rmk}
\end{comment}