### 2015-08-16 weakeningcats.tex

\begin{comment}
\subsection{Quotients of E-systems}

\begin{defn}
A regular congruence relation $\sim$ on a category $\cat{F}$ with term structure,
is a triple $(\sim_c,\sim_f,\sim_t)$ consisting of
\begin{enumerate}
\item an equivalence relation $\sim_c$ on the set of objects of $\cat{F}$,
\item an equivalence relation $\sim_f$ on the set of morphisms of $\cat{F}$,
\item an equivalence relation $\sim_t$ on the set of terms of $\cat{F}$,
\end{enumerate}
which are compatible with $\mathrm{cod}$, $\mathrm{dom}$, $\catid{}$, $\circ$
and $\ebd$
\end{defn}
\end{comment}