### 2015-10-01 main.tex

\documentclass{article}

\input{Preamble/preamble-articles}

%\includeonly{Algebras/article-models-cat}

\newcommand{\thetitle}{An essentially algebraic formulation of dependent type theory}
\newcommand{\theauthor}{Egbert Rijke}
\hypersetup{pdftitle={\thetitle},pdfauthor={\theauthor},pdfhighlight={/I}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\thetitle}
\author{\theauthor\footnote{{\sc Carnegie Mellon University}, department of
philosophy; \url{erijke@andrew.cmu.edu}}}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We present an algebraic formulation of dependent type theory with
three sorts for contexts, families and terms, with finitely many operations
between them and finitely many equality axioms. Algebras for this theory are
called E-systems. A distinctive feature of the dependent type theory of
E-systems is that it takes `telescopic families' as the primitive notion of
families.

We then compare Voevodsky's B-systems, Dybjer's categories
with families and Jacobs' comprehension categories with E-systems. There is
a full and faithful functor from B-systems to E-systems, of which the image is
the category of stratified E-systems. To have a functor from categories with
families to E-systems, we need to impose an additonal coherence condition, which
makes sure that the substitution operation of a category with families extends 
to telescopic families. 
\end{abstract}

\tableofcontents

\input{introduction}

%\tableofcontents

%\input{Introduction/article-models-intro}

%\part{Type theory}

\begin{comment}
In this part we develop type theory from the ground up. We start with a type
theory without any of the basic constructors. This is the theory of contexts
families and terms which has the basic operations of extension, weakening,
substitution and identity terms. Type theory before type constructors has not
been studied very much. Dependent product types or even universes tend to make
an early appearance in just about any presentation of type theory.
A noteable exception is the theory of categories with families of Dybjer
in \cite{Dybjer1996}, which has been elaborated on further by Dybjer and
Clairambault in \cite{DybjerClairambault2011} and in unpublished work by
Awodey \cite{Awodey2013} on natural models of type theory,
which makes the connection between categories with families and representable
transformations of presheaves. In the way type theory is presented by the
Univalent Foundations Project in \cite{TheBook}, which seems to have won the race of introducing universes
as early as possible hands down, it seems entirely unfeasible to study type
theory without type constructors. Also, the 
proof-assistant \Coq\ {\color{red}(and \Agda too?)} has universes and dependent product and pair types 
built-in, making it impossible to study type theory without type constructors in 
that environment.

Nevertheless, type theory without type constructors has received some attention
contributors to the Univalent Foundations Program recently. We name two
further investigations on this topic, other than the mentioned work by Awodey.
In \cite{Garner2014}, Garner describes the combinatorial structure
of the type operations of the weakening, substitution and projection monads
(their projections are our identity terms) and suggests lots of further research
that can be done on type theory without constructors. Also, Joyal has presented
his theory of tribes, which is a categorical explanation of type theory
without type constructors.

After we have described the E-system, which is the flavor of type theory without
type constructors presented in \autoref{tt}, we will demonstrate in \autoref{ttderived}
that the theory gives rise to a rich categorical structure. The notions and
properties we derive here will be essential for the further work on internal
models for E-systems, presented in \autoref{part:models}.
\end{comment}

\begin{comment}
\section{Introduction}
In this article we define the essentially algebraic objects which correspond to dependent
type theories, which are closely related to Voevodsky's B-systems. This provides
a new axiomatization of dependent type theory, with the following properties:
\begin{itemize}
\item The meta-theory should not require anything more than rules for handling
inferences. In particular, natural numbers, or an infinite set of variables 
should not be required.
\item Finitely many sorts. The basic ingredients are contexts, families and
terms. Thus there will be 3 sorts.
\item The theory is algebraic and the operations on it are homomorphisms. From
the requirement that type theoretic operations must be homomorphisms, we can
read off which judgmental equalities to require require.
\item Finite set of rules.
\item The theory is invariant under slicing in a canonical way.
\end{itemize}

We build
dependent type theory by imposing several structures of increasing complexities
on top of each other. We begin by the theory of extension, which incorporates
a context extension and a family extension operation to the theory which merely
has three sorts. On top of the theory of extension we can formulate the theories
of weakening, substitution and extension-units (i.e.~the empty context and families).
The theory of weakening can be enriched with units playing the role of identity
morphisms, which gives a theory of projections. The three resulting theories
can then be combined into dependent type theory. Thus, the theory of dependent
types is outlined in the following diagram.

\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=0em]
{} \& \textbf{Extension} \ar[very thick,-stealth]{dr} \ar[very thick,-stealth]{d} \ar[very thick,-stealth]{dl}
  \\
\makebox[.7cm]{\textbf{Weakening}} \ar[very thick,-stealth]{d} \& \textbf{Substitution} \ar[very thick,-stealth]{dd} \& \makebox[3cm]{\textbf{Empty context and families}} \ar[very thick,-stealth]{ddl}
\\
\makebox[.7cm]{\textbf{Projection}} \ar[very thick,-stealth]{dr}
\\
{} \& \textbf{Dependent Type Theory}
\end{tikzcd}
\end{equation*}

We give two presentations of the theory simultaneously. First, we directly
define the algebras for the corresponding theories in an arbitrary category
with finite limits and second, we describe the theory in terms of inference
rules.

\subsection{Categories with finite limits}
For the semantics of the theory of E-systems, we will assume that $\cat{C}$ is a 
category with finite limits and
whenever we write a pullback, we assume that it is chosen. Recall that for
any morphism $f:A\to B$ in a category $\cat{C}$ with chosen pullbacks, there
is a functor
\begin{equation*}
f^\ast : \cat{C}/B\to\cat{C}/A.
\end{equation*}
As usual, when $g:X\to B$ is a morphism, we will write $f^\ast(X)$ for the
domain of $f^\ast(g)$. When there is more than one morphism $X\to B$ involved,
as will be the case below, we will write $\pullback{A}{X}{f}{g}$. The projections
will be written as $\pullbackpr{1}{f}{g}$ and $\pullbackpr{2}{f}{g}$. So in this notation, a
typical pullback diagram has the following form:
\begin{equation*}
\begin{tikzcd}[column sep=large]
\pullback{A}{X}{f}{g}
  \ar{r}{\pullbackpr{1}{f}{g}}
  \ar{d}[swap]{\pullbackpr{2}{f}{g}}
  &
A \ar{d}{f}
  \\
X \ar{r}[swap]{g}
  &
B
\end{tikzcd}
\end{equation*}
Furthermore, given $h:Y\to A$ and $k:Y\to B$ such that $h\circ f=g\circ k$, we
denote the arrow $Y\to\pullback{A}{X}{f}{g}$ given by the universal property of
pullbacks, by $[h,k]$. Also, when we have a commutative diagram of the form
\begin{equation*}
\begin{tikzcd}
A \ar{r}{f}
  \ar{d}{a}
  &
X \ar{d}
  & 
B \ar{l}[swap]{g}
  \ar{d}{b}
  \\
A'
  \ar{r}[swap]{f'}
  &
X'
  &
B'
  \ar{l}{g'}
\end{tikzcd}
\end{equation*}
we will denote the unique map from $\pullback{A}{B}{f}{g}$ to $\pullback{A'}{B'}{f'}{g'}$
such that the diagram
\begin{equation*}
\begin{tikzcd}
  {}
  & 
\pullback{A'}{B'}{f'}{g'}
  \ar{dd}
  \ar{rr}
  &
  &
B'
  \ar{dd}{g'}
  \\
\pullback{A}{B}{f}{g}
  \ar{dd}
  \ar[crossing over]{rr}
  \ar[dotted]{ur}{\pullback{a}{b}{f'}{g'}}
  &
  &
B \ar{ur}{b}
  \\
  {}
  &
A'
  \ar{rr}
  &
  &
X'
  \\
A \ar{rr}[swap]{f}
  \ar{ur}{a}
  &
  &
X \ar[crossing over,leftarrow]{uu}[near end,swap]{g}
  \ar{ur}
\end{tikzcd}
\end{equation*}
commutes, by $\pullback{a}{b}{f'}{g'}$. In the current work, we shall
write $A\times B$ for the pullback of $A\rightarrow 1\leftarrow B$, and
$\pi_1$ and $\pi_2$ for its projections (thus, no separate choice of
cartesian products is made).

\subsection{Inference rules}
The syntactic theory of contexts, families and terms shall be described by means
of a declaration of valid inference rules. Although their appearance suggests 
otherwise, the notion of inference rule is ultimately informal, in the sense
that we state the explanations and possibilities with inference rules in
natural language. One may attempt to study systems of inference rules within
other systems to study the applications of the various concepts involved, but
this study will not make the basic notion of inference rules more formal. We do
not see this as a problem, as systems of inference rules have been implemented
in proof assistant software in which one essentially only manipulates
judgments. In this paper we use inference rules as an alternative presentation
of the theory, which some readers may prefer.

An inference rule is a finite list $\mathcal{H}_1,\ldots,\mathcal{H}_n;
\mathcal{J}$ of expressions called judgments. Usually we represent an inference
rule in the following form
\begin{equation*}
\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
\end{equation*}
The judgments $\mathcal{H}_1$, \ldots, $\mathcal{H}_n$ are called the
hypotheses and the judgment $\mathcal{J}$ is the conclusion of the inference
rule.

An inference rule is called \emph{valid} in either of the following cases:
\begin{description}
\item[(Axiom)] It is declared valid as an axiom;
\item[(Projection)] The inference rule
\begin{equation*}
\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{H}_i}
\end{equation*}
is a valid inference rule for any $1\leq i\leq n$;
\item[(Composition)] If
\begin{equation*}
\inference{\mathcal{I}_{1k}\quad\cdots\quad\mathcal{I}_{m_kk}}{\mathcal{H}_k}
\qquad\text{and}\qquad
\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{J}}
\end{equation*}
are valid inference rules for $1\leq k\leq n$, then
\begin{equation*}
\inference{\mathcal{I}_{11}\quad\cdots\quad\mathcal{I}_{m_11}\quad\cdots\quad\mathcal{I}_{1n}\quad\cdots\quad\mathcal{I}_{m_nn}}{\mathcal{J}}
\end{equation*}
is also a valid inference rule;
\item[(Cut elimination)] If
\begin{equation*}
\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_n}{\mathcal{H}_{n+1}}
\qquad
\text{and}
\qquad
\inference{\mathcal{H}_1\quad\cdots\quad\mathcal{H}_{n+m+1}}{\mathcal{J}}
\end{equation*}
are valid inference rules, then
\begin{equation*}
\inference{\mathcal{H}_1\quad\cdots\quad\hat{\mathcal{H}}_{n+1}\quad\cdots\quad\mathcal{H}_{n+m+1}}{\mathcal{J}}
\end{equation*}
where the hypothesis $\mathcal{H}_{n+1}$ is omitted, is a valid inference rule.
\end{description}
In categories with finite limits, valid inference rules will always be 
interpreted by  arrows.

\input{fundamental_structure}

\input{extension}

\input{weakening}

\input{projections}

\input{substitution}

\input{empty}

\input{dependent_type_theory}

\section{The relation between E-systems and B-systems}

In this section we answer the following questions posed by Voevodsky in private
correspondence.
\begin{itemize}
\item Is the construction that takes a (unital) B-system to an E-system a functor?
\item Is this functor fully faithful?
\item Is there a simple way to characterize the E-systems that arise from B-systems (the image of this functor on objects)?
\item Construct an example of an E-system that does not arise from a B-system.
\end{itemize}

\subsection{The functor from B-systems to E-systems}
Consider a B-system $\mathbb{B}$. We will construct an E-system $E(\mathbb{B})$
in the category of sets. This E-systems consists of sets $C(\mathbb{B})$,
$F(\mathbb{B})$ and $T(\mathbb{B})$ for the sorts of contexts, families and
terms respectively, with functions $\eft:F(\mathbb{B})\to C(\mathbb{B})$ and
$\ebd:T(\mathbb{B})\to T(\mathbb{B})$ and the operations of extension,
weakening, units, substitution and empty contexts and families.

The set $C(\mathbb{B})$ consists of pairs $(n,X)$, where $n\in\mathbb{N}$ and
$X\in B_n$. The set $F(\mathbb{B})$ consists of quadruples $(n,X,m,Y)$, where
$n,m\in\mathbb{N}$ such that $n<m$, $X\in B_n$ and $Y\in B_{m}$ such that
$ft^{m-n}:=ft^{n-1}\circ\cdots\circ ft^m(Y)=X$.

%\include{Derivations/article-models-ttderived}

%\part{Internal Models}\label{part:models}

%\include{Internal_models/article-models-ttinternal}

%\part{Type constructors}

%\include{Dependent_function_types/article-models-ttdp}

%\include{Universes/article-models-ttuniverses}

%\include{Junk/article-models-ttind}

%\part{Categorical semantics}

%\input{Algebras/article-models-cat}

\part{The new E-systems}
\end{comment}

\input{e-systems}

\input{internal_homs}

\input{compared}

\input{theories}

\begin{comment}
\section{Type constructors in E-systems}

\begin{defn}
An E-system $\mathbb{E}$ is said to have $\Pi$-types, if 
\begin{enumerate}
\item for any 
$A\in\cat{F}/\Gamma$, there is an E-homomorphism
\begin{equation*}
\Pi_A : \mathbb{E}/\ctxext{\Gamma}{A}\to\mathbb{E}/\Gamma
\end{equation*}
For terms $g:Q$ in context $\ctxext{{\Gamma}{A}}{P}$, we will usually write
$\lambda_{A}(g)$ for the term $\Pi_A(g)$ of $\Pi_A(Q)$ in context
$\ctxext{\Gamma}{\Pi_A(P)}$. Note that when $P$ is empty, then we retrieve the
usual lambda-abstraction.
\item for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$, 
there is a term 
\begin{equation*}
\tfev^{A,Q} : \ctxwk{\Pi_A(Q)}{{\Pi_A(P)}{Q}}
\end{equation*}
in context $\ctxext{{{{\Gamma}{\Pi_A(P)}}{\Pi_A(Q)}}{\ctxwk{\Pi_A(\ctxext{P}{Q})}{A}}}{\ctxwk{\Pi_A(\ctxext{P}{Q})}{P}}$.
Note that for each $g:\Pi_A(Q)$, we obtain a term $\subst{g}{\tfev^{A,Q}}: \ctxwk{\Pi_A(P)}{Q}$ in
context $\ctxext{{\Gamma}{\Pi_A(P)}}{\ctxwk{\Pi_A(P)}{A}}$.
\item for any $A\in\cat{F}/\Gamma$ there is a term
\begin{equation*}
{\tfev^{A,P}} : {\ctxwk{\Pi_A(P)}{P}}
\end{equation*}
in context ${\ctxext{{\Gamma}{\Pi_A(P)}}{\ctxwk{\Pi_A(P)}{A}}}$
\end{enumerate}
satisfying
\begin{enumerate}
\item 
\begin{enumerate}
\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ one has
\begin{equation*}
\Pi_{\ctxext{A}{P}} \jdeq \Pi_A\circ \Pi_P
\end{equation*}
\item For any $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$, one has
\begin{equation*}
{\evtm{\ctxext{A}{P}}{Q}}
  \jdeq
{\subst{\evtm{A}{\mprd{P}{Q}}}{\ctxwk{\mprd{A}{\mprd{P}{Q}}}{\evtm{P}{Q}}}}
\end{equation*}
\end{enumerate}
\item 
\begin{enumerate}
\item For any $\Gamma\in\mathbb{E}$, one has
\begin{equation*}
\Pi_{\catid{\Gamma}}\jdeq \catid{\mathbb{E}/\Gamma}.
\end{equation*}
\item For any $A\in\mathbb{F}/\Gamma$, one has
\begin{equation*}
\tfev^{\catid{\Gamma},A} \jdeq \tfid{A}
\end{equation*}
\end{enumerate}
\item for any term $f:P$ in context $\ctxext{\Gamma}{A}$, one has
\begin{equation*}
\subst{\lambda_A(f)}{\tfev^{A,P}}\jdeq f.
\end{equation*}
\end{enumerate}
\end{defn}

\begin{rmk}
The evaluation term determines a term
\begin{equation*}
\tilde\tfev^{A,P} : \thom{\ctxwk{A}{\Pi_A(P)}}{P}
\end{equation*}
\end{rmk}

\begin{defn}
A $\Pi$E-homomorphism from $\mathbb{E}$ to $\mathbb{E}'$ is an E-homomorphism
$H:\mathbb{E}\to\mathbb{E}'$ such that
\begin{enumerate}
\item for any $A\in\cat{F}/\Gamma$, one has
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{F}/\ctxext{\Gamma}{A} \arrow[d,swap,"\Pi_A"] \arrow[r,"H/\ctxext{\Gamma}{A}"] & \cat{F}'/\ctxext{H(\Gamma)}{H(A)} \arrow[d,"\Pi_{H(A)}"] \\
\cat{F}/\Gamma \arrow[r,swap,"H/\Gamma"] & \cat{F}'/F(\Gamma)
\end{tikzcd}
\end{equation*}
\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$, one has
\begin{equation*}
H(\tfev^{A,P})\jdeq \tfev^{H(A),H(P)}
\end{equation*}
\end{enumerate}
\end{defn}
\end{comment}

\bibliographystyle{plain}
%\phantomsection\addcontentsline{toc}{section}{References}
\bibliography{References/refs}

\end{document}